---
title: "other_scripts"
author: "Johan Zicola"
date: '2025-01-16'
output: html_document
---


# Remove empty reads

```{bash, eval=FALSE}

while read i; do
  srun -p medium -t 01:20:00 --cpus-per-task=10 cutadapt -j 10 --minimum-length 25 \
  -o ${i}_L1_1.trimmed.min25.fq -p ${i}_L1_2.trimmed.min25.fq \
  ${i}_L1_1.trimmed.fq ${i}_L1_2.trimmed.fq &
done < list.txt

```

# Trimming with trimmomatic

```{bash, eval=FALSE}

module load rev/21.12 trimmomatic openjdk

PATH_ADAPTER="/usr/product/bioinfo/SL_7.0/BIOINFORMATICS/TRIMMOMATIC/0.36/adapters/"

srun -p medium -t 00:30:00 --mem=12G --cpus-per-task=10 \
java -Xmx10G -jar /usr/product/bioinfo/SL_7.0/BIOINFORMATICS/TRIMMOMATIC/0.36/trimmomatic-0.36.jar \
PE -phred33 -threads 10 \
raw_data/01_Ad/01_Ad_L1_1.fq.gz \
raw_data/01_Ad/01_Ad_L1_2.fq.gz \
trimmed_reads_trimmomatic/01_Ad_L1_1.paired.fq.gz \
trimmed_reads_trimmomatic/01_Ad_L1_1.unpaired.fq.gz \
trimmed_reads_trimmomatic/01_Ad_L1_2.paired.fq.gz \
trimmed_reads_trimmomatic/01_Ad_L1_2.unpaired.fq.gz \
ILLUMINACLIP:"/usr/product/bioinfo/SL_7.0/BIOINFORMATICS/TRIMMOMATIC/0.36/adapters/TruSeq2-SE.fa":2:30:10 LEADING:3 TRAILING:3 SLIDINGWINDOW:4:15 MINLEN:50 &

Input Read Pairs: 64769548 Both Surviving: 63780516 (98.47%) Forward Only Surviving: 228552 (0.35%) Reverse Only Surviving: 744814 (1.15%) Dropped: 15666 (0.02%)
TrimmomaticPE: Completed successfully

seqkit stats 01_Ad_L1_1.paired.fq.gz 01_Ad_L1_2.paired.fq.gz
processed files:  2 / 2 [======================================] ETA: 0s. done
file                     format  type    num_seqs        sum_len  min_len  avg_len  max_len
01_Ad_L1_1.paired.fq.gz  FASTQ   DNA   63,780,516  6,351,162,309       50     99.6      100
01_Ad_L1_2.paired.fq.gz  FASTQ   DNA   63,780,516  6,350,850,563       50     99.6      100

# Same number of reads in each file

module load fastqc/0.11.4

srun -p medium -t 00:50:00 fastqc 01_Ad_L1_2.paired.fq.gz &
srun -p medium -t 00:50:00 fastqc 01_Ad_L1_1.paired.fq.gz &

```


# Remove rRNA

https://github.com/sortmerna/sortmerna

Download v4.3 database (https://github.com/biocore/sortmerna/releases/download/v4.3.4/database.tar.gz)

## Index database

The docu recommends to use `smr_v4.3_default_db.fasta`

```{r}
mkdir ~/genomes/silva_rRNA_maize/smr_database && cd ~/genomes/silva_rRNA_maize/smr_database

wget https://github.com/biocore/sortmerna/releases/download/v4.3.4/database.tar.gz

tar -zxvf database.tar.gz

seqkit stats smr_v4.3_default_db.fasta
file                       format  type  num_seqs      sum_len  min_len  avg_len  max_len
smr_v4.3_default_db.fasta  FASTA   RNA     73,977  133,533,894       91  1,805.1    4,380

module load rev/21.12 sortmerna/20170809

srun -t 00:50:00 --mem=20G -p medium indexdb_rna --ref smr_v4.3_default_db.fasta,smr_v4.3_default_db.idx -v

```


```{bash, eval=FALSE}
\ls -1sh smr_v4.3_default*
140M smr_v4.3_default_db.fasta
662M smr_v4.3_default_db.idx.bursttrie_0.dat
1.0M smr_v4.3_default_db.idx.kmer_0.dat
1.2G smr_v4.3_default_db.idx.pos_0.dat
4.0M smr_v4.3_default_db.idx.stats
```

4 index files are generated with a *dat extension.

## Merge reads

I tested with the R1 and R2 in separate files using two --read flags but the output contained only R2 sequences. When I merge the R1 and R2, it works. I just need to merge R1 and R2 and unmerge them afterwards

```{bash, eval=FALSE}

cd /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trimmed_reads

mkdir merged_fastq

module load rev/21.12 sortmerna/20170809

while read i; do
if [ -e merged_fastq/${i}.merged.fq ]; then
  echo "$i already processed"
else
  srun -p medium --mem=2Gb --time=00:20:00 bash /usr/product/bioinfo/SL_7.0/BIOINFORMATICS/SORTMERNA/sortmerna_20170809/scripts/merge-paired-reads.sh ${i}_L1_1.trimmed.fq ${i}_L1_2.trimmed.fq  merged_fastq/${i}.merged.fq &
fi
done < list.txt


```


```{bash, eval=FALSE}
#!/bin/bash
#SBATCH --job-name=merge_fastq
#SBATCH --nodes=1
#SBATCH --time=00:40:00
#SBATCH --cpus-per-task=1
#SBATCH --mem=10gb
#SBATCH --partition=medium
#SBATCH --mail-type=END
#SBATCH --mail-user=johan.zicola@uni-goettingen.de
#SBATCH --output=sortmerna.%J.out
#SBATCH --error=sortmerna.%J.err
#SBATCH --array=1-8

mem_gbytes=$(( $SLURM_MEM_PER_NODE/1024 ))

echo "Starting at $(date)"
echo "Job submitted to the ${SLURM_JOB_PARTITION} partition, the default partition on ${SLURM_CLUSTER_NAME}"
echo "Job name: ${SLURM_JOB_NAME}, Job ID: ${SLURM_JOB_ID}"
echo "I have ${SLURM_CPUS_ON_NODE} CPUs and ${mem_gbytes}GiB of RAM on compute node $(hostname)"

module purge

module load rev/21.12 sortmerna/20170809

i=$(sed -n ${SLURM_ARRAY_TASK_ID}p trimmed_reads/list.txt)

bash /usr/product/bioinfo/SL_7.0/BIOINFORMATICS/SORTMERNA/sortmerna_20170809/scripts/merge-paired-reads.sh trimmed_reads/${i}_L1_1.trimmed.fq trimmed_reads/${i}_L1_2.trimmed.fq  trimmed_reads/merged_fastq/${i}.merged.fq

```


## sortmerna

```{r}
#!/bin/bash
#SBATCH --job-name=sortmerna
#SBATCH --nodes=1
#SBATCH --time=01:00:00
#SBATCH --cpus-per-task=10
#SBATCH --mem=50gb
#SBATCH --partition=medium
#SBATCH --mail-type=END
#SBATCH --mail-user=johan.zicola@uni-goettingen.de
#SBATCH --output=sortmerna.%J.out
#SBATCH --error=sortmerna.%J.err
#SBATCH --array=1-8

mem_gbytes=$(( $SLURM_MEM_PER_NODE/1024 ))

echo "Starting at $(date)"
echo "Job submitted to the ${SLURM_JOB_PARTITION} partition, the default partition on ${SLURM_CLUSTER_NAME}"
echo "Job name: ${SLURM_JOB_NAME}, Job ID: ${SLURM_JOB_ID}"
echo "I have ${SLURM_CPUS_ON_NODE} CPUs and ${mem_gbytes}GiB of RAM on compute node $(hostname)"

module purge

module load rev/21.12 sortmerna/20170809

PATH_DB="/usr/users/zicola/genomes/silva_rRNA_maize/smr_database"

i=$(sed -n ${SLURM_ARRAY_TASK_ID}p trimmed_reads/list.txt)

sortmerna -a 1 --paired_in --fastx \
    --ref ${PATH_DB}/smr_v4.3_default_db.fasta,${PATH_DB}/smr_v4.3_default_db.idx \
    --reads trimmed_reads/merged_fastq/${i}.merged.fq \
    --aligned rRNA_removed/${i}.aligned \
    --other rRNA_removed/${i}.mRNA

```

```{bash, eval=FALSE}
srun --pty -p medium --mem=50G --cpus-per-task=5 --time=01:00:00  bash

i="02_Pm"

module purge

module load rev/21.12 sortmerna/20170809

PATH_DB="/usr/users/zicola/genomes/silva_rRNA_maize/smr_database"

sortmerna -a 5 --paired_in --fastx \
    --ref ${PATH_DB}/smr_v5.3_default_db.fasta,${PATH_DB}/smr_v4.3_default_db.idx \
    --reads trimmed_reads/merged_fastq/${i}.merged.fq \
    --aligned rRNA_removed/${i}.aligned \
    --other rRNA_removed/${i}.mRNA

```

Not one read mapped to the smr_v4.3_default_db database. This is very suspicious.


```{bash, eval=FALSE}

i="08_Ar"

sortmerna -a 5 --paired_in --fastx \
    --ref ${PATH_DB}/smr_v4.3_default_db.fasta,${PATH_DB}/smr_v4.3_default_db.idx \
    --reads trimmed_reads/merged_fastq/${i}.merged.fq \
    --aligned rRNA_removed/${i}.aligned \
    --other rRNA_removed/${i}.mRNA

```

Still nothing mapping to rRNA database.



## Unmerge fastq

```{bash, eval=FALSE}
#!/bin/bash
#SBATCH --job-name=unmerge_fastq
#SBATCH --nodes=1
#SBATCH --time=00:40:00
#SBATCH --cpus-per-task=1
#SBATCH --mem=20gb
#SBATCH --partition=medium
#SBATCH --mail-type=END
#SBATCH --mail-user=johan.zicola@uni-goettingen.de
#SBATCH --output=sortmerna.%J.out
#SBATCH --error=sortmerna.%J.err
#SBATCH --array=1-8

mem_gbytes=$(( $SLURM_MEM_PER_NODE/1024 ))

echo "Starting at $(date)"
echo "Job submitted to the ${SLURM_JOB_PARTITION} partition, the default partition on ${SLURM_CLUSTER_NAME}"
echo "Job name: ${SLURM_JOB_NAME}, Job ID: ${SLURM_JOB_ID}"
echo "I have ${SLURM_CPUS_ON_NODE} CPUs and ${mem_gbytes}GiB of RAM on compute node $(hostname)"

module purge

module load rev/21.12 sortmerna/20170809

i=$(sed -n ${SLURM_ARRAY_TASK_ID}p trimmed_reads/list.txt)

bash /usr/product/bioinfo/SL_7.0/BIOINFORMATICS/SORTMERNA/sortmerna_20170809/scripts/merge-paired-reads.sh rRNA_removed/${i}.mRNA.fq rRNA_removed/${i}.R1.mRNA.fq rRNA_removed/${i}.R2.mRNA.fq


```

## Use bowtie2

Index the database from sortmerna

```{bash, eval=FALSE}
cd /usr/users/zicola/genomes/silva_rRNA_maize/smr_database

bowtie2-build -f smr_v4.3_default_db.fasta smr_v4_3_default_db
```

```{bash, eval=FALSE}
#!/bin/bash
#SBATCH --job-name=rRNA_removal
#SBATCH --nodes=1
#SBATCH --time=02:00:00
#SBATCH --cpus-per-task=10
#SBATCH --mem=50gb
#SBATCH --partition=medium
#SBATCH --mail-type=END
#SBATCH --mail-user=johan.zicola@uni-goettingen.de
#SBATCH --output=rRNA_removal.%J.out
#SBATCH --error=rRNA_removal.%J.err
#SBATCH --array=1-8

mem_gbytes=$(( $SLURM_MEM_PER_NODE/1024 ))

echo "Starting at $(date)"
echo "Job submitted to the ${SLURM_JOB_PARTITION} partition, the default partition on ${SLURM_CLUSTER_NAME}"
echo "Job name: ${SLURM_JOB_NAME}, Job ID: ${SLURM_JOB_ID}"
echo "I have ${SLURM_CPUS_ON_NODE} CPUs and ${mem_gbytes}GiB of RAM on compute node $(hostname)"

module purge
module load rev/21.12 bowtie2

i=$(sed -n ${SLURM_ARRAY_TASK_ID}p trimmed_reads/list.txt)

bowtie2 -p 10 --nofw --quiet --very-sensitive-local --phred33  \
    -x /usr/users/zicola/genomes/silva_rRNA_maize/smr_database/smr_v4_3_default_db \
    -1 trimmed_reads/${i}_L1_1.trimmed.fq  \
    -2 trimmed_reads/${i}_L1_2.trimmed.fq \
    --un-conc rRNA_removed/${i}.rRNA_removed.fq > /dev/null

```


```{bash, eval=FALSE}

# Trimmed reads
seqkit stats 01_Ad*
processed files:  2 / 2 [======================================] ETA: 0s. done
file                   format  type    num_seqs        sum_len  min_len  avg_len  max_len
01_Ad_L1_1.trimmed.fq  FASTQ   DNA   63,577,851  6,342,339,512        1     99.8      100
01_Ad_L1_2.trimmed.fq  FASTQ   DNA   63,577,851  6,340,577,359        1     99.7      100


# Trimmed reads and rRNA removal

seqkit stats 01_Ad*
processed files:  2 / 2 [======================================] ETA: 0s. done
file                     format  type    num_seqs        sum_len  min_len  avg_len  max_len
01_Ad.rRNA_removed.1.fq  FASTQ   DNA   63,572,322  6,341,800,686        1     99.8      100
01_Ad.rRNA_removed.2.fq  FASTQ   DNA   63,572,322  6,340,036,951        1     99.7      100


```

That is 5529 (63577851-63572322) reads removed out of 63M (.00877619%) so I really doubt rRNA contamination would affect the assembly.



Do on all samples

```{bash, eval=FALSE}
cat stat.txt
processed files:  16 / 16  ETA: 0s. done
file                      format  type    num_seqs        sum_len  min_len  avg_len  max_len
01_Ad.rRNA_removed.1.fq   FASTQ   DNA   63,572,322  6,341,800,686        1     99.8      100
01_Ad.rRNA_removed.2.fq   FASTQ   DNA   63,572,322  6,340,036,951        1     99.7      100
02_Pm.rRNA_removed.1.fq   FASTQ   DNA   55,259,617  5,507,706,345       25     99.7      100
02_Pm.rRNA_removed.2.fq   FASTQ   DNA   55,259,617  5,505,907,480       25     99.6      100
03_Zt.rRNA_removed.1.fq   FASTQ   DNA   49,126,313  4,904,813,047       25     99.8      100
03_Zt.rRNA_removed.2.fq   FASTQ   DNA   49,126,313  4,903,228,530       25     99.8      100
04_Asp.rRNA_removed.1.fq  FASTQ   DNA   41,798,830  4,169,156,197       25     99.7      100
04_Asp.rRNA_removed.2.fq  FASTQ   DNA   41,798,830  4,168,954,996       25     99.7      100
05_G.rRNA_removed.1.fq    FASTQ   DNA   40,895,705  4,078,580,793       25     99.7      100
05_G.rRNA_removed.2.fq    FASTQ   DNA   40,895,705  4,077,039,215       25     99.7      100
06_Ol.rRNA_removed.1.fq   FASTQ   DNA   56,412,659  5,630,612,260       25     99.8      100
06_Ol.rRNA_removed.2.fq   FASTQ   DNA   56,412,659  5,628,472,850       25     99.8      100
07_Pr.rRNA_removed.1.fq   FASTQ   DNA   47,066,726  4,697,862,680       25     99.8      100
07_Pr.rRNA_removed.2.fq   FASTQ   DNA   47,066,726  4,696,252,248       25     99.8      100
08_Ar.rRNA_removed.1.fq   FASTQ   DNA   43,705,177  4,354,160,178       25     99.6      100
08_Ar.rRNA_removed.2.fq   FASTQ   DNA   43,705,177  4,352,752,476       25     99.6      100

```

Compare with number of reads after trimming

| sample      | reads after trimming | reads after bowtie2 rRNA filtering | difference | % trimmed reads |
|-------------|----------------------|------------------------------------|------------|-----------------|
| 01_Ad_L1_1  | 63,577,928           | 63,572,322                         | 5,606      | 0.0088%         |
| 01_Ad_L1_2  | 63,577,928           | 63,572,322                         | 5,606      | 0.0088%         |
| 02_Pm_L1_1  | 55,269,387           | 55,259,617                         | 9,770      | 0.0177%         |
| 02_Pm_L1_2  | 55,269,387           | 55,259,617                         | 9,770      | 0.0177%         |
| 03_Zt_L1_1  | 49,128,635           | 49,126,313                         | 2,322      | 0.0047%         |
| 03_Zt_L1_2  | 49,128,635           | 49,126,313                         | 2,322      | 0.0047%         |
| 04_Asp_L1_1 | 41,800,974           | 41,798,830                         | 2,144      | 0.0051%         |
| 04_Asp_L1_2 | 41,800,974           | 41,798,830                         | 2,144      | 0.0051%         |
| 05_G_L1_1   | 40,906,340           | 40,895,705                         | 10,635     | 0.0260%         |
| 05_G_L1_2   | 40,906,340           | 40,895,705                         | 10,635     | 0.0260%         |
| 06_Ol_L1_1  | 56,417,880           | 56,412,659                         | 5,221      | 0.0093%         |
| 06_Ol_L1_2  | 56,417,880           | 56,412,659                         | 5,221      | 0.0093%         |
| 07_Pr_L1_1  | 47,070,196           | 47,066,726                         | 3,470      | 0.0074%         |
| 07_Pr_L1_2  | 47,070,196           | 47,066,726                         | 3,470      | 0.0074%         |
| 08_Ar_L1_1  | 43,759,260           | 43,705,177                         | 54,083     | 0.1236%         |
| 08_Ar_L1_2  | 43,759,260           | 43,705,177                         | 54,083     | 0.1236%         |

The rRNA contamination


## Use bowtie2 on Arb Silva Coleoptera database

Search for taxonomy Coleoptera. Download fasta without gaps for the LSU and SSU hits (https://www.arb-silva.de/download)

+ 13765 SSU
+ 8776 LSU

```{bash, eval=FALSE}

cd /usr/users/zicola/genomes/rRNA_coleoptera

tar -zxvf arb-silva.de_2025-01-16_id1372508.tgz
mv arb-silva.de_2025-01-16_id1372508_tax_silva_trunc.fasta LSU_coleoptera.fasta

tar -zxvf arb-silva.de_2025-01-16_id1372510.tgz
mv arb-silva.de_2025-01-16_id1372510_tax_silva_trunc.fasta SSU_coleoptera.fasta

seqkit stats *fasta
processed files:  2 / 2 [======================================] ETA: 0s. done
file                  format  type  num_seqs     sum_len  min_len  avg_len  max_len
LSU_coleoptera.fasta  FASTA   RNA      8,776   5,944,179      300    677.3    3,238
SSU_coleoptera.fasta  FASTA   RNA     13,768  16,277,792      300  1,182.3    2,645

cat LSU_coleoptera.fasta SSU_coleoptera.fasta > LSU_SSU_coleoptera.fasta

seqkit stats LSU_SSU_coleoptera.fasta
file                      format  type  num_seqs     sum_len  min_len  avg_len  max_len
LSU_SSU_coleoptera.fasta  FASTA   RNA     22,544  22,221,971      300    985.7    3,238


bowtie2-build -f LSU_SSU_coleoptera.fasta LSU_SSU_coleoptera

```

Test with Ad

```{bash, eval=FALSE}
#!/bin/bash
#SBATCH --job-name=rRNA_removal
#SBATCH --nodes=1
#SBATCH --time=02:00:00
#SBATCH --cpus-per-task=10
#SBATCH --mem=50gb
#SBATCH --partition=medium
#SBATCH --mail-type=END
#SBATCH --mail-user=johan.zicola@uni-goettingen.de
#SBATCH --output=rRNA_removal.%J.out
#SBATCH --error=rRNA_removal.%J.err

mem_gbytes=$(( $SLURM_MEM_PER_NODE/1024 ))

echo "Starting at $(date)"
echo "Job submitted to the ${SLURM_JOB_PARTITION} partition, the default partition on ${SLURM_CLUSTER_NAME}"
echo "Job name: ${SLURM_JOB_NAME}, Job ID: ${SLURM_JOB_ID}"
echo "I have ${SLURM_CPUS_ON_NODE} CPUs and ${mem_gbytes}GiB of RAM on compute node $(hostname)"

module purge
module load rev/21.12 bowtie2

bowtie2 -p 10 --nofw --quiet --very-sensitive-local --phred33  \
    -x /usr/users/zicola/genomes/rRNA_coleoptera/LSU_SSU_coleoptera \
    -1 trimmed_reads/01_Ad_L1_1.trimmed.fq  \
    -2 trimmed_reads/01_Ad_L1_2.trimmed.fq \
    --un-conc rRNA_removed/01_Ad.rRNA_removed_silva_coleptera.fq > /dev/null

```

```{bash, eval=FALSE}
sbatch test_bowtie2.sh
```


```{bash, eval=FALSE}
# Bowtie using smr database

seqkit stats 01_Ad*
processed files:  2 / 2 [======================================] ETA: 0s. done
file                     format  type    num_seqs        sum_len  min_len  avg_len  max_len
01_Ad.rRNA_removed.1.fq  FASTQ   DNA   63,572,322  6,341,800,686        1     99.8      100
01_Ad.rRNA_removed.2.fq  FASTQ   DNA   63,572,322  6,340,036,951        1     99.7      100


# Bowtie using ARB Silva Coleoptera SSU + LSU database

seqkit stats 01_Ad.rRNA_removed_silva_coleptera.*
processed files:  2 / 2 [======================================] ETA: 0s. done
file                                     format  type    num_seqs        sum_len  min_len  avg_len  max_len
01_Ad.rRNA_removed_silva_coleptera.1.fq  FASTQ   DNA   63,574,766  6,342,041,048        1     99.8      100
01_Ad.rRNA_removed_silva_coleptera.2.fq  FASTQ   DNA   63,574,766  6,340,277,246        1     99.7      100


```

63574766-63572322 = 2444. I can actually less reads mapping rRNA when using the Silva database for Coleptera (22,544 sequences) compared to the smr default database (73,977 sequences).



## test sortmerna

```{bash, eval=FALSE}

# Take first entry of smr_v4.3_default_db.fasta and put it in seq_rRNA.fa


cd /usr/users/zicola/genomes/sortmerna_database/test

python ~/git_repositories/fastq_generator/fastq_generator.py generate_mapped_fastq_PE seq_rRNA.fa 50 10 10 > fastq_interleaved.fq

# Split in two files
seqkit stats fastq_interleaved.fq
file                  format  type  num_seqs  sum_len  min_len  avg_len  max_len
fastq_interleaved.fq  FASTQ   DNA        520   26,000       50       50       50

head -n 1040 fastq_interleaved.fq > fastq_interleaved.R1.fq
tail -n 1040 fastq_interleaved.fq > fastq_interleaved.R2.fq

module load rev/21.12 sortmerna/20170809

# Get them merged (each pair is together)
/usr/product/bioinfo/SL_7.0/BIOINFORMATICS/SORTMERNA/sortmerna_20170809/scripts/merge-paired-reads.sh fastq_interleaved.R1.fq fastq_interleaved.R2.fq  fastq_interleaved.merged.fq

sortmerna -a 1 --paired_in --fastx \
    --ref ../smr_v4.3_default_db.fasta,../smr_v4.3_default_db.idx \
    --reads fastq_interleaved.merged.fq \
    --aligned aligned \
    --other mRNA -v
# Nothing aligned

sortmerna -a 1 --fastx \
    --ref ../smr_v4.3_default_db.fasta,../smr_v4.3_default_db.idx \
    --reads fastq_interleaved.R1.fq --reads fastq_interleaved.R2.fq \
    --aligned aligned \
    --other mRNA -v




sortmerna -a 1 --fastx \
    --ref ../smr_v4.3_default_db.fasta,../smr_v4.3_default_db.idx \
    --reads fastq_interleaved.merged.fq \
    --aligned aligned \
    --other mRNA -v


sortmerna -a 1 --fastx \
    --ref ../smr_v4.3_default_db.fasta,../smr_v4.3_default_db.idx \
    --reads fastq_interleaved.R1.fq \
    --aligned aligned \
    --other mRNA -v
    
    
sortmerna -a 1 --fastx \
    --ref ../smr_v4.3_default_db.fasta,../smr_v4.3_default_db.idx \
    --reads tests/data/set2_environmental_study_550_amplicon.fasta \
    --aligned aligned \
    --other mRNA -v
    
sortmerna -a 1 --fastx \
    --ref ../smr_v4.3_default_db.fasta,../smr_v4.3_default_db.idx \
    --reads tests/data/set2_environmental_study_550_amplicon.fasta \
    --aligned aligned \
    --other mRNA -v
    
```

Not working

```{bash, eval=FALSE}
module load rev/21.12 sortmerna/20170809

indexdb --ref seq_rRNA.fa,seq_rRNA.idx -v

sortmerna -a 1 --fastx \
    --ref seq_rRNA.fa,seq_rRNA.idx \
    --reads fastq_interleaved.R1.fq \
    --aligned aligned \
    --other mRNA -v

# Now it worked! Everything is in aligned.fq

sortmerna -a 1 --fastx \
    --ref smr_v4.3_default_db.fasta,smr_v4.3_default_db.idx \
    --reads test/fastq_interleaved.R1.fq \
    --aligned aligned \
    --other mRNA -v

# Not working
```

I think the DB is corrupted.

Remake the indexes of smr_v4.3_default_db.fasta

```{bash, eval=FALSE}

srun --pty -t 00:50:00 --mem=20G -p medium bash

module load rev/21.12 sortmerna/20170809

indexdb --ref smr_v4.3_default_db.fasta,smr_v4.3_default_db.idx -v

```

```{bash, eval=FALSE}
# Single end mode
sortmerna -a 1 --fastx \
    --ref smr_v4.3_default_db.fasta,smr_v4.3_default_db.idx \
    --reads test/fastq_interleaved.R1.fq \
    --aligned aligned \
    --other mRNA -v
```

Now it works!

```{bash, eval=FALSE}
# Paired end mode
sortmerna -a 1 --paired_in --fastx \
    --ref ../smr_v4.3_default_db.fasta,../smr_v4.3_default_db.idx \
    --reads fastq_interleaved.merged.fq \
    --aligned aligned \
    --other mRNA -v
```

Works

Redo a test to see the fraction of reads removed.


### Test Anchomenus dorsalis 


```{bash, eval=FALSE}
srun --pty -p medium --mem=40G --time=04:00:00  bash

module load rev/21.12 sortmerna/20170809

cd /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trimmed_reads
mkdir test_sortmerna
gunzip  01_Ad_*

# Get them merged (each pair is together)
/usr/product/bioinfo/SL_7.0/BIOINFORMATICS/SORTMERNA/sortmerna_20170809/scripts/merge-paired-reads.sh 01_Ad_L1_1.trimmed.fq 01_Ad_L1_2.trimmed.fq  test_sortmerna/01_Ad_L1.merged.fq

cd test_sortmerna

DB_sortmerna="/usr/users/zicola/genomes/sortmerna_database"

sortmerna -a 1 --paired_in --fastx \
    --ref ${DB_sortmerna}/smr_v4.3_default_db.fasta,${DB_sortmerna}/smr_v4.3_default_db.idx \
    --reads 01_Ad_L1.merged.fq \
    --aligned 01_Ad_L1_aligned \
    --other 01_Ad_L1_mRNA -v

# Reseparate strands

```


# Trinotate

Wiki: https://github.com/Trinotate/Trinotate/wiki/Software-installation-and-data-required


Reuse script used for OPM analysis.


```{bash, eval=FALSE}

#!/bin/bash
#SBATCH --job-name=trinotate
#SBATCH --nodes=1
#SBATCH --partition=medium
#SBATCH --time=48:00:00
#SBATCH --cpus-per-task=8
#SBATCH --mem=40gb
#SBATCH --mail-type=END
#SBATCH --mail-user=johan.zicola@uni-goettingen.de
#SBATCH --output=trinotate.%J.out
#SBATCH --error=trinotate.%J.err


mem_gbytes=$(( $SLURM_MEM_PER_NODE/1024 ))

echo "Starting at $(date)"
echo "Job submitted to the ${SLURM_JOB_PARTITION} partition, the default partition on ${SLURM_CLUSTER_NAME}"
echo "Job name: ${SLURM_JOB_NAME}, Job ID: ${SLURM_JOB_ID}"
echo "I have ${SLURM_CPUS_ON_NODE} CPUs and ${mem_gbytes}GiB of RAM on compute node $(hostname)"

# Modules to load
module purge

module load rev/21.12 gcc/10.2.0 hmmer/3.3 ncbi-blast/2.11.0

# Test that all programs needed are installed
test_program(){
  command -v $1 >/dev/null || \
    { echo >&2 " $1 required but not installed. Aborting"; exit 1; }
}

test_program "diamond"
test_program "cmpress"
test_program "hmmpress"
test_program "tmhmm"
test_program "signalp6"
test_program "makeblastdb"

# Variables passed to sbatch
# NAME_SPECIES
# Output from Trinity $GENE_TRANS_MAP and $TRINITY_FASTA
# Output from Transdecoder $TRANSDECODER_PEP

# Variables
TRINOTATE_HOME="/usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1"

#NAME_SPECIES is provided as argument to sbatch
sqlite_db="/usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1/Trinotate_${NAME_SPECIES}.sqlite"


echo "############################"
echo "Trinotate create"
date
echo "############################"

$TRINOTATE_HOME/Trinotate --create \
    --db $sqlite_db \
    --trinotate_data_dir $TRINOTATE_HOME/TRINOTATE_DATA_DIR \
    --use_diamond

echo "###############################"
echo "Trinotate Init"
date
echo "###############################"


$TRINOTATE_HOME/Trinotate --db $sqlite_db --init \
           --gene_trans_map $GENE_TRANS_MAP \
           --transcript_fasta $TRINITY_FASTA \
           --transdecoder_pep $TRANSDECODER_PEP

echo "##############################"
echo  "Trinotate Run"
date
echo "##############################"
 
$TRINOTATE_HOME/Trinotate --db ${sqlite_db} --CPU ${SLURM_CPUS_ON_NODE} --run ALL \
  --trinotate_data_dir $TRINOTATE_HOME/TRINOTATE_DATA_DIR \
    --transcript_fasta $TRINITY_FASTA \
      --transdecoder_pep $TRANSDECODER_PEP \
      --use_diamond

echo "###########################"
echo "Generating report table"
date
echo "###########################"

$TRINOTATE_HOME/Trinotate --db ${sqlite_db} \
  --report --incl_pep --incl_trans > Trinotate_report.tsv

##################
## Misc value adds
##################

# Extract GO terms
${TRINOTATE_HOME}/util/extract_GO_assignments_from_Trinotate_xls.pl \
  --Trinotate_xls Trinotate_report.tsv -G -I > Trinotate_report.tsv.gene_ontology

# Generate trinotate report summary statistics
${TRINOTATE_HOME}/util/report_summary/trinotate_report_summary.pl \
  Trinotate_report.tsv Trinotate_report_stats

echo "##########################"
echo "done.  See annotation summary file:  Trinotate_report.tsv"
date
echo "##########################"



```

```{bash, eval=FALSE}


# Variables passed to sbatch
# Output from Trinity $GENE_TRANS_MAP and $TRINITY_FASTA
# Output from Transdecoder $TRANSDECODER_PEP

work_path="/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects"
while read i; do
  mkdir $i
  cd $i 
  sbatch --export=NAME_SPECIES=$i,GENE_TRANS_MAP=${work_path}/trinity_output/summary/${i}_trinity.fasta.gene_trans_map,TRINITY_FASTA=${work_path}/trinity_output/summary/${i}_trinity.fasta,TRANSDECODER_PEP=${work_path}/transdecoder_output/${i}_trinity.fasta.transdecoder.pep ../04_trinotate.sh
  cd ..
done < list_species.txt

```


Steps done before being killed

```{bash, eval=FALSE}

Running CMD: /usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1/Trinotate --db /usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1/Trinotate_Anchomenus_dorsalis.sqlite --LOAD_swissprot_blastp uniprot_sprot.diamond.blastp.outfmt6
-LOADING as per --LOAD_swissprot_blastp

CMD: /home/uni08/zicola/bin/Trinotate-Trinotate-v4.0.1/util/trinotateSeqLoader/Trinotate_BLAST_loader.pl --sqlite /usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1/Trinotate_Anchomenus_dorsalis.sqlite --outfmt6 uniprot_sprot.diamond.blastp.outfmt6 --prog blastp --dbtype Swissprot
Can't locate DBI.pm in @INC (you may need to install the DBI module) (@INC contains: /home/uni08/zicola/bin/Trinotate-Trinotate-v4.0.1/util/trinotateSeqLoader/../../PerlLib /usr/users/zicola/perl5/lib/perl5/x86_64-linux-thread-multi /usr/users/zicola/perl5/lib/perl5 /usr/local/lib64/perl5 /usr/local/share/perl5 /usr/lib64/perl5/vendor_perl /usr/share/perl5/vendor_perl /usr/lib64/perl5 /usr/share/perl5) at /home/uni08/zicola/bin/Trinotate-Trinotate-v4.0.1/util/trinotateSeqLoader/Trinotate_BLAST_loader.pl line 9.
BEGIN failed--compilation aborted at /home/uni08/zicola/bin/Trinotate-Trinotate-v4.0.1/util/trinotateSeqLoader/Trinotate_BLAST_loader.pl line 9.
Error, cmd: /home/uni08/zicola/bin/Trinotate-Trinotate-v4.0.1/util/trinotateSeqLoader/Trinotate_BLAST_loader.pl --sqlite /usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1/Trinotate_Anchomenus_dorsalis.sqlite --outfmt6 uniprot_sprot.diamond.blastp.outfmt6 --prog blastp --dbtype Swissprot died with ret 512 at /usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1/Trinotate line 338.
Error, cmd: /usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1/Trinotate --db /usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1/Trinotate_Anchomenus_dorsalis.sqlite --LOAD_swissprot_blastp uniprot_sprot.diamond.blastp.outfmt6 died with ret 512 No such file or directory at /home/uni08/zicola/bin/Trinotate-Trinotate-v4.0.1/PerlLib/Pipeliner.pm line 187.
        Pipeliner::run(Pipeliner=HASH(0x56249c3268b0)) called at /usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1/Trinotate line 736
        main::Trinotate_RUN_annotators("/usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1/Trinotate_An"..., ARRAY(0x56249bf92f20), 8, "/usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1/TRINOTATE_DA"..., "/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgs"..., "/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgs"...) called at /usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1/Trinotate line 308
-REPORT being generated.
CMD: /home/uni08/zicola/bin/Trinotate-Trinotate-v4.0.1/util/Trinotate_report_writer.pl --sqlite /usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1/Trinotate_Anchomenus_dorsalis.sqlite -E 1e-5 --pfam_cutoff DNC  --incl_pep  --incl_trans
Can't locate DBI.pm in @INC (you may need to install the DBI module) (@INC contains: /home/uni08/zicola/bin/Trinotate-Trinotate-v4.0.1/util/../PerlLib /usr/users/zicola/perl5/lib/perl5/x86_64-linux-thread-multi /usr/users/zicola/perl5/lib/perl5 /usr/local/lib64/perl5 /usr/local/share/perl5 /usr/lib64/perl5/vendor_perl /usr/share/perl5/vendor_perl /usr/lib64/perl5 /usr/share/perl5) at /home/uni08/zicola/bin/Trinotate-Trinotate-v4.0.1/util/../PerlLib/Sqlite_connect.pm line 15.


```



## Troubleshoot perl library installation

After the migration, I had to reinstall my perl libraries. I notice that some of the libraries I installed before cannot be reinstalled (`cpan install DBI`) due to a `cc1: fatal error: inaccessible plugin file /opt/sw/rev/21.12/haswell/gcc-4.8.5/gcc-10.2.0-potfoy/lib/gcc/x86_64-pc-linux-gnu/10.2.0/plugin/gcc-annobin.so expanded from short plugin name gcc-annobin: No such file or directory` error. I found on the web several posts about the annobin issue and one suggests installing a local Perl installation with Spack (https://github.com/openhpc/ohpc/issues/1364).

I may think this 'annobin' gcc plugin was present on the previous SCC nodes OS (Scientific Linux 7) but was probably removed in the current OS (Rocky Linux 8). It seems that the GCC plugin annobin can be installed on Rocky 8 (https://rockylinux.pkgs.org/8/rockylinux-devel-x86\_64/annobin-11.13-2.el8.x86\_64.rpm.html). Would it be possible to install it or do I have to install my own gcc?

Note that I tried with all versions of gcc available and all failed:

```
module purge
module load rev/20.12 gcc/7.5.0
cpan instal DBI
#cc1: error: unrecognized command line option '-fstack-clash-protection'
#cc1: error: unrecognized command line option '-fcf-protection'
#cc1: fatal error: inaccessible plugin file /opt/sw/rev/20.12/haswell/gcc-9.3.0/gcc-7.5.0-owlabd/lib/gcc/x86_64-pc-linux-gnu/7.5.0/plugin/gcc-annobin.so expanded from short plugin name gcc-annobin: No such file or directory

# Nope
module purge
module load rev/20.12 gcc/8.4.0
cpan instal DBI
#cc1: fatal error: inaccessible plugin file /opt/sw/rev/20.12/haswell/gcc-4.8.5/gcc-8.4.0-anizkf/lib/gcc/x86_64-pc-linux-gnu/8.4.0/plugin/gcc-annobin.so expanded from short plugin name gcc-annobin: No such file or directory

module purge
module load rev/20.12 gcc/9.3.0
cpan instal DBI
# cc1: fatal error: inaccessible plugin file /opt/sw/rev/20.12/haswell/gcc-8.4.0/gcc-9.3.0-a466o5/lib/gcc/x86_64-pc-linux-gnu/9.3.0/plugin/gcc-annobin.so expanded from short plugin name gcc-annobin: No such file or directory

module purge
module load rev/20.12  gcc/9.5.0
cpan instal DBI
# cc1: fatal error: inaccessible plugin file /opt/sw/rev/23.12/linux-scientific7-haswell/gcc-4.8.5/gcc-9.5.0-y5n2mx/lib/gcc/x86_64-pc-linux-gnu/9.5.0/plugin/gcc-annobin.so expanded from short plugin name gcc-annobin: No such file or directory


module purge
module load rev/20.12 gcc/10.2.0
cpan instal DBI
# cc1: fatal error: inaccessible plugin file /opt/sw/rev/20.12/haswell/gcc-8.4.0/gcc-10.2.0-tc66mg/lib/gcc/x86_64-pc-linux-gnu/10.2.0/plugin/gcc-annobin.so expanded from short plugin name gcc-annobin: No such file or directory

module purge
module load gcc/11.4.0
cpan instal DBI
#cc1: fatal error: inaccessible plugin file /opt/sw/rev/23.12/linux-scientific7-haswell/gcc-4.8.5/gcc-11.4.0-jgoe5h/lib/gcc/x86_64-pc-linux-gnu/11.4.0/plugin/gcc-annobin.so expanded from short plugin name gcc-annobin: No such file or directory
```

Red Hat documentation on Annobin: https://docs.redhat.com/en/documentation/red\_hat\_developer\_toolset/9/html/user\_guide/chap-annobin#chap-Annobin


I tried `spack install perl` but the installation failed

```
ImportError: cannot bootstrap the "clingo" Python module from spec "clingo-bootstrap@spack+python %gcc target=x86_64" due to the following failures:
    'spack-install' raised InstallError: Terminating after first install failure: ProcessError: Command exited with status 1:
    '/usr/users/zicola/.spack/bootstrap/store/linux-rocky8-x86_64/gcc-10.2.0/perl-5.34.0-hhpsqrp435rgen7twicgyvpz5tpy5tbs/bin/perl' 'Makefile.PL'
    Please run `spack -d spec zlib` for more verbose error messages
```

I tried several times `spack -d spec zlib` but it does not work.



Mnay posts about the issue with annobin. One recommendation is to install perl in local using spack.


Download archive https://metacpan.org/pod/DBI and install manually DBI-1.647.tgz


```{bash, eval=FALSE}
cd ~/bin
tar -zxvf DBI-1.647.tgz
cd DBI-1.647

perl Makefile.PL PREFIX=/usr/users/zicola/perl5/lib/perl5
make

#cc1: fatal error: inaccessible plugin file /opt/sw/rev/23.12/linux-scientific7-haswell/gcc-4.8.5/gcc-11.4.0-jgoe5h/lib/gcc/x86_64-pc-linux-gnu/11.4.0/plugin/gcc-annobin.so expanded from short plugin name gcc-annobin: No such file or directory

#make test
#make install

```


```{bash, eval=FALSE}

spack install perl

# Error message, recommends to type 'spack -d spec zlib'
spack -d spec zlib


```


### Solution => Use apptainer


# Test apptainer

sbatch script `sbatch_apptainer.sh`. Launch from dir `/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects`

```{bash, eval=FALSE}


#!/bin/bash
#SBATCH --job-name=trinotate
#SBATCH --nodes=1
#SBATCH --partition=medium
#SBATCH --time=00:05:00
#SBATCH --cpus-per-task=1
#SBATCH --mem=1G
#SBATCH --output=trinotate.%J.out
#SBATCH --error=trinotate.%J.err

module load apptainer

# Pass to apptainer the variable NEWHOME
export APPTAINERENV_NEWHOME=/usr/users/zicola/newhome
export APPTAINERENV_SPECIES=$NAMESPECIES

apptainer run -C -B /usr/users/npghpc/ \
   -B /usr/users/zicola /usr/users/zicola/bin/trinotate.v4.0.2.simg /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/test_apptainer.sh

```


shell script `test_apptainer.sh`

```{bash, eval=FALSE}

export HOME=$NEWHOME
export TRINOTATE_HOME=/usr/local/src/Trinotate
cd $HOME
pwd > output_pwd.txt

#${TRINOTATE_HOME}/Trinotate --help > help_trinotate.txt

```

Make the script executable `chmod +x test_apptainer.sh`

```{bash, eval=FALSE}

sbatch sbatch_apptainer.sh

cat ~/newhome/output_pwd.txt
/usr/users/zicola/newhome

# Worked
```




##Simple functions


Way to call Trinotate

```{bash, eval=FALSE}

module load apptainer

PATH_DIR="/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects"

export TRINOTATE_HOME=/usr/local/src/Trinotate

apptainer exec --bind $PATH_DIR -e ~/bin/trinotate.v4.0.2.simg $TRINOTATE_HOME/Trinotate 


# Direct run within the container
apptainer shell --bind "/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects" ~/bin/trinotate.v4.0.2.simg

cd ~
apptainer shell ~/bin/trinotate.v4.0.2.simg

```


Add extra step for tmhmm and signal in the pipeline used for OPM


## Create database

What this step does it just copy the `TrinotateBoilerplate.sqlite` (340 Mb) file from TRINOTATE_DATA_DIR into the name and folder we specify for the our database. Include --use_diamond if you plan to use diamond blast with Trinotate, otherwise NCBI blast+ will be used for database preparation.

https://github.com/Trinotate/Trinotate/wiki/Software-installation-and-data-required

```{bash, eval=FALSE}
PATH_DIR="/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinotate_output"

# Variables (I moved the folder TRINOTATE_DATA_DIR locally for the run to avoid path mess with apptainer)
# Was originally in /usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1

export TRINOTATE_DATA_DIR=/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinotate_output/TRINOTATE_DATA_DIR

export TRINOTATE_HOME=/usr/local/src/Trinotate

PATH_DIR="/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinotate_output"

while read NAME_SPECIES; do
   mkdir $NAME_SPECIES
   
   apptainer exec --bind $PATH_DIR -e ~/bin/trinotate.v4.0.2.simg $TRINOTATE_HOME/Trinotate \
      --db $NAME_SPECIES/database_${NAME_SPECIES}.sqlite \
      --create \
      --trinotate_data_dir $TRINOTATE_DATA_DIR \
      --use_diamond
done < list_species.txt

    
```


## Trinotate


```{bash, eval=FALSE}
#!/bin/bash
#SBATCH --job-name=trinotate
#SBATCH --nodes=1
#SBATCH --partition=medium
#SBATCH --time=16:00:00
#SBATCH --cpus-per-task=10
#SBATCH --mem=40gb
#SBATCH --mail-type=END
#SBATCH --mail-user=johan.zicola@uni-goettingen.de
#SBATCH --output=trinotate.%J.out
#SBATCH --error=trinotate.%J.err

mem_gbytes=$(( $SLURM_MEM_PER_NODE/1024 ))

echo "Starting at $(date)"
echo "Job submitted to the ${SLURM_JOB_PARTITION} partition, the default partition on ${SLURM_CLUSTER_NAME}"
echo "Job name: ${SLURM_JOB_NAME}, Job ID: ${SLURM_JOB_ID}"
echo "I have ${SLURM_CPUS_ON_NODE} CPUs and ${mem_gbytes}GiB of RAM on compute node $(hostname)"

# Modules to load
module purge
module load apptainer

# Test that all programs needed are installed
test_program(){
  command -v $1 >/dev/null || \
    { echo >&2 " $1 required but not installed. Aborting"; exit 1; }
}

test_program "tmhmm"
test_program "signalp6"

# Variables passed to sbatch
# NAME_SPECIES
# Output from Trinity $GENE_TRANS_MAP and $TRINITY_FASTA
# Output from Transdecoder $TRANSDECODER_PEP

PATH_DIR="/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects"

# Variables (I moved the folder TRINOTATE_DATA_DIR locally for the run to avoid path mess with apptainer)
# Was originally in /usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1
export TRINOTATE_DATA_DIR=${PATH_DIR}/trinotate_output/TRINOTATE_DATA_DIR

export TRINOTATE_HOME=/usr/local/src/Trinotate

#NAME_SPECIES is provided as argument to sbatch (Do not put a uppercase to trinotate_ suffix, it won't work)
sqlite_db="${PATH_DIR}/trinotate_output/${NAME_SPECIES}/database_${NAME_SPECIES}.sqlite"


echo "###############################"
echo "Trinotate Init"
date
echo "###############################"

apptainer exec --bind $PATH_DIR -e ~/bin/trinotate.v4.0.2.simg $TRINOTATE_HOME/Trinotate \
   --db $sqlite_db --init \
   --gene_trans_map $GENE_TRANS_MAP \
   --transcript_fasta $TRINITY_FASTA \
   --transdecoder_pep $TRANSDECODER_PEP

echo "##############################"
echo  "Trinotate Run"
date
echo "##############################"

apptainer exec --bind $PATH_DIR -e ~/bin/trinotate.v4.0.2.simg $TRINOTATE_HOME/Trinotate \
   --CPU ${SLURM_CPUS_ON_NODE} \
   --db $sqlite_db \
   --run ALL \
   --trinotate_data_dir $TRINOTATE_DATA_DIR \
   --transcript_fasta $TRINITY_FASTA \
   --transdecoder_pep $TRANSDECODER_PEP \
   --use_diamond

echo "##############################"
echo  "Trinotate Run signalp"
date
echo "##############################"

signalp6 --fastafile $TRANSDECODER_PEP \
   --output_dir ${PATH_DIR}/trinotate_output/${NAME_SPECIES}/sigP6outdir \
   --format none --organism euk --mode fast


apptainer exec --bind $PATH_DIR -e ~/bin/trinotate.v4.0.2.simg $TRINOTATE_HOME/Trinotate \
   --db $sqlite_db \
   --LOAD_signalp ${PATH_DIR}/trinotate_output/${NAME_SPECIES}/sigP6outdir/output.gff3


echo "##############################"
echo  "Trinotate Run TMHMM"
date
echo "##############################"

tmhmm --short $TRANSDECODER_PEP > ${PATH_DIR}/trinotate_output/${NAME_SPECIES}/tmhmm.v2.out

apptainer exec --bind $PATH_DIR -e ~/bin/trinotate.v4.0.2.simg $TRINOTATE_HOME/Trinotate \
   --db $sqlite_db \
   --LOAD_tmhmmv2 ${PATH_DIR}/trinotate_output/${NAME_SPECIES}/tmhmm.v2.out

echo "###########################"
echo "Generating report table"
date
echo "###########################"

apptainer exec --bind $PATH_DIR -e ~/bin/trinotate.v4.0.2.simg $TRINOTATE_HOME/Trinotate \
   --db ${sqlite_db} \
   --report \
   --incl_pep \
   --incl_trans > ${PATH_DIR}/trinotate_output/${NAME_SPECIES}/Trinotate_report.tsv


##################
## Misc value adds
date
##################

# Extract GO terms

apptainer exec --bind $PATH_DIR -e ~/bin/trinotate.v4.0.2.simg \
   $TRINOTATE_HOME/util/extract_GO_assignments_from_Trinotate_xls.pl \
  --Trinotate_xls Trinotate_report.tsv \
  -G -I > ${PATH_DIR}/trinotate_output/${NAME_SPECIES}/Trinotate_report.tsv.gene_ontology

# Generate trinotate report summary statistics
apptainer exec --bind $PATH_DIR -e ~/bin/trinotate.v4.0.2.simg \
   $TRINOTATE_HOME/util/report_summary/trinotate_report_summary.pl \
  ${PATH_DIR}/trinotate_output/${NAME_SPECIES}/Trinotate_report.tsv \
  ${PATH_DIR}/trinotate_output/${NAME_SPECIES}/Trinotate_report_stats


echo "##########################"
echo "done.  See annotation summary file:  Trinotate_report.tsv"
date
echo "##########################"

```

Run one by one because I notice intermediate files and report are written in $HOME without running number, which will lead to potential conflicts for parallel runs.

```{bash, eval=FALSE}

work_path="/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects"

i="Amara_similata"
cd $i
sbatch --export=NAME_SPECIES=$i,GENE_TRANS_MAP=${work_path}/trinity_output/summary/${i}_trinity.fasta.gene_trans_map,TRINITY_FASTA=${work_path}/trinity_output/summary/${i}_trinity.fasta,TRANSDECODER_PEP=${work_path}/transdecoder_output/${i}_trinity.fasta.transdecoder.pep ../04_trinotate.sh

```



Job killed due to error

```{bash, eval=FALSE}
CMD: /usr/local/src/Trinotate/util/report_summary/../gene_ontology/Trinotate_GO_to_SLIM.pl /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinotate_output/Amara_similata/Trinotate_report_stats.GO > /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinotate_output/Amara_similata/Trinotate_report_stats.GO.slim
CMD: /usr/local/src/Trinotate/util/report_summary/util/make_cXp_html.pl /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinotate_output/Amara_similata/Trinotate_report_stats > /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinotate_output/Amara_similata/Trinotate_report_stats.cXp_summary.html
Can't use an undefined value as an ARRAY reference at /usr/local/src/Trinotate/util/report_summary/util/../../../TrinotateWeb/cgi-bin/PerlLib/CanvasXpress/Sunburst.pm line 77.
Error, cmd:
/usr/local/src/Trinotate/util/report_summary/util/make_cXp_html.pl /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinotate_output/Amara_similata/Trinotate_report_stats > /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinotate_output/Amara_similata/Trinotate_report_stats.cXp_summary.html
 died with ret (65280) at /usr/local/src/Trinotate/util/report_summary/../../PerlLib/Process_cmd.pm line 19, <$fh> line 1.
        Process_cmd::process_cmd("/usr/local/src/Trinotate/util/report_summary/util/make_cXp_ht"...) called at /usr/local/src/Trinotate/util/report_summary/trinotate_report_summary.pl line 102

```


```{bash, eval=FALSE}

apptainer exec --bind $PATH_DIR -e ~/bin/trinotate.v4.0.2.simg $TRINOTATE_HOME/Trinotate \
   --db ${sqlite_db} \
   --report \
   --incl_pep \
   --incl_trans > ${PATH_DIR}/trinotate_output/${NAME_SPECIES}/Trinotate_report.tsv

```



```{bash, eval=FALSE}
#!/bin/bash
#SBATCH --job-name=test_pwd
#SBATCH --nodes=1
#SBATCH --partition=medium
#SBATCH --time=00:00:10
#SBATCH --cpus-per-task=1
#SBATCH --mem=40mb
#SBATCH --output=trinotate.%J.out
#SBATCH --chdir="/usr/users/zicola"

pwd


```



## Problem with DB download


## Remake full database

I think I screwed up something in TRINOTATE_DATA_DIR so download everything anew.

```{bash, eval=FALSE}
module load apptainer

export TRINOTATE_DATA_DIR=/scratch/users/zicola/TRINOTATE_DATA_DIR
export TRINOTATE_HOME=/usr/local/src/Trinotate

PATH_DIR="/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinotate_output"

apptainer exec --bind /scratch/users/zicola -e ~/bin/trinotate.v4.0.2.simg $TRINOTATE_HOME/Trinotate \
      --db test.sqlite \
      --create \
      --trinotate_data_dir $TRINOTATE_DATA_DIR \
      --use_diamond

cd /scratch/users/zicola

mkdir TRINOTATE_DATA_DIR

srun --time=06:00:00 -p medium --mem=20Gb bash create_database.sh

```


Rfam files did not download sucessfully so I downloaded them manually from https://ftp.ebi.ac.uk/pub/databases/Rfam/CURRENT/

Kill the job at this point, then do the steps my self and relaunch

```{bash, eval=FALSE}
cd /scratch/users/zicola/trinotate

wget ftp://ftp.ebi.ac.uk/pub/databases/Rfam/CURRENT/Rfam.cm.gz

gunzip Rfam.cm.gz
cmpress Rfam.cm

Working...    done.
Pressed and indexed 4178 CMs and p7 HMM filters (4178 names and 4178 accessions).
Covariance models and p7 filters pressed into binary file:  Rfam.cm.i1m
SSI index for binary covariance model file:                 Rfam.cm.i1i
Optimized p7 filter profiles (MSV part)  pressed into:      Rfam.cm.i1f
Optimized p7 filter profiles (remainder) pressed into:      Rfam.cm.i1p


wget https://ftp.ebi.ac.uk/pub/databases/Rfam/CURRENT/Rfam.clanin

mv Rfam.* /scratch/users/zicola/trinotate/TRINOTATE_DATA_DIR

# Create a fake checkpoint so that it does not try to download the files again
touch TRINOTATE_DATA_DIR/__chckpts/cmpress_Rfam_cm.ok

srun --time=06:00:00 -o output2.txt -p medium --mem=20Gb bash create_database.sh

```


I also moved the EGGNOG files as they are heavy and I have alread downloaded them recently

```{bash, eval=FALSE}
cd /scratch/users/zicola/TRINOTATE_DATA_DIR
mv /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinotate_output/TRINOTATE_DATA_DIR_OLD/EGGNOG_DATA_DIR .
```


```{bash, eval=FALSE}
module load apptainer

export TRINOTATE_DATA_DIR=/scratch/users/zicola/TRINOTATE_DATA_DIR
export TRINOTATE_HOME=/usr/local/src/Trinotate

PATH_DIR="/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinotate_output"

apptainer exec --bind /scratch/users/zicola -e ~/bin/trinotate.v4.0.2.simg $TRINOTATE_HOME/Trinotate \
      --db test.sqlite \
      --create \
      --trinotate_data_dir $TRINOTATE_DATA_DIR \
      --use_diamond
```

```

-CREATING test.sqlite and populating data dir: /scratch/users/zicola/TRINOTATE_DATA_DIR
WARNING: SQLITE database test.sqlite already exists and wont be replaced.
-sqlite db test.sqlite already exists and is not being replaced. If a new boilerplate is required, copy db from: /scratch/users/zicola/TRINOTATE_DATA_DIR/TrinotateBoilerplate.sqlite
-- Skipping CMD: /usr/local/src/Trinotate/util/admin/Build_Trinotate_Boilerplate_SQLite_db.pl TrinotateBoilerplate, checkpoint [/scratch/users/zicola/TRINOTATE_DATA_DIR/__chckpts/build_boilerplate.ok] exists.
-- Skipping CMD: mv TrinotateBoilerplate.sqlite pfam2go go-basic.obo pfam2go.tab NOG.annotations.tsv.gz /scratch/users/zicola/TRINOTATE_DATA_DIR/, checkpoint [/scratch/users/zicola/TRINOTATE_DATA_DIR/__chckpts/copy_boilerplate_to_datadir.ok] exists.
-- Skipping CMD: makeblastdb -in uniprot_sprot.pep -dbtype prot, checkpoint [/scratch/users/zicola/TRINOTATE_DATA_DIR/__chckpts/make_sprot_blast_db.ok] exists.
-- Skipping CMD: diamond makedb --in uniprot_sprot.pep -d uniprot_sprot, checkpoint [/scratch/users/zicola/TRINOTATE_DATA_DIR/__chckpts/make_diamond_sprot_db.ok] exists.
-- Skipping CMD: mv uniprot_sprot* /scratch/users/zicola/TRINOTATE_DATA_DIR/, checkpoint [/scratch/users/zicola/TRINOTATE_DATA_DIR/__chckpts/relocate_sprot_db.ok] exists.
-- Skipping CMD: gunzip Pfam-A.hmm.gz, checkpoint [/scratch/users/zicola/TRINOTATE_DATA_DIR/__chckpts/gunzip_pfam.ok] exists.
-- Skipping CMD: hmmpress Pfam-A.hmm, checkpoint [/scratch/users/zicola/TRINOTATE_DATA_DIR/__chckpts/hmmpress_pfam_db.ok] exists.
-- Skipping CMD: mv Pfam-A.hmm* /scratch/users/zicola/TRINOTATE_DATA_DIR/, checkpoint [/scratch/users/zicola/TRINOTATE_DATA_DIR/__chckpts/relocate_pfam_db.ok] exists.
-- Skipping CMD: gunzip Pfam-A.hmm.gz, checkpoint [/scratch/users/zicola/TRINOTATE_DATA_DIR/__chckpts/gunzip_pfam.ok] exists.
-- Skipping CMD: hmmpress Pfam-A.hmm, checkpoint [/scratch/users/zicola/TRINOTATE_DATA_DIR/__chckpts/hmmpress_pfam_db.ok] exists.
-- Skipping CMD: mv Pfam-A.hmm* /scratch/users/zicola/TRINOTATE_DATA_DIR/, checkpoint [/scratch/users/zicola/TRINOTATE_DATA_DIR/__chckpts/relocate_pfam_db.ok] exists.
-- Skipping CMD: wget ftp://ftp.ebi.ac.uk/pub/databases/Rfam/CURRENT/Rfam.cm.gz, checkpoint [/scratch/users/zicola/TRINOTATE_DATA_DIR/__chckpts/wget_Rfam.cm.gz] exists.
-- Skipping CMD: gunzip Rfam.cm.gz, checkpoint [/scratch/users/zicola/TRINOTATE_DATA_DIR/__chckpts/gunzip_Rfam.cm.gz] exists.
-- Skipping CMD: wget ftp://ftp.ebi.ac.uk/pub/databases/Rfam/CURRENT/Rfam.clanin, checkpoint [/scratch/users/zicola/TRINOTATE_DATA_DIR/__chckpts/wget_clanin.cm.gz] exists.
-- Skipping CMD: cmpress Rfam.cm, checkpoint [/scratch/users/zicola/TRINOTATE_DATA_DIR/__chckpts/cmpress_Rfam_cm.ok] exists.
-- Skipping CMD: mv Rfam.* /scratch/users/zicola/TRINOTATE_DATA_DIR/, checkpoint [/scratch/users/zicola/TRINOTATE_DATA_DIR/__chckpts/relocate_Rfam_db.ok] exists.
-EGGNOG_MAPPER: installing eggnogmapper data at: /scratch/users/zicola/TRINOTATE_DATA_DIR/EGGNOG_DATA_DIR
* [Thu Jan 23 09:49:49 2025] Running CMD: download_eggnog_data.py --data_dir /scratch/users/zicola/TRINOTATE_DATA_DIR/EGGNOG_DATA_DIR -y
Skipping eggnog.db database (already present). Use -f to force download
Skipping eggnog.taxa.db database (already present). Use -f to force download
Skipping diamond database (or already present). Use -f to force download
Skipping novel families diamond database (or already present). Use -F and -f to force download
Skipping Pfam database (or already present). Use -P and -f to force download
Skipping MMseqs2 database (or already present). Use -M and -f to force download
No HMMER database requested. Use "-H -d taxid" to download the hmmer database for taxid
Finished.
```

It seems every it worked but I don't think the boiler plate included all files because of this checkpoint files.


```{bash, eval=FALSE}

module load apptainer

export TRINOTATE_DATA_DIR=/scratch/users/zicola/trinotate/TRINOTATE_DATA_DIR
export TRINOTATE_HOME=/usr/local/src/Trinotate

apptainer exec --bind /scratch/users/zicola -e ~/bin/trinotate.v4.0.2.simg $TRINOTATE_HOME/Trinotate \
      --db test.sqlite \
      --create \
      --trinotate_data_dir $TRINOTATE_DATA_DIR \
      --use_diamond


cd /scratch/users/zicola/trinotate

srun --time=06:00:00 -p medium --mem=10Gb bash create_database.sh

```


## Amara_similata

```{bash, eval=FALSE}
i="Amara_similata"
cd $i
sbatch --export=NAME_SPECIES=$i,GENE_TRANS_MAP=${work_path}/trinity_output/summary/${i}_trinity.fasta.gene_trans_map,TRINITY_FASTA=${work_path}/trinity_output/summary/${i}_trinity.fasta,TRANSDECODER_PEP=${work_path}/transdecoder_output/${i}_trinity.fasta.transdecoder.pep ../04_trinotate.sh
```


What happens in the background (using err file from slurm)

```{bash, eval=FALSE}

# Load transcripts into sqlite database
/usr/local/src/Trinotate/util/trinotateSeqLoader/TrinotateSeqLoader.pl \
   --sqlite /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinotate_output/Amara_si
milata/database_Amara_similata.sqlite \
   --gene_trans_map /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinity_output/summary/Amara_similata_trinity.fasta.gen
e_trans_map \
   --transcript_fasta /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinity_output/summary/Amara_similata_trinity.fasta \
      --transdecoder_pep /usr/use
rs/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/transdecoder_output/Amara_similata_trinity.fasta.transdecoder.pep \
   --bulk_load

# Diamong blastp
diamond blastp -d /scratch/users/zicola/TRINOTATE_DATA_DIR/uniprot_sprot \
   -q /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/transdecoder_output/Amara_similata_trinity.fasta.transdecoder.pep \
   -p 8 -k 1 -e 1e-5 -o \
   uniprot_sprot.diamond.blastp.outfmt6 --outfmt 6

# Load blastp
/usr/local/src/Trinotate/util/trinotateSeqLoader/Trinotate_BLAST_loader.pl --sqlite /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinotate_output/Amara_similata/database_Amara_similata.sqlite --outfmt6 uniprot_sprot.diamond.blastp.outfmt6 \
   --prog blastp \
   --dbtype Swissprot

# Diamong blastx
diamond blastx -d /scratch/users/zicola/TRINOTATE_DATA_DIR/uniprot_sprot -q /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinity_output/summary/Amara_similata_trinity.fasta \
   -p 8 -k 1 -e 1e-5 \
   -o uniprot_sprot.diamond.blastx.outfmt6 --outfmt 6

# Load blastx
/usr/local/src/Trinotate/util/trinotateSeqLoader/Trinotate_BLAST_loader.pl \
--sqlite /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinotate_output/Amara_similata/database_Amara_similata.sqlite --outfmt6 uniprot_sprot.diamond.blastx.outfmt6 \
   --prog blastx \
   --dbtype Swissprot

# hmmsearch (PFAM)
hmmsearch --cpu 8 --noali --domtblout TrinotatePFAM.out /scratch/users/zicola/TRINOTATE_DATA_DIR/Pfam-A.hmm /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/transdecoder_output/Amara_similata_trinity.fasta.transdecoder.pep > pfam.log

# PFAM
/usr/local/src/Trinotate/Trinotate --db /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinotate_output/Amara_similata/database_Amara_similata.sqlite --LOAD_pfam  TrinotatePFAM.out

# Load PFAM result
/usr/local/src/Trinotate/util/trinotateSeqLoader/Trinotate_PFAM_loader.pl --sqlite /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinotate_output/Amara_similata/database_Amara_similata.sqlite --pfam TrinotatePFAM.out

# Infernal
cmscan -Z 90 --cut_ga --rfam --nohmmonly --tblout infernal.out --fmt 2 --cpu 8 --clanin /scratch/users/zicola/TRINOTATE_DATA_DIR/Rfam.clanin /scratch/users/zicola/TRINOTATE_DATA_DIR/Rfam.cm /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinity_output/summary/Amara_similata_trinity.fasta > infernal.log

# Load
/usr/local/src/Trinotate/Trinotate --db /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinotate_output/Amara_similata/database_Amara_similata.sqlite --LOAD_infernal infernal.out

# emapper
emapper.py -i /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/transdecoder_output/Amara_similata_trinity.fasta.transdecoder.pep --cpu 8 --data_dir /scratch/users/zicola/TRINOTATE_DATA_DIR/EGGNOG_DATA_DIR -o eggnog_mapper

# Load
/usr/local/src/Trinotate/util/trinotateSeqLoader/Trinotate_EggnogMapper_loader.pl --sqlite /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinotate_output/Amara_similata/database_Amara_similata.sqlite --emapper eggnog_mapper.emapper.annotations

```




Still creating files in $HOME

```{bash, eval=FALSE}
zicola@gwdu101:~
$ ls -h | grep "Jan 23"
drwx------  5 zicola UANP         2 Jan 23  2024 R
-rw-r-----  1 zicola UANP      120M Jan 23 13:26 TrinotatePFAM.out
-rw-r-----  1 zicola UANP         0 Jan 23 12:50 __init.ok
drwxr-x---  2 zicola UANP         2 Jan 23 22:00 __trinotate_run_checkpts
-rw-r-----  1 zicola UANP       43M Jan 23 22:00 eggnog_mapper.emapper.annotations
-rw-r-----  1 zicola UANP      9.1M Jan 23 19:54 eggnog_mapper.emapper.hits
-rw-r-----  1 zicola UANP      2.3M Jan 23 22:00 eggnog_mapper.emapper.seed_orthologs
-rw-r-----  1 zicola UANP      129M Jan 23 19:41 infernal.log
-rw-r-----  1 zicola UANP       79K Jan 23 19:41 infernal.out
-rw-r-----  1 zicola UANP      134M Jan 23 13:26 pfam.log
-rw-r-----  1 zicola UANP      1.6M Jan 23 12:50 uniprot_sprot.diamond.blastp.outfmt6
-rw-r-----  1 zicola UANP      1.6M Jan 23 12:50 uniprot_sprot.diamond.blastx.outfmt6

```

It seems the flag `--workdir ${PATH_DIR}/trinotate_output/${NAME_SPECIES}` does not help to make Trinotate generate its output in the working directory


Check how many transcripts were annotated as rRNA:

```{bash, eval=FALSE}
grep "rRNA" infernal.out | wc -l
111


```




```{bash, eval=FALSE}

#!/bin/bash
#SBATCH --job-name=trinotate
#SBATCH --nodes=1
#SBATCH --partition=medium
#SBATCH --time=16:00:00
#SBATCH --cpus-per-task=1
#SBATCH --mem=20gb
#SBATCH --mail-type=END
#SBATCH --mail-user=johan.zicola@uni-goettingen.de
#SBATCH --output=trinotate.%J.out
#SBATCH --error=trinotate.%J.err

mem_gbytes=$(( $SLURM_MEM_PER_NODE/1024 ))

echo "Starting at $(date)"
echo "Job submitted to the ${SLURM_JOB_PARTITION} partition, the default partition on ${SLURM_CLUSTER_NAME}"
echo "Job name: ${SLURM_JOB_NAME}, Job ID: ${SLURM_JOB_ID}"
echo "I have ${SLURM_CPUS_ON_NODE} CPUs and ${mem_gbytes}GiB of RAM on compute node $(hostname)"

# Modules to load
module purge
module load apptainer

# Test that all programs needed are installed
test_program(){
  command -v $1 >/dev/null || \
    { echo >&2 " $1 required but not installed. Aborting"; exit 1; }
}

test_program "tmhmm"
test_program "signalp6"

# Variables passed to sbatch
# NAME_SPECIES
# Output from Trinity $GENE_TRANS_MAP and $TRINITY_FASTA
# Output from Transdecoder $TRANSDECODER_PEP

PATH_DIR="/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects"

# Variables (I moved the folder TRINOTATE_DATA_DIR locally for the run to avoid path mess with apptainer)
# Was originally in /usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1
export TRINOTATE_DATA_DIR=/scratch/users/zicola/TRINOTATE_DATA_DIR

export TRINOTATE_HOME=/usr/local/src/Trinotate

#NAME_SPECIES is provided as argument to sbatch (Do not put a uppercase to trinotate_ suffix, it won't work)
sqlite_db="${PATH_DIR}/trinotate_output/${NAME_SPECIES}/database_${NAME_SPECIES}.sqlite"


echo "##############################"
echo  "Trinotate Run signalp"
date
echo "##############################"

signalp6 --fastafile $TRANSDECODER_PEP \
   --output_dir ${PATH_DIR}/trinotate_output/${NAME_SPECIES}/sigP6outdir \
   --format none --organism euk --mode fast


apptainer exec --bind $PATH_DIR,$TRINOTATE_DATA_DIR -e ~/bin/trinotate.v4.0.2.simg $TRINOTATE_HOME/Trinotate \
   --db $sqlite_db \
   --LOAD_signalp ${PATH_DIR}/trinotate_output/${NAME_SPECIES}/sigP6outdir/output.gff3


echo "##############################"
echo  "Trinotate Run TMHMM"
date
echo "##############################"

tmhmm --short $TRANSDECODER_PEP > ${PATH_DIR}/trinotate_output/${NAME_SPECIES}/tmhmm.v2.out

apptainer exec --bind $PATH_DIR,$TRINOTATE_DATA_DIR -e ~/bin/trinotate.v4.0.2.simg $TRINOTATE_HOME/Trinotate \
   --db $sqlite_db \
   --LOAD_tmhmmv2 ${PATH_DIR}/trinotate_output/${NAME_SPECIES}/tmhmm.v2.out

echo "###########################"
echo "Generating report table"
date
echo "###########################"

apptainer exec --bind $PATH_DIR,$TRINOTATE_DATA_DIR -e ~/bin/trinotate.v4.0.2.simg $TRINOTATE_HOME/Trinotate \
   --db ${sqlite_db} \
   --report \
   --incl_pep \
   --incl_trans > ${PATH_DIR}/trinotate_output/${NAME_SPECIES}/Trinotate_report.tsv


##################
## Misc value adds
date
##################

# Extract GO terms

apptainer exec --bind $PATH_DIR,$TRINOTATE_DATA_DIR -e ~/bin/trinotate.v4.0.2.simg \
   $TRINOTATE_HOME/util/extract_GO_assignments_from_Trinotate_xls.pl \
  --Trinotate_xls Trinotate_report.tsv \
  -G -I > ${PATH_DIR}/trinotate_output/${NAME_SPECIES}/Trinotate_report.tsv.gene_ontology

# Generate trinotate report summary statistics
apptainer exec --bind $PATH_DIR,$TRINOTATE_DATA_DIR -e ~/bin/trinotate.v4.0.2.simg \
   $TRINOTATE_HOME/util/report_summary/trinotate_report_summary.pl \
  ${PATH_DIR}/trinotate_output/${NAME_SPECIES}/Trinotate_report.tsv \
  ${PATH_DIR}/trinotate_output/${NAME_SPECIES}/Trinotate_report_stats


echo "##########################"
echo "done.  See annotation summary file:  Trinotate_report.tsv"
date
echo "##########################"
```



# tmhmm alone


I just need to do tmhmm as I've already done signalp for Amara similara


```{bash, eval=FALSE}

#!/bin/bash
#SBATCH --job-name=trinotate_tmhmm
#SBATCH --nodes=1
#SBATCH --partition=medium
#SBATCH --time=06:00:00
#SBATCH --cpus-per-task=1
#SBATCH --mem=5G
#SBATCH --mail-type=END
#SBATCH --mail-user=johan.zicola@uni-goettingen.de
#SBATCH --output=trinotate_tmhmm.%J.out
#SBATCH --error=trinotate_tmhmm..%J.err

mem_gbytes=$(( $SLURM_MEM_PER_NODE/1024 ))

echo "Starting at $(date)"
echo "Job submitted to the ${SLURM_JOB_PARTITION} partition, the default partition on ${SLURM_CLUSTER_NAME}"
echo "Job name: ${SLURM_JOB_NAME}, Job ID: ${SLURM_JOB_ID}"
echo "I have ${SLURM_CPUS_ON_NODE} CPUs and ${mem_gbytes}GiB of RAM on compute node $(hostname)"

# Modules to load
module purge
module load apptainer

# Test that all programs needed are installed
test_program(){
  command -v $1 >/dev/null || \
    { echo >&2 " $1 required but not installed. Aborting"; exit 1; }
}

test_program "tmhmm"

# Path to trinotate executable
export TRINOTATE_HOME=/usr/local/src/Trinotate

# Path to Trinity and Transdecoder files
export WORKPATH=/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects
export GENE_TRANS_MAP=${WORKPATH}/trinity_output/summary/${SPECIES}_trinity.fasta.gene_trans_map
export TRINITY_FASTA=${WORKPATH}/trinity_output/summary/${SPECIES}_trinity.fasta
export TRANSDECODER_PEP=${WORKPATH}/transdecoder_output/${SPECIES}_trinity.fasta.transdecoder.pep 

# Database
export TRINOTATE_DATA_DIR=/scratch/users/zicola/TRINOTATE_DATA_DIR
export EGGNOG_DATA_DIR=/scratch/users/zicola/TRINOTATE_DATA_DIR/EGGNOG_DATA_DIR
export SQLITE_DB=${WORKPATH}/trinotate_output/${SPECIES}/database_${SPECIES}.sqlite

cd $WORKPATH

tmhmm --short $TRANSDECODER_PEP > ${WORKPATH}/trinotate_output/${SPECIES}/tmhmm.v2.out

apptainer exec --bind $WORKPATH -e ~/bin/trinotate.v4.0.2.simg $TRINOTATE_HOME/Trinotate \
   --db $SQLITE_DB \
   --LOAD_tmhmmv2 ${WORKPATH}/trinotate_output/${SPECIES}/tmhmm.v2.out
   

```

```{bash, eval=FALSE}

sbatch --export=SPECIES="Amara_similata" 05_tmhmm.sh

```

```{bash, eval=FALSE}

sacct -j 5431422   -o jobid,jobname,AllocCPUS,ReqMem,MaxRSS,Elapsed

sacct -j 5431422   -o jobid,jobname,AllocCPUS,ReqMem,MaxRSS,Elapsed
JobID           JobName  AllocCPUS     ReqMem     MaxRSS    Elapsed
------------ ---------- ---------- ---------- ---------- ----------
5431422      trinotate+          1         5G              00:27:02
5431422.bat+      batch          1                48284K   00:27:02
5431422.ext+     extern          1                     0   00:27:02

```

Took only 48 Mb of RAM and 27 min
