---
title: "other_scripts"
author: "Johan Zicola"
date: '2025-01-16'
output: html_document
---


# Remove empty reads

```{bash, eval=FALSE}

while read i; do
  srun -p medium -t 01:20:00 --cpus-per-task=10 cutadapt -j 10 --minimum-length 25 \
  -o ${i}_L1_1.trimmed.min25.fq -p ${i}_L1_2.trimmed.min25.fq \
  ${i}_L1_1.trimmed.fq ${i}_L1_2.trimmed.fq &
done < list.txt

```

# Trimming with trimmomatic

```{bash, eval=FALSE}

module load rev/21.12 trimmomatic openjdk

PATH_ADAPTER="/usr/product/bioinfo/SL_7.0/BIOINFORMATICS/TRIMMOMATIC/0.36/adapters/"

srun -p medium -t 00:30:00 --mem=12G --cpus-per-task=10 \
java -Xmx10G -jar /usr/product/bioinfo/SL_7.0/BIOINFORMATICS/TRIMMOMATIC/0.36/trimmomatic-0.36.jar \
PE -phred33 -threads 10 \
raw_data/01_Ad/01_Ad_L1_1.fq.gz \
raw_data/01_Ad/01_Ad_L1_2.fq.gz \
trimmed_reads_trimmomatic/01_Ad_L1_1.paired.fq.gz \
trimmed_reads_trimmomatic/01_Ad_L1_1.unpaired.fq.gz \
trimmed_reads_trimmomatic/01_Ad_L1_2.paired.fq.gz \
trimmed_reads_trimmomatic/01_Ad_L1_2.unpaired.fq.gz \
ILLUMINACLIP:"/usr/product/bioinfo/SL_7.0/BIOINFORMATICS/TRIMMOMATIC/0.36/adapters/TruSeq2-SE.fa":2:30:10 LEADING:3 TRAILING:3 SLIDINGWINDOW:4:15 MINLEN:50 &

Input Read Pairs: 64769548 Both Surviving: 63780516 (98.47%) Forward Only Surviving: 228552 (0.35%) Reverse Only Surviving: 744814 (1.15%) Dropped: 15666 (0.02%)
TrimmomaticPE: Completed successfully

seqkit stats 01_Ad_L1_1.paired.fq.gz 01_Ad_L1_2.paired.fq.gz
processed files:  2 / 2 [======================================] ETA: 0s. done
file                     format  type    num_seqs        sum_len  min_len  avg_len  max_len
01_Ad_L1_1.paired.fq.gz  FASTQ   DNA   63,780,516  6,351,162,309       50     99.6      100
01_Ad_L1_2.paired.fq.gz  FASTQ   DNA   63,780,516  6,350,850,563       50     99.6      100

# Same number of reads in each file

module load fastqc/0.11.4

srun -p medium -t 00:50:00 fastqc 01_Ad_L1_2.paired.fq.gz &
srun -p medium -t 00:50:00 fastqc 01_Ad_L1_1.paired.fq.gz &

```


# Remove rRNA

https://github.com/sortmerna/sortmerna

Download v4.3 database (https://github.com/biocore/sortmerna/releases/download/v4.3.4/database.tar.gz)

## Index database

The docu recommends to use `smr_v4.3_default_db.fasta`

```{r}
mkdir ~/genomes/silva_rRNA_maize/smr_database && cd ~/genomes/silva_rRNA_maize/smr_database

wget https://github.com/biocore/sortmerna/releases/download/v4.3.4/database.tar.gz

tar -zxvf database.tar.gz

seqkit stats smr_v4.3_default_db.fasta
file                       format  type  num_seqs      sum_len  min_len  avg_len  max_len
smr_v4.3_default_db.fasta  FASTA   RNA     73,977  133,533,894       91  1,805.1    4,380

module load rev/21.12 sortmerna/20170809

srun -t 00:50:00 --mem=20G -p medium indexdb_rna --ref smr_v4.3_default_db.fasta,smr_v4.3_default_db.idx -v

```


```{bash, eval=FALSE}
\ls -1sh smr_v4.3_default*
140M smr_v4.3_default_db.fasta
662M smr_v4.3_default_db.idx.bursttrie_0.dat
1.0M smr_v4.3_default_db.idx.kmer_0.dat
1.2G smr_v4.3_default_db.idx.pos_0.dat
4.0M smr_v4.3_default_db.idx.stats
```

4 index files are generated with a *dat extension.

## Merge reads

I tested with the R1 and R2 in separate files using two --read flags but the output contained only R2 sequences. When I merge the R1 and R2, it works. I just need to merge R1 and R2 and unmerge them afterwards

```{bash, eval=FALSE}

cd /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trimmed_reads

mkdir merged_fastq

module load rev/21.12 sortmerna/20170809

while read i; do
if [ -e merged_fastq/${i}.merged.fq ]; then
  echo "$i already processed"
else
  srun -p medium --mem=2Gb --time=00:20:00 bash /usr/product/bioinfo/SL_7.0/BIOINFORMATICS/SORTMERNA/sortmerna_20170809/scripts/merge-paired-reads.sh ${i}_L1_1.trimmed.fq ${i}_L1_2.trimmed.fq  merged_fastq/${i}.merged.fq &
fi
done < list.txt


```


```{bash, eval=FALSE}
#!/bin/bash
#SBATCH --job-name=merge_fastq
#SBATCH --nodes=1
#SBATCH --time=00:40:00
#SBATCH --cpus-per-task=1
#SBATCH --mem=10gb
#SBATCH --partition=medium
#SBATCH --mail-type=END
#SBATCH --mail-user=johan.zicola@uni-goettingen.de
#SBATCH --output=sortmerna.%J.out
#SBATCH --error=sortmerna.%J.err
#SBATCH --array=1-8

mem_gbytes=$(( $SLURM_MEM_PER_NODE/1024 ))

echo "Starting at $(date)"
echo "Job submitted to the ${SLURM_JOB_PARTITION} partition, the default partition on ${SLURM_CLUSTER_NAME}"
echo "Job name: ${SLURM_JOB_NAME}, Job ID: ${SLURM_JOB_ID}"
echo "I have ${SLURM_CPUS_ON_NODE} CPUs and ${mem_gbytes}GiB of RAM on compute node $(hostname)"

module purge

module load rev/21.12 sortmerna/20170809

i=$(sed -n ${SLURM_ARRAY_TASK_ID}p trimmed_reads/list.txt)

bash /usr/product/bioinfo/SL_7.0/BIOINFORMATICS/SORTMERNA/sortmerna_20170809/scripts/merge-paired-reads.sh trimmed_reads/${i}_L1_1.trimmed.fq trimmed_reads/${i}_L1_2.trimmed.fq  trimmed_reads/merged_fastq/${i}.merged.fq

```


## sortmerna

```{r}
#!/bin/bash
#SBATCH --job-name=sortmerna
#SBATCH --nodes=1
#SBATCH --time=01:00:00
#SBATCH --cpus-per-task=10
#SBATCH --mem=50gb
#SBATCH --partition=medium
#SBATCH --mail-type=END
#SBATCH --mail-user=johan.zicola@uni-goettingen.de
#SBATCH --output=sortmerna.%J.out
#SBATCH --error=sortmerna.%J.err
#SBATCH --array=1-8

mem_gbytes=$(( $SLURM_MEM_PER_NODE/1024 ))

echo "Starting at $(date)"
echo "Job submitted to the ${SLURM_JOB_PARTITION} partition, the default partition on ${SLURM_CLUSTER_NAME}"
echo "Job name: ${SLURM_JOB_NAME}, Job ID: ${SLURM_JOB_ID}"
echo "I have ${SLURM_CPUS_ON_NODE} CPUs and ${mem_gbytes}GiB of RAM on compute node $(hostname)"

module purge

module load rev/21.12 sortmerna/20170809

PATH_DB="/usr/users/zicola/genomes/silva_rRNA_maize/smr_database"

i=$(sed -n ${SLURM_ARRAY_TASK_ID}p trimmed_reads/list.txt)

sortmerna -a 1 --paired_in --fastx \
    --ref ${PATH_DB}/smr_v4.3_default_db.fasta,${PATH_DB}/smr_v4.3_default_db.idx \
    --reads trimmed_reads/merged_fastq/${i}.merged.fq \
    --aligned rRNA_removed/${i}.aligned \
    --other rRNA_removed/${i}.mRNA

```

```{bash, eval=FALSE}
srun --pty -p medium --mem=50G --cpus-per-task=5 --time=01:00:00  bash

i="02_Pm"

module purge

module load rev/21.12 sortmerna/20170809

PATH_DB="/usr/users/zicola/genomes/silva_rRNA_maize/smr_database"

sortmerna -a 5 --paired_in --fastx \
    --ref ${PATH_DB}/smr_v5.3_default_db.fasta,${PATH_DB}/smr_v4.3_default_db.idx \
    --reads trimmed_reads/merged_fastq/${i}.merged.fq \
    --aligned rRNA_removed/${i}.aligned \
    --other rRNA_removed/${i}.mRNA

```

Not one read mapped to the smr_v4.3_default_db database. This is very suspicious.


```{bash, eval=FALSE}

i="08_Ar"

sortmerna -a 5 --paired_in --fastx \
    --ref ${PATH_DB}/smr_v4.3_default_db.fasta,${PATH_DB}/smr_v4.3_default_db.idx \
    --reads trimmed_reads/merged_fastq/${i}.merged.fq \
    --aligned rRNA_removed/${i}.aligned \
    --other rRNA_removed/${i}.mRNA

```

Still nothing mapping to rRNA database.



## Unmerge fastq

```{bash, eval=FALSE}
#!/bin/bash
#SBATCH --job-name=unmerge_fastq
#SBATCH --nodes=1
#SBATCH --time=00:40:00
#SBATCH --cpus-per-task=1
#SBATCH --mem=20gb
#SBATCH --partition=medium
#SBATCH --mail-type=END
#SBATCH --mail-user=johan.zicola@uni-goettingen.de
#SBATCH --output=sortmerna.%J.out
#SBATCH --error=sortmerna.%J.err
#SBATCH --array=1-8

mem_gbytes=$(( $SLURM_MEM_PER_NODE/1024 ))

echo "Starting at $(date)"
echo "Job submitted to the ${SLURM_JOB_PARTITION} partition, the default partition on ${SLURM_CLUSTER_NAME}"
echo "Job name: ${SLURM_JOB_NAME}, Job ID: ${SLURM_JOB_ID}"
echo "I have ${SLURM_CPUS_ON_NODE} CPUs and ${mem_gbytes}GiB of RAM on compute node $(hostname)"

module purge

module load rev/21.12 sortmerna/20170809

i=$(sed -n ${SLURM_ARRAY_TASK_ID}p trimmed_reads/list.txt)

bash /usr/product/bioinfo/SL_7.0/BIOINFORMATICS/SORTMERNA/sortmerna_20170809/scripts/merge-paired-reads.sh rRNA_removed/${i}.mRNA.fq rRNA_removed/${i}.R1.mRNA.fq rRNA_removed/${i}.R2.mRNA.fq


```

## Use bowtie2

Index the database from sortmerna

```{bash, eval=FALSE}
cd /usr/users/zicola/genomes/silva_rRNA_maize/smr_database

bowtie2-build -f smr_v4.3_default_db.fasta smr_v4_3_default_db
```

```{bash, eval=FALSE}
#!/bin/bash
#SBATCH --job-name=rRNA_removal
#SBATCH --nodes=1
#SBATCH --time=02:00:00
#SBATCH --cpus-per-task=10
#SBATCH --mem=50gb
#SBATCH --partition=medium
#SBATCH --mail-type=END
#SBATCH --mail-user=johan.zicola@uni-goettingen.de
#SBATCH --output=rRNA_removal.%J.out
#SBATCH --error=rRNA_removal.%J.err
#SBATCH --array=1-8

mem_gbytes=$(( $SLURM_MEM_PER_NODE/1024 ))

echo "Starting at $(date)"
echo "Job submitted to the ${SLURM_JOB_PARTITION} partition, the default partition on ${SLURM_CLUSTER_NAME}"
echo "Job name: ${SLURM_JOB_NAME}, Job ID: ${SLURM_JOB_ID}"
echo "I have ${SLURM_CPUS_ON_NODE} CPUs and ${mem_gbytes}GiB of RAM on compute node $(hostname)"

module purge
module load rev/21.12 bowtie2

i=$(sed -n ${SLURM_ARRAY_TASK_ID}p trimmed_reads/list.txt)

bowtie2 -p 10 --nofw --quiet --very-sensitive-local --phred33  \
    -x /usr/users/zicola/genomes/silva_rRNA_maize/smr_database/smr_v4_3_default_db \
    -1 trimmed_reads/${i}_L1_1.trimmed.fq  \
    -2 trimmed_reads/${i}_L1_2.trimmed.fq \
    --un-conc rRNA_removed/${i}.rRNA_removed.fq > /dev/null

```


```{bash, eval=FALSE}

# Trimmed reads
seqkit stats 01_Ad*
processed files:  2 / 2 [======================================] ETA: 0s. done
file                   format  type    num_seqs        sum_len  min_len  avg_len  max_len
01_Ad_L1_1.trimmed.fq  FASTQ   DNA   63,577,851  6,342,339,512        1     99.8      100
01_Ad_L1_2.trimmed.fq  FASTQ   DNA   63,577,851  6,340,577,359        1     99.7      100


# Trimmed reads and rRNA removal

seqkit stats 01_Ad*
processed files:  2 / 2 [======================================] ETA: 0s. done
file                     format  type    num_seqs        sum_len  min_len  avg_len  max_len
01_Ad.rRNA_removed.1.fq  FASTQ   DNA   63,572,322  6,341,800,686        1     99.8      100
01_Ad.rRNA_removed.2.fq  FASTQ   DNA   63,572,322  6,340,036,951        1     99.7      100


```

That is 5529 (63577851-63572322) reads removed out of 63M (.00877619%) so I really doubt rRNA contamination would affect the assembly.



Do on all samples

```{bash, eval=FALSE}
cat stat.txt
processed files:  16 / 16  ETA: 0s. done
file                      format  type    num_seqs        sum_len  min_len  avg_len  max_len
01_Ad.rRNA_removed.1.fq   FASTQ   DNA   63,572,322  6,341,800,686        1     99.8      100
01_Ad.rRNA_removed.2.fq   FASTQ   DNA   63,572,322  6,340,036,951        1     99.7      100
02_Pm.rRNA_removed.1.fq   FASTQ   DNA   55,259,617  5,507,706,345       25     99.7      100
02_Pm.rRNA_removed.2.fq   FASTQ   DNA   55,259,617  5,505,907,480       25     99.6      100
03_Zt.rRNA_removed.1.fq   FASTQ   DNA   49,126,313  4,904,813,047       25     99.8      100
03_Zt.rRNA_removed.2.fq   FASTQ   DNA   49,126,313  4,903,228,530       25     99.8      100
04_Asp.rRNA_removed.1.fq  FASTQ   DNA   41,798,830  4,169,156,197       25     99.7      100
04_Asp.rRNA_removed.2.fq  FASTQ   DNA   41,798,830  4,168,954,996       25     99.7      100
05_G.rRNA_removed.1.fq    FASTQ   DNA   40,895,705  4,078,580,793       25     99.7      100
05_G.rRNA_removed.2.fq    FASTQ   DNA   40,895,705  4,077,039,215       25     99.7      100
06_Ol.rRNA_removed.1.fq   FASTQ   DNA   56,412,659  5,630,612,260       25     99.8      100
06_Ol.rRNA_removed.2.fq   FASTQ   DNA   56,412,659  5,628,472,850       25     99.8      100
07_Pr.rRNA_removed.1.fq   FASTQ   DNA   47,066,726  4,697,862,680       25     99.8      100
07_Pr.rRNA_removed.2.fq   FASTQ   DNA   47,066,726  4,696,252,248       25     99.8      100
08_Ar.rRNA_removed.1.fq   FASTQ   DNA   43,705,177  4,354,160,178       25     99.6      100
08_Ar.rRNA_removed.2.fq   FASTQ   DNA   43,705,177  4,352,752,476       25     99.6      100

```

Compare with number of reads after trimming

| sample      | reads after trimming | reads after bowtie2 rRNA filtering | difference | % trimmed reads |
|-------------|----------------------|------------------------------------|------------|-----------------|
| 01_Ad_L1_1  | 63,577,928           | 63,572,322                         | 5,606      | 0.0088%         |
| 01_Ad_L1_2  | 63,577,928           | 63,572,322                         | 5,606      | 0.0088%         |
| 02_Pm_L1_1  | 55,269,387           | 55,259,617                         | 9,770      | 0.0177%         |
| 02_Pm_L1_2  | 55,269,387           | 55,259,617                         | 9,770      | 0.0177%         |
| 03_Zt_L1_1  | 49,128,635           | 49,126,313                         | 2,322      | 0.0047%         |
| 03_Zt_L1_2  | 49,128,635           | 49,126,313                         | 2,322      | 0.0047%         |
| 04_Asp_L1_1 | 41,800,974           | 41,798,830                         | 2,144      | 0.0051%         |
| 04_Asp_L1_2 | 41,800,974           | 41,798,830                         | 2,144      | 0.0051%         |
| 05_G_L1_1   | 40,906,340           | 40,895,705                         | 10,635     | 0.0260%         |
| 05_G_L1_2   | 40,906,340           | 40,895,705                         | 10,635     | 0.0260%         |
| 06_Ol_L1_1  | 56,417,880           | 56,412,659                         | 5,221      | 0.0093%         |
| 06_Ol_L1_2  | 56,417,880           | 56,412,659                         | 5,221      | 0.0093%         |
| 07_Pr_L1_1  | 47,070,196           | 47,066,726                         | 3,470      | 0.0074%         |
| 07_Pr_L1_2  | 47,070,196           | 47,066,726                         | 3,470      | 0.0074%         |
| 08_Ar_L1_1  | 43,759,260           | 43,705,177                         | 54,083     | 0.1236%         |
| 08_Ar_L1_2  | 43,759,260           | 43,705,177                         | 54,083     | 0.1236%         |

The rRNA contamination


## Use bowtie2 on Arb Silva Coleoptera database

Search for taxonomy Coleoptera. Download fasta without gaps for the LSU and SSU hits (https://www.arb-silva.de/download)

+ 13765 SSU
+ 8776 LSU

```{bash, eval=FALSE}

cd /usr/users/zicola/genomes/rRNA_coleoptera

tar -zxvf arb-silva.de_2025-01-16_id1372508.tgz
mv arb-silva.de_2025-01-16_id1372508_tax_silva_trunc.fasta LSU_coleoptera.fasta

tar -zxvf arb-silva.de_2025-01-16_id1372510.tgz
mv arb-silva.de_2025-01-16_id1372510_tax_silva_trunc.fasta SSU_coleoptera.fasta

seqkit stats *fasta
processed files:  2 / 2 [======================================] ETA: 0s. done
file                  format  type  num_seqs     sum_len  min_len  avg_len  max_len
LSU_coleoptera.fasta  FASTA   RNA      8,776   5,944,179      300    677.3    3,238
SSU_coleoptera.fasta  FASTA   RNA     13,768  16,277,792      300  1,182.3    2,645

cat LSU_coleoptera.fasta SSU_coleoptera.fasta > LSU_SSU_coleoptera.fasta

seqkit stats LSU_SSU_coleoptera.fasta
file                      format  type  num_seqs     sum_len  min_len  avg_len  max_len
LSU_SSU_coleoptera.fasta  FASTA   RNA     22,544  22,221,971      300    985.7    3,238


bowtie2-build -f LSU_SSU_coleoptera.fasta LSU_SSU_coleoptera

```

Test with Ad

```{bash, eval=FALSE}
#!/bin/bash
#SBATCH --job-name=rRNA_removal
#SBATCH --nodes=1
#SBATCH --time=02:00:00
#SBATCH --cpus-per-task=10
#SBATCH --mem=50gb
#SBATCH --partition=medium
#SBATCH --mail-type=END
#SBATCH --mail-user=johan.zicola@uni-goettingen.de
#SBATCH --output=rRNA_removal.%J.out
#SBATCH --error=rRNA_removal.%J.err

mem_gbytes=$(( $SLURM_MEM_PER_NODE/1024 ))

echo "Starting at $(date)"
echo "Job submitted to the ${SLURM_JOB_PARTITION} partition, the default partition on ${SLURM_CLUSTER_NAME}"
echo "Job name: ${SLURM_JOB_NAME}, Job ID: ${SLURM_JOB_ID}"
echo "I have ${SLURM_CPUS_ON_NODE} CPUs and ${mem_gbytes}GiB of RAM on compute node $(hostname)"

module purge
module load rev/21.12 bowtie2

bowtie2 -p 10 --nofw --quiet --very-sensitive-local --phred33  \
    -x /usr/users/zicola/genomes/rRNA_coleoptera/LSU_SSU_coleoptera \
    -1 trimmed_reads/01_Ad_L1_1.trimmed.fq  \
    -2 trimmed_reads/01_Ad_L1_2.trimmed.fq \
    --un-conc rRNA_removed/01_Ad.rRNA_removed_silva_coleptera.fq > /dev/null

```

```{bash, eval=FALSE}
sbatch test_bowtie2.sh
```


```{bash, eval=FALSE}
# Bowtie using smr database

seqkit stats 01_Ad*
processed files:  2 / 2 [======================================] ETA: 0s. done
file                     format  type    num_seqs        sum_len  min_len  avg_len  max_len
01_Ad.rRNA_removed.1.fq  FASTQ   DNA   63,572,322  6,341,800,686        1     99.8      100
01_Ad.rRNA_removed.2.fq  FASTQ   DNA   63,572,322  6,340,036,951        1     99.7      100


# Bowtie using ARB Silva Coleoptera SSU + LSU database

seqkit stats 01_Ad.rRNA_removed_silva_coleptera.*
processed files:  2 / 2 [======================================] ETA: 0s. done
file                                     format  type    num_seqs        sum_len  min_len  avg_len  max_len
01_Ad.rRNA_removed_silva_coleptera.1.fq  FASTQ   DNA   63,574,766  6,342,041,048        1     99.8      100
01_Ad.rRNA_removed_silva_coleptera.2.fq  FASTQ   DNA   63,574,766  6,340,277,246        1     99.7      100


```

63574766-63572322 = 2444. I can actually less reads mapping rRNA when using the Silva database for Coleptera (22,544 sequences) compared to the smr default database (73,977 sequences).



## test sortmerna

```{bash, eval=FALSE}

# Take first entry of smr_v4.3_default_db.fasta and put it in seq_rRNA.fa


cd /usr/users/zicola/genomes/sortmerna_database/test

python ~/git_repositories/fastq_generator/fastq_generator.py generate_mapped_fastq_PE seq_rRNA.fa 50 10 10 > fastq_interleaved.fq

# Split in two files
seqkit stats fastq_interleaved.fq
file                  format  type  num_seqs  sum_len  min_len  avg_len  max_len
fastq_interleaved.fq  FASTQ   DNA        520   26,000       50       50       50

head -n 1040 fastq_interleaved.fq > fastq_interleaved.R1.fq
tail -n 1040 fastq_interleaved.fq > fastq_interleaved.R2.fq

module load rev/21.12 sortmerna/20170809

# Get them merged (each pair is together)
/usr/product/bioinfo/SL_7.0/BIOINFORMATICS/SORTMERNA/sortmerna_20170809/scripts/merge-paired-reads.sh fastq_interleaved.R1.fq fastq_interleaved.R2.fq  fastq_interleaved.merged.fq

sortmerna -a 1 --paired_in --fastx \
    --ref ../smr_v4.3_default_db.fasta,../smr_v4.3_default_db.idx \
    --reads fastq_interleaved.merged.fq \
    --aligned aligned \
    --other mRNA -v
# Nothing aligned

sortmerna -a 1 --fastx \
    --ref ../smr_v4.3_default_db.fasta,../smr_v4.3_default_db.idx \
    --reads fastq_interleaved.R1.fq --reads fastq_interleaved.R2.fq \
    --aligned aligned \
    --other mRNA -v




sortmerna -a 1 --fastx \
    --ref ../smr_v4.3_default_db.fasta,../smr_v4.3_default_db.idx \
    --reads fastq_interleaved.merged.fq \
    --aligned aligned \
    --other mRNA -v


sortmerna -a 1 --fastx \
    --ref ../smr_v4.3_default_db.fasta,../smr_v4.3_default_db.idx \
    --reads fastq_interleaved.R1.fq \
    --aligned aligned \
    --other mRNA -v
    
    
sortmerna -a 1 --fastx \
    --ref ../smr_v4.3_default_db.fasta,../smr_v4.3_default_db.idx \
    --reads tests/data/set2_environmental_study_550_amplicon.fasta \
    --aligned aligned \
    --other mRNA -v
    
sortmerna -a 1 --fastx \
    --ref ../smr_v4.3_default_db.fasta,../smr_v4.3_default_db.idx \
    --reads tests/data/set2_environmental_study_550_amplicon.fasta \
    --aligned aligned \
    --other mRNA -v
    
```

Not working

```{bash, eval=FALSE}
module load rev/21.12 sortmerna/20170809

indexdb --ref seq_rRNA.fa,seq_rRNA.idx -v

sortmerna -a 1 --fastx \
    --ref seq_rRNA.fa,seq_rRNA.idx \
    --reads fastq_interleaved.R1.fq \
    --aligned aligned \
    --other mRNA -v

# Now it worked! Everything is in aligned.fq

sortmerna -a 1 --fastx \
    --ref smr_v4.3_default_db.fasta,smr_v4.3_default_db.idx \
    --reads test/fastq_interleaved.R1.fq \
    --aligned aligned \
    --other mRNA -v

# Not working
```

I think the DB is corrupted.

Remake the indexes of smr_v4.3_default_db.fasta

```{bash, eval=FALSE}

srun --pty -t 00:50:00 --mem=20G -p medium bash

module load rev/21.12 sortmerna/20170809

indexdb --ref smr_v4.3_default_db.fasta,smr_v4.3_default_db.idx -v

```

```{bash, eval=FALSE}
# Single end mode
sortmerna -a 1 --fastx \
    --ref smr_v4.3_default_db.fasta,smr_v4.3_default_db.idx \
    --reads test/fastq_interleaved.R1.fq \
    --aligned aligned \
    --other mRNA -v
```

Now it works!

```{bash, eval=FALSE}
# Paired end mode
sortmerna -a 1 --paired_in --fastx \
    --ref ../smr_v4.3_default_db.fasta,../smr_v4.3_default_db.idx \
    --reads fastq_interleaved.merged.fq \
    --aligned aligned \
    --other mRNA -v
```
 Works

# Trinotate

Wiki: https://github.com/Trinotate/Trinotate/wiki/Software-installation-and-data-required


Reuse script used for OPM analysis.


```{bash, eval=FALSE}

#!/bin/bash
#SBATCH --job-name=trinotate
#SBATCH --nodes=1
#SBATCH --partition=medium
#SBATCH --time=48:00:00
#SBATCH --cpus-per-task=8
#SBATCH --mem=40gb
#SBATCH --mail-type=END
#SBATCH --mail-user=johan.zicola@uni-goettingen.de
#SBATCH --output=trinotate.%J.out
#SBATCH --error=trinotate.%J.err


mem_gbytes=$(( $SLURM_MEM_PER_NODE/1024 ))

echo "Starting at $(date)"
echo "Job submitted to the ${SLURM_JOB_PARTITION} partition, the default partition on ${SLURM_CLUSTER_NAME}"
echo "Job name: ${SLURM_JOB_NAME}, Job ID: ${SLURM_JOB_ID}"
echo "I have ${SLURM_CPUS_ON_NODE} CPUs and ${mem_gbytes}GiB of RAM on compute node $(hostname)"

# Modules to load
module purge

module load rev/21.12 gcc/10.2.0 hmmer/3.3 ncbi-blast/2.11.0

# Test that all programs needed are installed
test_program(){
  command -v $1 >/dev/null || \
    { echo >&2 " $1 required but not installed. Aborting"; exit 1; }
}

test_program "diamond"
test_program "cmpress"
test_program "hmmpress"
test_program "tmhmm"
test_program "signalp6"
test_program "makeblastdb"

# Variables passed to sbatch
# NAME_SPECIES
# Output from Trinity $GENE_TRANS_MAP and $TRINITY_FASTA
# Output from Transdecoder $TRANSDECODER_PEP

# Variables
TRINOTATE_HOME="/usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1"

#NAME_SPECIES is provided as argument to sbatch
sqlite_db="/usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1/Trinotate_${NAME_SPECIES}.sqlite"


echo "############################"
echo "Trinotate create"
date
echo "############################"

$TRINOTATE_HOME/Trinotate --create \
    --db $sqlite_db \
    --trinotate_data_dir $TRINOTATE_HOME/TRINOTATE_DATA_DIR \
    --use_diamond

echo "###############################"
echo "Trinotate Init"
date
echo "###############################"


$TRINOTATE_HOME/Trinotate --db $sqlite_db --init \
           --gene_trans_map $GENE_TRANS_MAP \
           --transcript_fasta $TRINITY_FASTA \
           --transdecoder_pep $TRANSDECODER_PEP

echo "##############################"
echo  "Trinotate Run"
date
echo "##############################"
 
$TRINOTATE_HOME/Trinotate --db ${sqlite_db} --CPU ${SLURM_CPUS_ON_NODE} --run ALL \
  --trinotate_data_dir $TRINOTATE_HOME/TRINOTATE_DATA_DIR \
    --transcript_fasta $TRINITY_FASTA \
      --transdecoder_pep $TRANSDECODER_PEP \
      --use_diamond

echo "###########################"
echo "Generating report table"
date
echo "###########################"

$TRINOTATE_HOME/Trinotate --db ${sqlite_db} \
  --report --incl_pep --incl_trans > Trinotate_report.tsv

##################
## Misc value adds
##################

# Extract GO terms
${TRINOTATE_HOME}/util/extract_GO_assignments_from_Trinotate_xls.pl \
  --Trinotate_xls Trinotate_report.tsv -G -I > Trinotate_report.tsv.gene_ontology

# Generate trinotate report summary statistics
${TRINOTATE_HOME}/util/report_summary/trinotate_report_summary.pl \
  Trinotate_report.tsv Trinotate_report_stats

echo "##########################"
echo "done.  See annotation summary file:  Trinotate_report.tsv"
date
echo "##########################"



```

```{bash, eval=FALSE}


# Variables passed to sbatch
# Output from Trinity $GENE_TRANS_MAP and $TRINITY_FASTA
# Output from Transdecoder $TRANSDECODER_PEP

work_path="/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects"
while read i; do
  mkdir $i
  cd $i 
  sbatch --export=NAME_SPECIES=$i,GENE_TRANS_MAP=${work_path}/trinity_output/summary/${i}_trinity.fasta.gene_trans_map,TRINITY_FASTA=${work_path}/trinity_output/summary/${i}_trinity.fasta,TRANSDECODER_PEP=${work_path}/transdecoder_output/${i}_trinity.fasta.transdecoder.pep ../04_trinotate.sh
  cd ..
done < list_species.txt

```


Steps done before being killed

```{bash, eval=FALSE}

Running CMD: /usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1/Trinotate --db /usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1/Trinotate_Anchomenus_dorsalis.sqlite --LOAD_swissprot_blastp uniprot_sprot.diamond.blastp.outfmt6
-LOADING as per --LOAD_swissprot_blastp

CMD: /home/uni08/zicola/bin/Trinotate-Trinotate-v4.0.1/util/trinotateSeqLoader/Trinotate_BLAST_loader.pl --sqlite /usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1/Trinotate_Anchomenus_dorsalis.sqlite --outfmt6 uniprot_sprot.diamond.blastp.outfmt6 --prog blastp --dbtype Swissprot
Can't locate DBI.pm in @INC (you may need to install the DBI module) (@INC contains: /home/uni08/zicola/bin/Trinotate-Trinotate-v4.0.1/util/trinotateSeqLoader/../../PerlLib /usr/users/zicola/perl5/lib/perl5/x86_64-linux-thread-multi /usr/users/zicola/perl5/lib/perl5 /usr/local/lib64/perl5 /usr/local/share/perl5 /usr/lib64/perl5/vendor_perl /usr/share/perl5/vendor_perl /usr/lib64/perl5 /usr/share/perl5) at /home/uni08/zicola/bin/Trinotate-Trinotate-v4.0.1/util/trinotateSeqLoader/Trinotate_BLAST_loader.pl line 9.
BEGIN failed--compilation aborted at /home/uni08/zicola/bin/Trinotate-Trinotate-v4.0.1/util/trinotateSeqLoader/Trinotate_BLAST_loader.pl line 9.
Error, cmd: /home/uni08/zicola/bin/Trinotate-Trinotate-v4.0.1/util/trinotateSeqLoader/Trinotate_BLAST_loader.pl --sqlite /usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1/Trinotate_Anchomenus_dorsalis.sqlite --outfmt6 uniprot_sprot.diamond.blastp.outfmt6 --prog blastp --dbtype Swissprot died with ret 512 at /usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1/Trinotate line 338.
Error, cmd: /usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1/Trinotate --db /usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1/Trinotate_Anchomenus_dorsalis.sqlite --LOAD_swissprot_blastp uniprot_sprot.diamond.blastp.outfmt6 died with ret 512 No such file or directory at /home/uni08/zicola/bin/Trinotate-Trinotate-v4.0.1/PerlLib/Pipeliner.pm line 187.
        Pipeliner::run(Pipeliner=HASH(0x56249c3268b0)) called at /usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1/Trinotate line 736
        main::Trinotate_RUN_annotators("/usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1/Trinotate_An"..., ARRAY(0x56249bf92f20), 8, "/usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1/TRINOTATE_DA"..., "/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgs"..., "/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgs"...) called at /usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1/Trinotate line 308
-REPORT being generated.
CMD: /home/uni08/zicola/bin/Trinotate-Trinotate-v4.0.1/util/Trinotate_report_writer.pl --sqlite /usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1/Trinotate_Anchomenus_dorsalis.sqlite -E 1e-5 --pfam_cutoff DNC  --incl_pep  --incl_trans
Can't locate DBI.pm in @INC (you may need to install the DBI module) (@INC contains: /home/uni08/zicola/bin/Trinotate-Trinotate-v4.0.1/util/../PerlLib /usr/users/zicola/perl5/lib/perl5/x86_64-linux-thread-multi /usr/users/zicola/perl5/lib/perl5 /usr/local/lib64/perl5 /usr/local/share/perl5 /usr/lib64/perl5/vendor_perl /usr/share/perl5/vendor_perl /usr/lib64/perl5 /usr/share/perl5) at /home/uni08/zicola/bin/Trinotate-Trinotate-v4.0.1/util/../PerlLib/Sqlite_connect.pm line 15.


```



## Troubleshoot perl library installation

After the migration, I had to reinstall my perl libraries. I notice that some of the libraries I installed before cannot be reinstalled (`cpan install DBI`) due to a `cc1: fatal error: inaccessible plugin file /opt/sw/rev/21.12/haswell/gcc-4.8.5/gcc-10.2.0-potfoy/lib/gcc/x86_64-pc-linux-gnu/10.2.0/plugin/gcc-annobin.so expanded from short plugin name gcc-annobin: No such file or directory` error. I found on the web several posts about the annobin issue and one suggests installing a local Perl installation with Spack (https://github.com/openhpc/ohpc/issues/1364).

I may think this 'annobin' gcc plugin was present on the previous SCC nodes OS (Scientific Linux 7) but was probably removed in the current OS (Rocky Linux 8). It seems that the GCC plugin annobin can be installed on Rocky 8 (https://rockylinux.pkgs.org/8/rockylinux-devel-x86\_64/annobin-11.13-2.el8.x86\_64.rpm.html). Would it be possible to install it or do I have to install my own gcc?

Note that I tried with all versions of gcc available and all failed:

```
module purge
module load rev/20.12 gcc/7.5.0
cpan instal DBI
#cc1: error: unrecognized command line option '-fstack-clash-protection'
#cc1: error: unrecognized command line option '-fcf-protection'
#cc1: fatal error: inaccessible plugin file /opt/sw/rev/20.12/haswell/gcc-9.3.0/gcc-7.5.0-owlabd/lib/gcc/x86_64-pc-linux-gnu/7.5.0/plugin/gcc-annobin.so expanded from short plugin name gcc-annobin: No such file or directory

# Nope
module purge
module load rev/20.12 gcc/8.4.0
cpan instal DBI
#cc1: fatal error: inaccessible plugin file /opt/sw/rev/20.12/haswell/gcc-4.8.5/gcc-8.4.0-anizkf/lib/gcc/x86_64-pc-linux-gnu/8.4.0/plugin/gcc-annobin.so expanded from short plugin name gcc-annobin: No such file or directory

module purge
module load rev/20.12 gcc/9.3.0
cpan instal DBI
# cc1: fatal error: inaccessible plugin file /opt/sw/rev/20.12/haswell/gcc-8.4.0/gcc-9.3.0-a466o5/lib/gcc/x86_64-pc-linux-gnu/9.3.0/plugin/gcc-annobin.so expanded from short plugin name gcc-annobin: No such file or directory

module purge
module load rev/20.12  gcc/9.5.0
cpan instal DBI
# cc1: fatal error: inaccessible plugin file /opt/sw/rev/23.12/linux-scientific7-haswell/gcc-4.8.5/gcc-9.5.0-y5n2mx/lib/gcc/x86_64-pc-linux-gnu/9.5.0/plugin/gcc-annobin.so expanded from short plugin name gcc-annobin: No such file or directory


module purge
module load rev/20.12 gcc/10.2.0
cpan instal DBI
# cc1: fatal error: inaccessible plugin file /opt/sw/rev/20.12/haswell/gcc-8.4.0/gcc-10.2.0-tc66mg/lib/gcc/x86_64-pc-linux-gnu/10.2.0/plugin/gcc-annobin.so expanded from short plugin name gcc-annobin: No such file or directory

module purge
module load gcc/11.4.0
cpan instal DBI
#cc1: fatal error: inaccessible plugin file /opt/sw/rev/23.12/linux-scientific7-haswell/gcc-4.8.5/gcc-11.4.0-jgoe5h/lib/gcc/x86_64-pc-linux-gnu/11.4.0/plugin/gcc-annobin.so expanded from short plugin name gcc-annobin: No such file or directory
```

Red Hat documentation on Annobin: https://docs.redhat.com/en/documentation/red\_hat\_developer\_toolset/9/html/user\_guide/chap-annobin#chap-Annobin


I tried `spack install perl` but the installation failed

```
ImportError: cannot bootstrap the "clingo" Python module from spec "clingo-bootstrap@spack+python %gcc target=x86_64" due to the following failures:
    'spack-install' raised InstallError: Terminating after first install failure: ProcessError: Command exited with status 1:
    '/usr/users/zicola/.spack/bootstrap/store/linux-rocky8-x86_64/gcc-10.2.0/perl-5.34.0-hhpsqrp435rgen7twicgyvpz5tpy5tbs/bin/perl' 'Makefile.PL'
    Please run `spack -d spec zlib` for more verbose error messages
```

I tried several times `spack -d spec zlib` but it does not work.



Mnay posts about the issue with annobin. One recommendation is to install perl in local using spack.


Download archive https://metacpan.org/pod/DBI and install manually DBI-1.647.tgz


```{bash, eval=FALSE}
cd ~/bin
tar -zxvf DBI-1.647.tgz
cd DBI-1.647

perl Makefile.PL PREFIX=/usr/users/zicola/perl5/lib/perl5
make

#cc1: fatal error: inaccessible plugin file /opt/sw/rev/23.12/linux-scientific7-haswell/gcc-4.8.5/gcc-11.4.0-jgoe5h/lib/gcc/x86_64-pc-linux-gnu/11.4.0/plugin/gcc-annobin.so expanded from short plugin name gcc-annobin: No such file or directory

#make test
#make install

```


```{bash, eval=FALSE}

spack install perl

# Error message, recommends to type 'spack -d spec zlib'
spack -d spec zlib


```





