---
title: "other_scripts"
author: "Johan Zicola"
date: '2025-01-16'
output: html_document
---

# Pipeline

https://southgreenplatform.github.io/tutorials//bioanalysis/trinity/

https://training.galaxyproject.org/topics/transcriptomics/tutorials/full-de-novo/tutorial.html

* Trim data with cutadapt to remove adapters

# Remove empty reads

```{bash, eval=FALSE}

while read i; do
  srun -p medium -t 01:20:00 --cpus-per-task=10 cutadapt -j 10 --minimum-length 25 \
  -o ${i}_L1_1.trimmed.min25.fq -p ${i}_L1_2.trimmed.min25.fq \
  ${i}_L1_1.trimmed.fq ${i}_L1_2.trimmed.fq &
done < list.txt

```

# Trimming with trimmomatic

```{bash, eval=FALSE}

module load rev/21.12 trimmomatic openjdk

PATH_ADAPTER="/usr/product/bioinfo/SL_7.0/BIOINFORMATICS/TRIMMOMATIC/0.36/adapters/"

srun -p medium -t 00:30:00 --mem=12G --cpus-per-task=10 \
java -Xmx10G -jar /usr/product/bioinfo/SL_7.0/BIOINFORMATICS/TRIMMOMATIC/0.36/trimmomatic-0.36.jar \
PE -phred33 -threads 10 \
raw_data/01_Ad/01_Ad_L1_1.fq.gz \
raw_data/01_Ad/01_Ad_L1_2.fq.gz \
trimmed_reads_trimmomatic/01_Ad_L1_1.paired.fq.gz \
trimmed_reads_trimmomatic/01_Ad_L1_1.unpaired.fq.gz \
trimmed_reads_trimmomatic/01_Ad_L1_2.paired.fq.gz \
trimmed_reads_trimmomatic/01_Ad_L1_2.unpaired.fq.gz \
ILLUMINACLIP:"/usr/product/bioinfo/SL_7.0/BIOINFORMATICS/TRIMMOMATIC/0.36/adapters/TruSeq2-SE.fa":2:30:10 LEADING:3 TRAILING:3 SLIDINGWINDOW:4:15 MINLEN:50 &

Input Read Pairs: 64769548 Both Surviving: 63780516 (98.47%) Forward Only Surviving: 228552 (0.35%) Reverse Only Surviving: 744814 (1.15%) Dropped: 15666 (0.02%)
TrimmomaticPE: Completed successfully

seqkit stats 01_Ad_L1_1.paired.fq.gz 01_Ad_L1_2.paired.fq.gz
processed files:  2 / 2 [======================================] ETA: 0s. done
file                     format  type    num_seqs        sum_len  min_len  avg_len  max_len
01_Ad_L1_1.paired.fq.gz  FASTQ   DNA   63,780,516  6,351,162,309       50     99.6      100
01_Ad_L1_2.paired.fq.gz  FASTQ   DNA   63,780,516  6,350,850,563       50     99.6      100

# Same number of reads in each file

module load fastqc/0.11.4

srun -p medium -t 00:50:00 fastqc 01_Ad_L1_2.paired.fq.gz &
srun -p medium -t 00:50:00 fastqc 01_Ad_L1_1.paired.fq.gz &

```


# Remove rRNA

https://github.com/sortmerna/sortmerna

Download v4.3 database (https://github.com/biocore/sortmerna/releases/download/v4.3.4/database.tar.gz)

## Index database

The docu recommends to use `smr_v4.3_default_db.fasta`

```{r}
mkdir ~/genomes/silva_rRNA_maize/smr_database && cd ~/genomes/silva_rRNA_maize/smr_database

wget https://github.com/biocore/sortmerna/releases/download/v4.3.4/database.tar.gz

tar -zxvf database.tar.gz

seqkit stats smr_v4.3_default_db.fasta
file                       format  type  num_seqs      sum_len  min_len  avg_len  max_len
smr_v4.3_default_db.fasta  FASTA   RNA     73,977  133,533,894       91  1,805.1    4,380

module load rev/21.12 sortmerna/20170809

srun -t 00:50:00 --mem=20G -p medium indexdb_rna --ref smr_v4.3_default_db.fasta,smr_v4.3_default_db.idx -v

```


```{bash, eval=FALSE}
\ls -1sh smr_v4.3_default*
140M smr_v4.3_default_db.fasta
662M smr_v4.3_default_db.idx.bursttrie_0.dat
1.0M smr_v4.3_default_db.idx.kmer_0.dat
1.2G smr_v4.3_default_db.idx.pos_0.dat
4.0M smr_v4.3_default_db.idx.stats
```

4 index files are generated with a *dat extension.

## Merge reads

I tested with the R1 and R2 in separate files using two --read flags but the output contained only R2 sequences. When I merge the R1 and R2, it works. I just need to merge R1 and R2 and unmerge them afterwards

```{bash, eval=FALSE}

cd /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trimmed_reads

mkdir merged_fastq

module load rev/21.12 sortmerna/20170809

while read i; do
if [ -e merged_fastq/${i}.merged.fq ]; then
  echo "$i already processed"
else
  srun -p medium --mem=2Gb --time=00:20:00 bash /usr/product/bioinfo/SL_7.0/BIOINFORMATICS/SORTMERNA/sortmerna_20170809/scripts/merge-paired-reads.sh ${i}_L1_1.trimmed.fq ${i}_L1_2.trimmed.fq  merged_fastq/${i}.merged.fq &
fi
done < list.txt


```


```{bash, eval=FALSE}
#!/bin/bash
#SBATCH --job-name=merge_fastq
#SBATCH --nodes=1
#SBATCH --time=00:40:00
#SBATCH --cpus-per-task=1
#SBATCH --mem=10gb
#SBATCH --partition=medium
#SBATCH --mail-type=END
#SBATCH --mail-user=johan.zicola@uni-goettingen.de
#SBATCH --output=sortmerna.%J.out
#SBATCH --error=sortmerna.%J.err
#SBATCH --array=1-8

mem_gbytes=$(( $SLURM_MEM_PER_NODE/1024 ))

echo "Starting at $(date)"
echo "Job submitted to the ${SLURM_JOB_PARTITION} partition, the default partition on ${SLURM_CLUSTER_NAME}"
echo "Job name: ${SLURM_JOB_NAME}, Job ID: ${SLURM_JOB_ID}"
echo "I have ${SLURM_CPUS_ON_NODE} CPUs and ${mem_gbytes}GiB of RAM on compute node $(hostname)"

module purge

module load rev/21.12 sortmerna/20170809

i=$(sed -n ${SLURM_ARRAY_TASK_ID}p trimmed_reads/list.txt)

bash /usr/product/bioinfo/SL_7.0/BIOINFORMATICS/SORTMERNA/sortmerna_20170809/scripts/merge-paired-reads.sh trimmed_reads/${i}_L1_1.trimmed.fq trimmed_reads/${i}_L1_2.trimmed.fq  trimmed_reads/merged_fastq/${i}.merged.fq

```


## sortmerna

```{r}
#!/bin/bash
#SBATCH --job-name=sortmerna
#SBATCH --nodes=1
#SBATCH --time=01:00:00
#SBATCH --cpus-per-task=10
#SBATCH --mem=50gb
#SBATCH --partition=medium
#SBATCH --mail-type=END
#SBATCH --mail-user=johan.zicola@uni-goettingen.de
#SBATCH --output=sortmerna.%J.out
#SBATCH --error=sortmerna.%J.err
#SBATCH --array=1-8

mem_gbytes=$(( $SLURM_MEM_PER_NODE/1024 ))

echo "Starting at $(date)"
echo "Job submitted to the ${SLURM_JOB_PARTITION} partition, the default partition on ${SLURM_CLUSTER_NAME}"
echo "Job name: ${SLURM_JOB_NAME}, Job ID: ${SLURM_JOB_ID}"
echo "I have ${SLURM_CPUS_ON_NODE} CPUs and ${mem_gbytes}GiB of RAM on compute node $(hostname)"

module purge

module load rev/21.12 sortmerna/20170809

PATH_DB="/usr/users/zicola/genomes/silva_rRNA_maize/smr_database"

i=$(sed -n ${SLURM_ARRAY_TASK_ID}p trimmed_reads/list.txt)

sortmerna -a 1 --paired_in --fastx \
    --ref ${PATH_DB}/smr_v4.3_default_db.fasta,${PATH_DB}/smr_v4.3_default_db.idx \
    --reads trimmed_reads/merged_fastq/${i}.merged.fq \
    --aligned rRNA_removed/${i}.aligned \
    --other rRNA_removed/${i}.mRNA

```

```{bash, eval=FALSE}
srun --pty -p medium --mem=50G --cpus-per-task=5 --time=01:00:00  bash

i="02_Pm"

module purge

module load rev/21.12 sortmerna/20170809

PATH_DB="/usr/users/zicola/genomes/silva_rRNA_maize/smr_database"

sortmerna -a 5 --paired_in --fastx \
    --ref ${PATH_DB}/smr_v5.3_default_db.fasta,${PATH_DB}/smr_v4.3_default_db.idx \
    --reads trimmed_reads/merged_fastq/${i}.merged.fq \
    --aligned rRNA_removed/${i}.aligned \
    --other rRNA_removed/${i}.mRNA

```

Not one read mapped to the smr_v4.3_default_db database. This is very suspicious.


```{bash, eval=FALSE}

i="08_Ar"

sortmerna -a 5 --paired_in --fastx \
    --ref ${PATH_DB}/smr_v4.3_default_db.fasta,${PATH_DB}/smr_v4.3_default_db.idx \
    --reads trimmed_reads/merged_fastq/${i}.merged.fq \
    --aligned rRNA_removed/${i}.aligned \
    --other rRNA_removed/${i}.mRNA

```

Still nothing mapping to rRNA database.



## Unmerge fastq

```{bash, eval=FALSE}
#!/bin/bash
#SBATCH --job-name=unmerge_fastq
#SBATCH --nodes=1
#SBATCH --time=00:40:00
#SBATCH --cpus-per-task=1
#SBATCH --mem=20gb
#SBATCH --partition=medium
#SBATCH --mail-type=END
#SBATCH --mail-user=johan.zicola@uni-goettingen.de
#SBATCH --output=sortmerna.%J.out
#SBATCH --error=sortmerna.%J.err
#SBATCH --array=1-8

mem_gbytes=$(( $SLURM_MEM_PER_NODE/1024 ))

echo "Starting at $(date)"
echo "Job submitted to the ${SLURM_JOB_PARTITION} partition, the default partition on ${SLURM_CLUSTER_NAME}"
echo "Job name: ${SLURM_JOB_NAME}, Job ID: ${SLURM_JOB_ID}"
echo "I have ${SLURM_CPUS_ON_NODE} CPUs and ${mem_gbytes}GiB of RAM on compute node $(hostname)"

module purge

module load rev/21.12 sortmerna/20170809

i=$(sed -n ${SLURM_ARRAY_TASK_ID}p trimmed_reads/list.txt)

bash /usr/product/bioinfo/SL_7.0/BIOINFORMATICS/SORTMERNA/sortmerna_20170809/scripts/merge-paired-reads.sh rRNA_removed/${i}.mRNA.fq rRNA_removed/${i}.R1.mRNA.fq rRNA_removed/${i}.R2.mRNA.fq


```

## Use bowtie2

Index the database from sortmerna

```{bash, eval=FALSE}
cd /usr/users/zicola/genomes/silva_rRNA_maize/smr_database

bowtie2-build -f smr_v4.3_default_db.fasta smr_v4_3_default_db
```

```{bash, eval=FALSE}
#!/bin/bash
#SBATCH --job-name=rRNA_removal
#SBATCH --nodes=1
#SBATCH --time=02:00:00
#SBATCH --cpus-per-task=10
#SBATCH --mem=50gb
#SBATCH --partition=medium
#SBATCH --mail-type=END
#SBATCH --mail-user=johan.zicola@uni-goettingen.de
#SBATCH --output=rRNA_removal.%J.out
#SBATCH --error=rRNA_removal.%J.err
#SBATCH --array=1-8

mem_gbytes=$(( $SLURM_MEM_PER_NODE/1024 ))

echo "Starting at $(date)"
echo "Job submitted to the ${SLURM_JOB_PARTITION} partition, the default partition on ${SLURM_CLUSTER_NAME}"
echo "Job name: ${SLURM_JOB_NAME}, Job ID: ${SLURM_JOB_ID}"
echo "I have ${SLURM_CPUS_ON_NODE} CPUs and ${mem_gbytes}GiB of RAM on compute node $(hostname)"

module purge
module load rev/21.12 bowtie2

i=$(sed -n ${SLURM_ARRAY_TASK_ID}p trimmed_reads/list.txt)

bowtie2 -p 10 --nofw --quiet --very-sensitive-local --phred33  \
    -x /usr/users/zicola/genomes/silva_rRNA_maize/smr_database/smr_v4_3_default_db \
    -1 trimmed_reads/${i}_L1_1.trimmed.fq  \
    -2 trimmed_reads/${i}_L1_2.trimmed.fq \
    --un-conc rRNA_removed/${i}.rRNA_removed.fq > /dev/null

```


```{bash, eval=FALSE}

# Trimmed reads
seqkit stats 01_Ad*
processed files:  2 / 2 [======================================] ETA: 0s. done
file                   format  type    num_seqs        sum_len  min_len  avg_len  max_len
01_Ad_L1_1.trimmed.fq  FASTQ   DNA   63,577,851  6,342,339,512        1     99.8      100
01_Ad_L1_2.trimmed.fq  FASTQ   DNA   63,577,851  6,340,577,359        1     99.7      100


# Trimmed reads and rRNA removal

seqkit stats 01_Ad*
processed files:  2 / 2 [======================================] ETA: 0s. done
file                     format  type    num_seqs        sum_len  min_len  avg_len  max_len
01_Ad.rRNA_removed.1.fq  FASTQ   DNA   63,572,322  6,341,800,686        1     99.8      100
01_Ad.rRNA_removed.2.fq  FASTQ   DNA   63,572,322  6,340,036,951        1     99.7      100


```

That is 5529 (63577851-63572322) reads removed out of 63M (.00877619%) so I really doubt rRNA contamination would affect the assembly.



Do on all samples

```{bash, eval=FALSE}
cat stat.txt
processed files:  16 / 16  ETA: 0s. done
file                      format  type    num_seqs        sum_len  min_len  avg_len  max_len
01_Ad.rRNA_removed.1.fq   FASTQ   DNA   63,572,322  6,341,800,686        1     99.8      100
01_Ad.rRNA_removed.2.fq   FASTQ   DNA   63,572,322  6,340,036,951        1     99.7      100
02_Pm.rRNA_removed.1.fq   FASTQ   DNA   55,259,617  5,507,706,345       25     99.7      100
02_Pm.rRNA_removed.2.fq   FASTQ   DNA   55,259,617  5,505,907,480       25     99.6      100
03_Zt.rRNA_removed.1.fq   FASTQ   DNA   49,126,313  4,904,813,047       25     99.8      100
03_Zt.rRNA_removed.2.fq   FASTQ   DNA   49,126,313  4,903,228,530       25     99.8      100
04_Asp.rRNA_removed.1.fq  FASTQ   DNA   41,798,830  4,169,156,197       25     99.7      100
04_Asp.rRNA_removed.2.fq  FASTQ   DNA   41,798,830  4,168,954,996       25     99.7      100
05_G.rRNA_removed.1.fq    FASTQ   DNA   40,895,705  4,078,580,793       25     99.7      100
05_G.rRNA_removed.2.fq    FASTQ   DNA   40,895,705  4,077,039,215       25     99.7      100
06_Ol.rRNA_removed.1.fq   FASTQ   DNA   56,412,659  5,630,612,260       25     99.8      100
06_Ol.rRNA_removed.2.fq   FASTQ   DNA   56,412,659  5,628,472,850       25     99.8      100
07_Pr.rRNA_removed.1.fq   FASTQ   DNA   47,066,726  4,697,862,680       25     99.8      100
07_Pr.rRNA_removed.2.fq   FASTQ   DNA   47,066,726  4,696,252,248       25     99.8      100
08_Ar.rRNA_removed.1.fq   FASTQ   DNA   43,705,177  4,354,160,178       25     99.6      100
08_Ar.rRNA_removed.2.fq   FASTQ   DNA   43,705,177  4,352,752,476       25     99.6      100

```

Compare with number of reads after trimming

| sample      | reads after trimming | reads after bowtie2 rRNA filtering | difference | % trimmed reads |
|-------------|----------------------|------------------------------------|------------|-----------------|
| 01_Ad_L1_1  | 63,577,928           | 63,572,322                         | 5,606      | 0.0088%         |
| 01_Ad_L1_2  | 63,577,928           | 63,572,322                         | 5,606      | 0.0088%         |
| 02_Pm_L1_1  | 55,269,387           | 55,259,617                         | 9,770      | 0.0177%         |
| 02_Pm_L1_2  | 55,269,387           | 55,259,617                         | 9,770      | 0.0177%         |
| 03_Zt_L1_1  | 49,128,635           | 49,126,313                         | 2,322      | 0.0047%         |
| 03_Zt_L1_2  | 49,128,635           | 49,126,313                         | 2,322      | 0.0047%         |
| 04_Asp_L1_1 | 41,800,974           | 41,798,830                         | 2,144      | 0.0051%         |
| 04_Asp_L1_2 | 41,800,974           | 41,798,830                         | 2,144      | 0.0051%         |
| 05_G_L1_1   | 40,906,340           | 40,895,705                         | 10,635     | 0.0260%         |
| 05_G_L1_2   | 40,906,340           | 40,895,705                         | 10,635     | 0.0260%         |
| 06_Ol_L1_1  | 56,417,880           | 56,412,659                         | 5,221      | 0.0093%         |
| 06_Ol_L1_2  | 56,417,880           | 56,412,659                         | 5,221      | 0.0093%         |
| 07_Pr_L1_1  | 47,070,196           | 47,066,726                         | 3,470      | 0.0074%         |
| 07_Pr_L1_2  | 47,070,196           | 47,066,726                         | 3,470      | 0.0074%         |
| 08_Ar_L1_1  | 43,759,260           | 43,705,177                         | 54,083     | 0.1236%         |
| 08_Ar_L1_2  | 43,759,260           | 43,705,177                         | 54,083     | 0.1236%         |

The rRNA contamination


## Use bowtie2 on Arb Silva Coleoptera database

Search for taxonomy Coleoptera. Download fasta without gaps for the LSU and SSU hits (https://www.arb-silva.de/download)

+ 13765 SSU
+ 8776 LSU

```{bash, eval=FALSE}

cd /usr/users/zicola/genomes/rRNA_coleoptera

tar -zxvf arb-silva.de_2025-01-16_id1372508.tgz
mv arb-silva.de_2025-01-16_id1372508_tax_silva_trunc.fasta LSU_coleoptera.fasta

tar -zxvf arb-silva.de_2025-01-16_id1372510.tgz
mv arb-silva.de_2025-01-16_id1372510_tax_silva_trunc.fasta SSU_coleoptera.fasta

seqkit stats *fasta
processed files:  2 / 2 [======================================] ETA: 0s. done
file                  format  type  num_seqs     sum_len  min_len  avg_len  max_len
LSU_coleoptera.fasta  FASTA   RNA      8,776   5,944,179      300    677.3    3,238
SSU_coleoptera.fasta  FASTA   RNA     13,768  16,277,792      300  1,182.3    2,645

cat LSU_coleoptera.fasta SSU_coleoptera.fasta > LSU_SSU_coleoptera.fasta

seqkit stats LSU_SSU_coleoptera.fasta
file                      format  type  num_seqs     sum_len  min_len  avg_len  max_len
LSU_SSU_coleoptera.fasta  FASTA   RNA     22,544  22,221,971      300    985.7    3,238


bowtie2-build -f LSU_SSU_coleoptera.fasta LSU_SSU_coleoptera

```

Test with Ad

```{bash, eval=FALSE}
#!/bin/bash
#SBATCH --job-name=rRNA_removal
#SBATCH --nodes=1
#SBATCH --time=02:00:00
#SBATCH --cpus-per-task=10
#SBATCH --mem=50gb
#SBATCH --partition=medium
#SBATCH --mail-type=END
#SBATCH --mail-user=johan.zicola@uni-goettingen.de
#SBATCH --output=rRNA_removal.%J.out
#SBATCH --error=rRNA_removal.%J.err

mem_gbytes=$(( $SLURM_MEM_PER_NODE/1024 ))

echo "Starting at $(date)"
echo "Job submitted to the ${SLURM_JOB_PARTITION} partition, the default partition on ${SLURM_CLUSTER_NAME}"
echo "Job name: ${SLURM_JOB_NAME}, Job ID: ${SLURM_JOB_ID}"
echo "I have ${SLURM_CPUS_ON_NODE} CPUs and ${mem_gbytes}GiB of RAM on compute node $(hostname)"

module purge
module load rev/21.12 bowtie2

bowtie2 -p 10 --nofw --quiet --very-sensitive-local --phred33  \
    -x /usr/users/zicola/genomes/rRNA_coleoptera/LSU_SSU_coleoptera \
    -1 trimmed_reads/01_Ad_L1_1.trimmed.fq  \
    -2 trimmed_reads/01_Ad_L1_2.trimmed.fq \
    --un-conc rRNA_removed/01_Ad.rRNA_removed_silva_coleptera.fq > /dev/null

```

```{bash, eval=FALSE}
sbatch test_bowtie2.sh
```


```{bash, eval=FALSE}
# Bowtie using smr database

seqkit stats 01_Ad*
processed files:  2 / 2 [======================================] ETA: 0s. done
file                     format  type    num_seqs        sum_len  min_len  avg_len  max_len
01_Ad.rRNA_removed.1.fq  FASTQ   DNA   63,572,322  6,341,800,686        1     99.8      100
01_Ad.rRNA_removed.2.fq  FASTQ   DNA   63,572,322  6,340,036,951        1     99.7      100


# Bowtie using ARB Silva Coleoptera SSU + LSU database

seqkit stats 01_Ad.rRNA_removed_silva_coleptera.*
processed files:  2 / 2 [======================================] ETA: 0s. done
file                                     format  type    num_seqs        sum_len  min_len  avg_len  max_len
01_Ad.rRNA_removed_silva_coleptera.1.fq  FASTQ   DNA   63,574,766  6,342,041,048        1     99.8      100
01_Ad.rRNA_removed_silva_coleptera.2.fq  FASTQ   DNA   63,574,766  6,340,277,246        1     99.7      100


```

63574766-63572322 = 2444. I can actually less reads mapping rRNA when using the Silva database for Coleptera (22,544 sequences) compared to the smr default database (73,977 sequences).



## test sortmerna

```{bash, eval=FALSE}

# Take first entry of smr_v4.3_default_db.fasta and put it in seq_rRNA.fa


cd /usr/users/zicola/genomes/sortmerna_database/test

python ~/git_repositories/fastq_generator/fastq_generator.py generate_mapped_fastq_PE seq_rRNA.fa 50 10 10 > fastq_interleaved.fq

# Split in two files
seqkit stats fastq_interleaved.fq
file                  format  type  num_seqs  sum_len  min_len  avg_len  max_len
fastq_interleaved.fq  FASTQ   DNA        520   26,000       50       50       50

head -n 1040 fastq_interleaved.fq > fastq_interleaved.R1.fq
tail -n 1040 fastq_interleaved.fq > fastq_interleaved.R2.fq

module load rev/21.12 sortmerna/20170809

# Get them merged (each pair is together)
/usr/product/bioinfo/SL_7.0/BIOINFORMATICS/SORTMERNA/sortmerna_20170809/scripts/merge-paired-reads.sh fastq_interleaved.R1.fq fastq_interleaved.R2.fq  fastq_interleaved.merged.fq

sortmerna -a 1 --paired_in --fastx \
    --ref ../smr_v4.3_default_db.fasta,../smr_v4.3_default_db.idx \
    --reads fastq_interleaved.merged.fq \
    --aligned aligned \
    --other mRNA -v
# Nothing aligned

sortmerna -a 1 --fastx \
    --ref ../smr_v4.3_default_db.fasta,../smr_v4.3_default_db.idx \
    --reads fastq_interleaved.R1.fq --reads fastq_interleaved.R2.fq \
    --aligned aligned \
    --other mRNA -v




sortmerna -a 1 --fastx \
    --ref ../smr_v4.3_default_db.fasta,../smr_v4.3_default_db.idx \
    --reads fastq_interleaved.merged.fq \
    --aligned aligned \
    --other mRNA -v


sortmerna -a 1 --fastx \
    --ref ../smr_v4.3_default_db.fasta,../smr_v4.3_default_db.idx \
    --reads fastq_interleaved.R1.fq \
    --aligned aligned \
    --other mRNA -v
    
    
sortmerna -a 1 --fastx \
    --ref ../smr_v4.3_default_db.fasta,../smr_v4.3_default_db.idx \
    --reads tests/data/set2_environmental_study_550_amplicon.fasta \
    --aligned aligned \
    --other mRNA -v
    
sortmerna -a 1 --fastx \
    --ref ../smr_v4.3_default_db.fasta,../smr_v4.3_default_db.idx \
    --reads tests/data/set2_environmental_study_550_amplicon.fasta \
    --aligned aligned \
    --other mRNA -v
    
```

Not working

```{bash, eval=FALSE}
module load rev/21.12 sortmerna/20170809

indexdb --ref seq_rRNA.fa,seq_rRNA.idx -v

sortmerna -a 1 --fastx \
    --ref seq_rRNA.fa,seq_rRNA.idx \
    --reads fastq_interleaved.R1.fq \
    --aligned aligned \
    --other mRNA -v

# Now it worked! Everything is in aligned.fq

sortmerna -a 1 --fastx \
    --ref smr_v4.3_default_db.fasta,smr_v4.3_default_db.idx \
    --reads test/fastq_interleaved.R1.fq \
    --aligned aligned \
    --other mRNA -v

# Not working
```

I think the DB is corrupted.

Remake the indexes of smr_v4.3_default_db.fasta

```{bash, eval=FALSE}

srun --pty -t 00:50:00 --mem=20G -p medium bash

module load rev/21.12 sortmerna/20170809

indexdb --ref smr_v4.3_default_db.fasta,smr_v4.3_default_db.idx -v

```

```{bash, eval=FALSE}
# Single end mode
sortmerna -a 1 --fastx \
    --ref smr_v4.3_default_db.fasta,smr_v4.3_default_db.idx \
    --reads test/fastq_interleaved.R1.fq \
    --aligned aligned \
    --other mRNA -v
```

Now it works!

```{bash, eval=FALSE}
# Paired end mode
sortmerna -a 1 --paired_in --fastx \
    --ref ../smr_v4.3_default_db.fasta,../smr_v4.3_default_db.idx \
    --reads fastq_interleaved.merged.fq \
    --aligned aligned \
    --other mRNA -v
```

Works

Redo a test to see the fraction of reads removed.


## Test Anchomenus dorsalis 


```{bash, eval=FALSE}
srun --pty -p medium --mem=40G --time=04:00:00  bash

module load rev/21.12 sortmerna/20170809

cd /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trimmed_reads
mkdir test_sortmerna
gunzip  01_Ad_*

# Get them merged (each pair is together)
/usr/product/bioinfo/SL_7.0/BIOINFORMATICS/SORTMERNA/sortmerna_20170809/scripts/merge-paired-reads.sh 01_Ad_L1_1.trimmed.fq 01_Ad_L1_2.trimmed.fq  test_sortmerna/01_Ad_L1.merged.fq

cd test_sortmerna

DB_sortmerna="/usr/users/zicola/genomes/sortmerna_database"

sortmerna -a 1 --paired_in --fastx \
    --ref ${DB_sortmerna}/smr_v4.3_default_db.fasta,${DB_sortmerna}/smr_v4.3_default_db.idx \
    --reads 01_Ad_L1.merged.fq \
    --aligned 01_Ad_L1_aligned \
    --other 01_Ad_L1_mRNA -v

# Many messages showing that read length is smaller than 18 nt
# WARNING: At least one of the reads is shorter than 18 nucleotides, by default it will not be searched

# Reseparate strands

```

Already running for 3 hours and still not done. Job was killed because overtiming the 4 hours!


I think it is slowed down by the presence of reads smaller than 18 bp. Remove these and try again


```{bash, eval=FALSE}

cd /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trimmed_reads

module load cutadapt/4.4

srun -p medium --time=00:40:00 --mem=20G --cpus-per-task=5 cutadapt --cores 5 --quality-base 33 --max-n 0 --minimum-length 25 -o test_sortmerna/01_Ad_L1_1.trimmed.min25.fq -p test_sortmerna/01_Ad_L1_2.trimmed.min25.fq 01_Ad_L1_1.trimmed.fq 01_Ad_L1_2.trimmed.fq


Processing paired-end reads on 5 cores ...
Finished in 601.723 s (9.464 µs/read; 6.34 M reads/minute).

=== Summary ===

Total read pairs processed:         63,577,851

== Read fate breakdown ==
Pairs that were too short:                 267 (0.0%)
Pairs with too many N:                       0 (0.0%)
Pairs written (passing filters):    63,577,584 (100.0%)

Total basepairs processed: 12,682,916,871 bp
  Read 1: 6,342,339,512 bp
  Read 2: 6,340,577,359 bp
Total written (filtered):  12,682,907,729 bp (100.0%)
  Read 1: 6,342,334,981 bp
  Read 2: 6,340,572,748 bp


srun -p medium --time=00:60:00 --mem=40G /usr/product/bioinfo/SL_7.0/BIOINFORMATICS/SORTMERNA/sortmerna_20170809/scripts/merge-paired-reads.sh 01_Ad_L1_1.trimmed.min25.fq 01_Ad_L1_2.trimmed.min25.fq  01_Ad_L1.merged.fq

```


```{bash, eval=FALSE}

#!/bin/bash
#SBATCH --job-name=rRNA_removal
#SBATCH --time=10:00:00
#SBATCH --cpus-per-task=1
#SBATCH --mem=40gb
#SBATCH --partition=medium
#SBATCH --mail-type=END
#SBATCH --mail-user=johan.zicola@uni-goettingen.de
#SBATCH --output=rRNA_removal.%J.out
#SBATCH --error=rRNA_removal.%J.err

module load rev/21.12 sortmerna/20170809

DB_sortmerna="/usr/users/zicola/genomes/sortmerna_database"

sortmerna -a 1 --paired_in --fastx \
    --ref ${DB_sortmerna}/smr_v4.3_default_db.fasta,${DB_sortmerna}/smr_v4.3_default_db.idx \
    --reads 01_Ad_L1.merged.fq \
    --aligned 01_Ad_L1_aligned \
    --other 01_Ad_L1_mRNA -v

```


```{bash, eval=FALSE}
sbatch run_sortmerna.sh
```

Done after 08:26 hours and used 38.31 Gb of RAM so I was very close to the limit set. The input fastq was 32 Gb in size

```{bash, eval=FALSE}
conda activate py3.7

srun -p medium --time=00:20:00 --mem=10G seqkit stats 01_Ad_L1_mRNA.fq 01_Ad_L1_aligned.fq

processed files:  2 / 2 [] ETA: 0s. done
file                 format  type     num_seqs         sum_len  min_len  avg_len  max_len
01_Ad_L1_mRNA.fq     FASTQ   DNA   120,612,416  12,029,946,265       25     99.7      100
01_Ad_L1_aligned.fq  FASTQ   DNA     6,542,752     652,961,464       35     99.8      100


```

6542752/127155168
[1] 0.05145486

> 120612416/2
[1] 60306208

60306208 pairs remain after filtering

With sortmerna, I get about 5% of the reads that map rRNA while I found only 0.0088% using the bowtie approach.

Unmerge reads

```{bash, eval=FALSE}

module load rev/21.12 sortmerna/20170809

srun -p medium --time=00:40:00 --mem=40G --output=unmerge.log bash /usr/product/bioinfo/SL_7.0/BIOINFORMATICS/SORTMERNA/sortmerna_20170809/scripts/unmerge-paired-reads.sh 01_Ad_L1_mRNA.fq 01_Ad_L1_mRNA.R1.fq 01_Ad_L1_mRNA.R2.fq &

```

16 minutes needed.




## Ribodetector

sortmerna is very long, try the newly developped ribodetector (https://github.com/hzi-bifo/RiboDetector)

Paper: https://academic.oup.com/nar/article/50/10/e60/6533611


> To demonstrate that rRNA removal is the most time-consuming step in processing with the most widely used method (SortMeRNA), we ran the entire data processing workflow on the oral metatranscriptome dataset that was used in the previous benchmark with 40 CPU cores. We performed read quality control with Fastp (46), then removed rRNA reads with SortMeRNA, filtered out human reads by mapping the non-rRNA reads against the human reference with BWA, and finally mapped the cleaned reads against HOMD reference CDSs with BWA. The read quality control required 1 min, rRNA removal with SortMeRNA took 109 min, human reads removal took 4 min, and read alignment took 2 min of computing time. In comparison, with RiboDetector, rRNA removal required only 4 min in GPU mode and 15 min in CPU mode.

It uses GPU (see GWDG ressource https://docs.hpc.gwdg.de/how_to_use/compute_partitions/gpu_partitions/index.html)


```{bash, eval=FALSE}
# Check gpus available
sinfo -o "%25N  %5c  %10m  %32f  %10G %18P " | grep gpu | grep scc

```


```{bash, eval=FALSE}


module load rev/23.12 miniconda3/22.11.1

conda create -n ribodetector python=3.8
conda activate ribodetector

# conda install -c bioconda ribodetector
# Doesn't work

pip install ribodetector
# Worked

ribodetector --version
ribodetector 0.3.1

# Doc recommends 20 Gb of RAM for PE data and 20 threads

```


### CPU version


```{bash, eval=FALSE}
#!/bin/bash
#SBATCH --job-name=ribodetector_cpu
#SBATCH --nodes=1
#SBATCH --partition=medium
#SBATCH --time=06:00:00
#SBATCH --cpus-per-task=20
#SBATCH --threads-per-core=1
#SBATCH --mem=40gb
#SBATCH --mail-type=END
#SBATCH --mail-user=johan.zicola@uni-goettingen.de
#SBATCH --output=job-%x-%j.out
#SBATCH --error=job-%x-%j.err

module load anaconda3/2021.05

source activate ribodetector

ribodetector_cpu -t 20 \
  -l 100 \
  -i 01_Ad_L1_1.trimmed.min25.fq 01_Ad_L1_2.trimmed.min25.fq \
  -e rrna \
  --chunk_size 256 \
  -o reads.nonrrna.1.fq reads.nonrrna.2.fq

```

```{bash, eval=FALSE}
2025-01-30 15:54:46 : INFO  Processed 63577584 sequences in total
2025-01-30 15:54:46 : INFO  Detected 62903350 non-rRNA sequences
2025-01-30 15:54:46 : INFO  Detected 674234 rRNA sequences

```

```{bash, eval=FALSE}
sa 5518685
JobID           JobName  AllocCPUS     ReqMem     MaxRSS  Timelimit    Elapsed
------------ ---------- ---------- ---------- ---------- ---------- ----------
5518685      ribodetec+         20        40G              06:00:00   01:43:54
5518685.bat+      batch         20                22.39G              01:43:54
5518685.ext+     extern         20                     0              01:43:54

```

Took 22 GB and 1 hour 43min (compared to 08:26 hours and used 38.31 Gb of RAM for sortmerna (excluding gunzipping, merging, and unmerging).


Let's compare number of reads sorted as rRNA with sortmerna vs ribodetector

```{bash, eval=FALSE}


# Sort me rRNA
srun -p medium --time=00:20:00 --mem=10G seqkit stats 01_Ad_L1_mRNA.fq 01_Ad_L1_aligned.fq

processed files:  2 / 2 [] ETA: 0s. done
file                 format  type     num_seqs         sum_len  min_len  avg_len  max_len
01_Ad_L1_mRNA.fq     FASTQ   DNA   120,612,416  12,029,946,265       25     99.7      100
01_Ad_L1_aligned.fq  FASTQ   DNA     6,542,752     652,961,464       35     99.8      100


srun -p medium --time=00:10:00 --mem=10G seqkit stats reads.nonrrna.1.fq
reads.nonrrna.1.fq  FASTQ   DNA   62,903,350  6,275,070,014       25     99.8      100

```

Total number of reads: 63577584 (63.5M)

Sortmerna

> 120612416/2
[1] 60306208

60,306,208 => 60M reads retrieved as not rRNA.

ribodetector

62,903,350 => 62.9M reads retrieved as not rRNA.

Sortmerna found 2.6M extra reads mapping rRNAs.


```{bash, eval=FALSE}
#!/bin/bash
#SBATCH --job-name=ribodetector_cpu
#SBATCH --nodes=1
#SBATCH --partition=medium
#SBATCH --time=06:00:00
#SBATCH --cpus-per-task=20
#SBATCH --threads-per-core=1
#SBATCH --mem=40gb
#SBATCH --mail-type=END
#SBATCH --mail-user=johan.zicola@uni-goettingen.de
#SBATCH --output=job-%x-%j.out
#SBATCH --error=job-%x-%j.err

module load anaconda3/2021.05

source activate ribodetector

ribodetector_cpu -t 20 \
  -l 100 \
  -i $READ1 $READ2 \
  -e rrna \
  --chunk_size 256 \
  -o $OUTPUT_READ1 $OUTPUT_READ2

```


```{bash, eval=FALSE}
cd /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trimmed_reads

sbatch sbatch_ribodetector.sh --export=READ1=02_Pm_L1_1.trimmed.fq,READ2=02_Pm_L1_1.trimmed.fq,OUTPUT_READ1=rrna_removed/02_Pm_L1_1.trimmed.rm_rrna.fq,OUTPUT_READ2=rrna_removed/02_Pm_L1_2.trimmed.rm_rrna.fq

```

### GPU version

From the SCC, I need to use the scc-gpu partion, which has access to NVIDIA A100 Tensor Core GPU (https://www.nvidia.com/en-us/data-center/a100/)


```{bash, eval=FALSE}

sinfo -o "%25N  %5c  %10m  %32f  %10G %18P " | grep gpu | grep scc-gpu
ggpu150                    128    515810      ssd,inet,wlcg,cpu_fixed,80gb,zen  gpu:A100:4 scc-gpu
ggpu[135-136]              128    515810      ssd,inet,wlcg,cpu_fixed,zen3,mil  gpu:A100:4 scc-gpu
ggpu[138-149,151-158,211-  64+    515810+     ssd,inet,wlcg,cpu_fixed,80gb,zen  gpu:A100:4 scc-gpu
ggpu[194-199]              128    515810+     ssd,inet,80gb,zen3,milan          gpu:A100:4 scc-gpu
ggpu237                    192    1031000     ssd,inet,96gb,sapphirerapids,vas  gpu:H100:4 scc-gpu

```

```{bash, eval=FALSE}
srun --time=00:10:00 --pty -p gpu /bin/bash
# Works

zicola@agq008
```

agq nodes (https://docs.hpc.gwdg.de/how_to_use/spack/index.html)

```{bash, eval=FALSE}
# https://chat.uni-goettingen.de/#/room/#hpc-users:gwdg.de/$S3hbla6bqoWuLfoIghwkRoEO1MrVsJtrryGg-4modr0

sinfo -p gpu --format=%N,%G NODELIST,GRES agq[002-014]
NODELIST,GRES
agq[007-012],gpu:RTX5000:4(S:0-1)
agq[013-014],gpu:RTX5000:4(S:0-1),shard:RTX5000:8(S:0-1)

```

Nodes	CPU and OS Architecture/Target	Fabric	cuda_arch	Partitions
agqXXX, agtXXX	linux-scientific7-cascadelake	OPA	70	gpu

https://gwdg.de/hpc/systems/scc/

CPU: 2 x Xeon Gold 6242
GPU: 4 x Quadro RTX5000 (16 Gb VRAM, 3072 FP32 cores, 384 tensor cores, 75 compute capacity)

agq node has 192 Gb of GPU RAM


```{bash, eval=FALSE}
srun --time=00:10:00 --pty -p gpu -G 1 /bin/bash

# logged on node agq009

echo $CUDA_VISIBLE_DEVICES
0

nvidia-smi
Sat Feb  1 13:55:20 2025
+-----------------------------------------------------------------------------------------+
| NVIDIA-SMI 550.144.03             Driver Version: 550.144.03     CUDA Version: 12.4     |
|-----------------------------------------+------------------------+----------------------+
| GPU  Name                 Persistence-M | Bus-Id          Disp.A | Volatile Uncorr. ECC |
| Fan  Temp   Perf          Pwr:Usage/Cap |           Memory-Usage | GPU-Util  Compute M. |
|                                         |                        |               MIG M. |
|=========================================+========================+======================|
|   0  Quadro RTX 5000                On  |   00000000:86:00.0 Off |                  Off |
| 34%   29C    P8              8W /  230W |       1MiB /  16384MiB |      0%      Default |
|                                         |                        |                  N/A |
+-----------------------------------------+------------------------+----------------------+

+-----------------------------------------------------------------------------------------+
| Processes:                                                                              |
|  GPU   GI   CI        PID   Type   Process name                              GPU Memory |
|        ID   ID                                                               Usage      |
|=========================================================================================|
|  No running processes found                                                             |
+-----------------------------------------------------------------------------------------+

```

The RAM of the GPU is 16 Gb (16384 Mb)


Use default parameter of ribodetect

```{bash, eval=FALSE}
#!/bin/bash
#SBATCH --job-name=ribodetector_gpu
#SBATCH -t 02:00:00
#SBATCH --threads-per-core=1
#SBATCH -p gpu
#SBATCG -G 1
#SBATCH --mem=15G
#SBATCH --cpus-per-task=10
#SBATCH -o job-%x-%j.out
#SBATCH -e job-%x-%j.err

module load anaconda3/2021.05

source activate ribodetector

# Load cuda, otherwise, error message  ERROR  No visible CUDA devices! from ribodetector
module load cuda/12.1.1

ribodetector --threads 10 \
  --len 100 \
  --input 01_Ad_L1_1.trimmed.min25.fq 01_Ad_L1_2.trimmed.min25.fq \
  --memory 12 \
  --ensure rrna \
  --output reads.gpu.nonrrna.1.fq reads.gpu.nonrrna.2.fq && date

```



```{bash, eval=FALSE}
#!/bin/bash
#SBATCH --job-name=ribodetector_gpu
#SBATCH -t 02:00:00
#SBATCH --threads-per-core=1
#SBATCH -p gpu
#SBATCH -G 1
#SBATCH --gres=gpu:1
#SBATCH --mem=15G
#SBATCH --cpus-per-task=10
#SBATCH -o job-%x-%j.out
#SBATCH -e job-%x-%j.err

mem_gbytes=$(( $SLURM_MEM_PER_NODE/1024 ))

echo "Starting at $(date)"
echo "Job submitted to the ${SLURM_JOB_PARTITION} partition, the default partition on ${SLURM_CLUSTER_NAME}"
echo "Job name: ${SLURM_JOB_NAME}, Job ID: ${SLURM_JOB_ID}"
echo "  I have ${SLURM_CPUS_ON_NODE} CPUs and ${mem_gbytes}GiB of RAM on compute node $(hostname)"
echo "Submitting job with sbatch from directory: ${SLURM_SUBMIT_DIR}"

module load anaconda3/2021.05

source activate ribodetector

# Load cuda, otherwise, error message  ERROR  No visible CUDA devices! from ribodetector
module load cuda/12.1.1

ribodetector --threads 10 \
  --len 100 \
  --input 01_Ad_L1_1.trimmed.min25.fq 01_Ad_L1_2.trimmed.min25.fq \
  --memory 12 \
  --ensure rrna \
  --output reads.gpu.nonrrna.1.fq reads.gpu.nonrrna.2.fq

```

```{bash, eval=FALSE}
$ cat job-ribodetector_gpu-5638529.err
2025-02-04 21:16:35 : INFO  Using high MCC model
2025-02-04 21:16:35 : INFO  Log file: None
2025-02-04 21:16:50 : INFO  Model using cuda for read length 100 loaded
2025-02-04 21:16:53 : INFO  Choose batch size: 8192 based on the given GPU RAM size 12GB and max read length 100
slurmstepd: error: *** JOB 5638529 ON agq013 CANCELLED AT 2025-02-04T23:16:09 DUE TO TIME LIMIT ***
slurmstepd: error: Detected 2 oom_kill events in StepId=5638529.batch. Some of the step tasks have been OOM Killed.

sa 5638529
JobID           JobName  AllocCPUS     ReqMem     MaxRSS  Timelimit    Elapsed      State
------------ ---------- ---------- ---------- ---------- ---------- ---------- ----------
5638529      ribodetec+         10        15G              02:00:00   02:00:02    TIMEOUT
5638529.bat+      batch         10                15.00G              02:00:04 OUT_OF_ME+
5638529.ext+     extern         10                 0.00G              02:00:04  COMPLETED


```


Add extra RAM (30Gb) and add --mem-per-gpu=15G

`sbatch: fatal: --mem, --mem-per-cpu, and --mem-per-gpu are mutually exclusive.`


```{bash, eval=FALSE}
#!/bin/bash
#SBATCH --job-name=ribodetector_gpu
#SBATCH --threads-per-core=1
#SBATCH -p gpu
#SBATCH -G 1
#SBATCH --gres=gpu:1
#SBATCH --mem=30G
#SBATCH --cpus-per-task=10
#SBATCH -o job-%x-%j.out
#SBATCH -e job-%x-%j.err

mem_gbytes=$(( $SLURM_MEM_PER_NODE/1024 ))

echo "Starting at $(date)"
echo "Job submitted to the ${SLURM_JOB_PARTITION} partition, the default partition on ${SLURM_CLUSTER_NAME}"
echo "Job name: ${SLURM_JOB_NAME}, Job ID: ${SLURM_JOB_ID}"
echo "  I have ${SLURM_CPUS_ON_NODE} CPUs and ${mem_gbytes}GiB of RAM on compute node $(hostname)"
echo "Submitting job with sbatch from directory: ${SLURM_SUBMIT_DIR}"

module load anaconda3/2021.05

source activate ribodetector

# Load cuda, otherwise, error message  ERROR  No visible CUDA devices! from ribodetector
module load cuda/12.1.1

ribodetector --threads 10 \
  --len 100 \
  --input 01_Ad_L1_1.trimmed.min25.fq 01_Ad_L1_2.trimmed.min25.fq \
  --memory 12 \
  --ensure rrna \
  --output reads.gpu.nonrrna.1.fq reads.gpu.nonrrna.2.fq
```


Still not done after 4 hours. Kill the job.



#### Interactive run

```{bash, eval=FALSE}
srun --time=02:00:00 --mem=15G --cpus-per-task=10 --threads-per-core=1 --pty -p gpu -G 1 /bin/bash

module load anaconda3/2021.05

source activate ribodetector

# Load cuda, otherwise, error message  ERROR  No visible CUDA devices! from ribodetector
module load cuda/12.1.1

ribodetector --threads 10 \
  --len 100 \
  --input 01_Ad_L1_1.trimmed.min25.fq 01_Ad_L1_2.trimmed.min25.fq \
  --memory 12 \
  --ensure rrna \
  --output reads.gpu.nonrrna.1.fq reads.gpu.nonrrna.2.fq

```

Start run Sat Feb  1 14:01:37 CET 2025

The screen was killed and there were not output files.



# Trinity w/o rRNA

```{bash, eval=FALSE}
sbatch  --export=INPUT_FASTQ_LEFT=trimmed_reads/test_sortmerna/reads.nonrrna.1.fq,INPUT_FASTQ_RIGHT=trimmed_reads/test_sortmerna/reads.nonrrna.2.fq,OUTPUT_DIR=trinity_output/01_Ad_wo_rRNA/01_Ad_trinity 02_trinity.sh

```

```{bash, eval=FALSE}
============ Job Information ==========================
JobID: 5675836
JobName: Trinity
Submitted: 2025-02-06T16:48:17
Started: 2025-02-06T17:17:32
Ended: 2025-02-07T02:50:02
Elapsed: 573 min, Limit: 720 min, Difference: 147 min
CPUs: 8, Nodes: 1
Estimated consumption: 152.80 core-hours
=======================================================

```


```{bash, eval=FALSE}
# Without rRNA
seqkit stats 01_Ad_trinity.Trinity.fasta
file                         format  type  num_seqs      sum_len  min_len  avg_len  max_len
01_Ad_trinity.Trinity.fasta  FASTA   DNA    102,969  110,492,695      185  1,073.1   24,547

cut -f1 01_Ad_trinity.Trinity.fasta.gene_trans_map |sort |uniq |wc -l
66955


# With rRNA
seqkit stats 01_Ad_trinity.Trinity.fasta
file                         format  type  num_seqs      sum_len  min_len  avg_len  max_len
01_Ad_trinity.Trinity.fasta  FASTA   DNA    103,137  110,493,798      184  1,071.3   33,453

cut -f1 01_Ad_trinity.Trinity.fasta.gene_trans_map |sort |uniq |wc -l
67047

```

168 extra transcripts and 92 extra genes when rRNAs are not removed.


# Trinity with rRNA

## 01_Ad

```{bash, eval=FALSE}
============ Job Information ==========================
JobID: 4909774
JobName: Trinity
Submitted: 2025-01-09T11:00:42
Started: 2025-01-09T11:01:32
Ended: 2025-01-09T20:00:00
Elapsed: 539 min, Limit: 2880 min, Difference: 2341 min
CPUs: 8, Nodes: 1
Estimated consumption: 143.73 core-hours
=======================================================
```


## 02_Pm

```{bash, eval=FALSE}
============ Job Information ==========================
JobID: 5037510
JobName: Trinity
Submitted: 2025-01-15T18:04:04
Started: 2025-01-15T18:30:58
Ended: 2025-01-16T05:29:00
Elapsed: 659 min, Limit: 720 min, Difference: 61 min
CPUs: 8, Nodes: 1
Estimated consumption: 175.73 core-hours
=======================================================

```

## 03_Zt

```{bash, eval=FALSE}
============ Job Information ==========================
JobID: 5037511
JobName: Trinity
Submitted: 2025-01-15T18:04:04
Started: 2025-01-15T18:33:29
Ended: 2025-01-16T02:21:09
Elapsed: 468 min, Limit: 720 min, Difference: 252 min
CPUs: 8, Nodes: 1
Estimated consumption: 124.80 core-hours
=======================================================
```


## 04_Asp

```{bash, eval=FALSE}
============ Job Information ==========================
JobID: 5037512
JobName: Trinity
Submitted: 2025-01-15T18:04:04
Started: 2025-01-15T18:43:26
Ended: 2025-01-16T01:05:36
Elapsed: 383 min, Limit: 720 min, Difference: 337 min
CPUs: 8, Nodes: 1
Estimated consumption: 102.13 core-hours
=======================================================

```


## 05_G

```{bash, eval=FALSE}
============ Job Information ==========================
JobID: 5037513
JobName: Trinity
Submitted: 2025-01-15T18:04:04
Started: 2025-01-16T01:05:50
Ended: 2025-01-16T09:00:03
Elapsed: 475 min, Limit: 720 min, Difference: 245 min
CPUs: 8, Nodes: 1
Estimated consumption: 126.67 core-hours
=======================================================
```


## 06_Ol

```{bash, eval=FALSE}
============ Job Information ==========================
JobID: 5037514
JobName: Trinity
Submitted: 2025-01-15T18:04:04
Started: 2025-01-16T09:00:47
Ended: 2025-01-16T16:55:44
Elapsed: 475 min, Limit: 720 min, Difference: 245 min
CPUs: 8, Nodes: 1
Estimated consumption: 126.67 core-hours
=======================================================
```

## 07_Pr

```{bash, eval=FALSE}

============ Job Information ==========================
JobID: 5037515
JobName: Trinity
Submitted: 2025-01-15T18:04:04
Started: 2025-01-16T09:51:54
Ended: 2025-01-16T17:41:27
Elapsed: 470 min, Limit: 720 min, Difference: 250 min
CPUs: 8, Nodes: 1
Estimated consumption: 125.33 core-hours
=======================================================
```


## 08_Ar

Time out

```{bash, eval=FALSE}
============ Job Information ==========================
JobID: 5037516
JobName: Trinity
Submitted: 2025-01-15T18:04:05
Started: 2025-01-16T09:51:54
Ended: 2025-01-16T21:52:17
Elapsed: 721 min, Limit: 720 min, Difference: -1 min
CPUs: 8, Nodes: 1
Estimated consumption: 192.27 core-hours
=======================================================
```

Based on the forum, I can rerun Trinity command and it will start from the last completed checkpoint


```{bash, eval=FALSE}
i="08_Ar"
sbatch  --export=INPUT_FASTQ_LEFT=trimmed_reads/${i}_L1_1.trimmed.fq,INPUT_FASTQ_RIGHT=trimmed_reads/${i}_L1_2.trimmed.fq,OUTPUT_DIR=trinity_output/${i}/${i}_trinity trinity_run.sh
```


```{bash, eval=FALSE}
============ Job Information ==========================
JobID: 5056408
JobName: Trinity
Submitted: 2025-01-17T08:47:18
Started: 2025-01-17T08:47:24
Ended: 2025-01-17T11:05:41
Elapsed: 139 min, Limit: 720 min, Difference: 581 min
CPUs: 8, Nodes: 1
Estimated consumption: 37.07 core-hours
=======================================================
```

Finished in 139 min for the remaining steps.



# Quality check TrinityStats

```{bash, eval=FALSE}

path_output="/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinity_output"

module load apptainer

apptainer exec --bind $path_output -e /usr/users/zicola/bin/trinityrnaseq.v2.15.1.simg /usr/local/bin/util/TrinityStats.pl 01_Ad/01_Ad_trinity.Trinity.fasta

```

## Amara_similata

```
################################
## Counts of transcripts, etc.
################################
Total trinity 'genes':  45431
Total trinity transcripts:   72481
Percent GC: 37.03

########################################
Stats based on ALL transcript contigs:
########################################

   Contig N10: 6191
   Contig N20: 4539
   Contig N30: 3567
   Contig N40: 2817
   Contig N50: 2233

   Median contig length: 533
   Average contig: 1139.59
   Total assembled bases: 82598703


#####################################################
## Stats based on ONLY LONGEST ISOFORM per 'GENE':
#####################################################

   Contig N10: 5577
   Contig N20: 3920
   Contig N30: 2963
   Contig N40: 2276
   Contig N50: 1750

   Median contig length: 390
   Average contig: 862.61
   Total assembled bases: 39189450

```

## Anchomenus_dorsalis

```
################################
## Counts of transcripts, etc.
################################
Total trinity 'genes':  67047
Total trinity transcripts:   103137
Percent GC: 36.52

########################################
Stats based on ALL transcript contigs:
########################################

   Contig N10: 6347
   Contig N20: 4615
   Contig N30: 3602
   Contig N40: 2835
   Contig N50: 2225

   Median contig length: 460
   Average contig: 1071.33
   Total assembled bases: 110493798


#####################################################
## Stats based on ONLY LONGEST ISOFORM per 'GENE':
#####################################################

   Contig N10: 5485
   Contig N20: 3842
   Contig N30: 2821
   Contig N40: 2091
   Contig N50: 1507

   Median contig length: 348
   Average contig: 751.81
   Total assembled bases: 50406345

```

## Aphidius_rhopalosiphi

```
################################
## Counts of transcripts, etc.
################################
Total trinity 'genes':  202721
Total trinity transcripts:   249888
Percent GC: 27.13

########################################
Stats based on ALL transcript contigs:
########################################

   Contig N10: 3219
   Contig N20: 2170
   Contig N30: 1504
   Contig N40: 986
   Contig N50: 680

   Median contig length: 350
   Average contig: 564.43
   Total assembled bases: 141043361


#####################################################
## Stats based on ONLY LONGEST ISOFORM per 'GENE':
#####################################################

   Contig N10: 2421
   Contig N20: 1436
   Contig N30: 897
   Contig N40: 645
   Contig N50: 507

   Median contig length: 334
   Average contig: 478.33
   Total assembled bases: 96967451
```

## Harpalus_affinis

```
################################
## Counts of transcripts, etc.
################################
Total trinity 'genes':  63019
Total trinity transcripts:   104390
Percent GC: 36.59

########################################
Stats based on ALL transcript contigs:
########################################

   Contig N10: 5600
   Contig N20: 3984
   Contig N30: 3006
   Contig N40: 2372
   Contig N50: 1843

   Median contig length: 453
   Average contig: 961.65
   Total assembled bases: 100387029


#####################################################
## Stats based on ONLY LONGEST ISOFORM per 'GENE':
#####################################################

   Contig N10: 4898
   Contig N20: 3317
   Contig N30: 2445
   Contig N40: 1830
   Contig N50: 1357

   Median contig length: 352
   Average contig: 728.03
   Total assembled bases: 45879553

```

## Harpalus_eous

```
################################
## Counts of transcripts, etc.
################################
Total trinity 'genes':  64282
Total trinity transcripts:   100481
Percent GC: 35.70

########################################
Stats based on ALL transcript contigs:
########################################

   Contig N10: 5486
   Contig N20: 3965
   Contig N30: 3028
   Contig N40: 2375
   Contig N50: 1866

   Median contig length: 447
   Average contig: 959.27
   Total assembled bases: 96388352


#####################################################
## Stats based on ONLY LONGEST ISOFORM per 'GENE':
#####################################################

   Contig N10: 4825
   Contig N20: 3342
   Contig N30: 2471
   Contig N40: 1872
   Contig N50: 1369

   Median contig length: 357
   Average contig: 731.99
   Total assembled bases: 47053619

```


## Poecilus_cupreus

```
################################
## Counts of transcripts, etc.
################################
Total trinity 'genes':  56798
Total trinity transcripts:   92572
Percent GC: 37.40

########################################
Stats based on ALL transcript contigs:
########################################

   Contig N10: 6481
   Contig N20: 4676
   Contig N30: 3736
   Contig N40: 2960
   Contig N50: 2341

   Median contig length: 538
   Average contig: 1169.79
   Total assembled bases: 108290220


#####################################################
## Stats based on ONLY LONGEST ISOFORM per 'GENE':
#####################################################

   Contig N10: 5761
   Contig N20: 4107
   Contig N30: 3008
   Contig N40: 2283
   Contig N50: 1693

   Median contig length: 376
   Average contig: 829.30
   Total assembled bases: 47102779
```

## Pterostichus_melanarius


```
################################
## Counts of transcripts, etc.
################################
Total trinity 'genes':  94235
Total trinity transcripts:   140145
Percent GC: 36.70

########################################
Stats based on ALL transcript contigs:
########################################

   Contig N10: 5401
   Contig N20: 3962
   Contig N30: 2946
   Contig N40: 2241
   Contig N50: 1687

   Median contig length: 372
   Average contig: 829.48
   Total assembled bases: 116247368


#####################################################
## Stats based on ONLY LONGEST ISOFORM per 'GENE':
#####################################################

   Contig N10: 4755
   Contig N20: 3178
   Contig N30: 2225
   Contig N40: 1580
   Contig N50: 1027

   Median contig length: 318
   Average contig: 620.79
   Total assembled bases: 58500213

```

## Zabrus_tenebrioides

```
################################
## Counts of transcripts, etc.
################################
Total trinity 'genes':  51952
Total trinity transcripts:   76934
Percent GC: 36.81

########################################
Stats based on ALL transcript contigs:
########################################

   Contig N10: 6342
   Contig N20: 4654
   Contig N30: 3615
   Contig N40: 2885
   Contig N50: 2285

   Median contig length: 505
   Average contig: 1129.58
   Total assembled bases: 86902887


#####################################################
## Stats based on ONLY LONGEST ISOFORM per 'GENE':
#####################################################

   Contig N10: 5679
   Contig N20: 3983
   Contig N30: 3002
   Contig N40: 2313
   Contig N50: 1753

   Median contig length: 377
   Average contig: 837.57
   Total assembled bases: 43513583

```




# Fungal presence

```{bash, eval=FALSE}

for i in *_Taxonomy_report.tabular; do 
  echo -e "\n======$i"
  awk '$9=="fung:fungi" {print $0}' $i | \
  cut -f6,7 | sort | uniq -c | sort | sed 's/^\s*//' | \
  awk '{print $2,$3,$1}'  | sort -k3nr | head
done

======01_Ad_Taxonomy_report.tabular
Mucor circinelloides 454
Mucor velutinosus 414
Mucor lusitanicus 410
Mucor irregularis 267
Mucor racemosus 118
Pilaira anomala 100
Mucor plumbeus 90
Thamnidium elegans 80
Mucor piriformis 69
Mucor mucedo 66

======02_Pm_Taxonomy_report.tabular
Mucor indicus 225
Mucor circinelloides 82
Gigaspora margarita 79
Mucor racemosus 68
Mucor lusitanicus 65
Mucor velutinosus 62
Choanephora cucurbitarum 56
Mucor irregularis 56
Massospora cicadina 54
Gilbertella persicaria 53

======03_Zt_Taxonomy_report.tabular
Massospora cicadina 32
Gigaspora margarita 25
Entomophthora muscae 21
Gigaspora rosea 21
Mucor irregularis 16
Diversispora epigaea 15
Rhizophagus clarus 12
Mucor lusitanicus 11
Mucor mucedo 11
Mucor racemosus 7

======04_Asp_Taxonomy_report.tabular
Massospora cicadina 32
Gigaspora margarita 31
Gigaspora rosea 28
Diversispora epigaea 17
Entomophthora muscae 15
Mucor circinelloides 7
Thamnidium elegans 7
Mucor irregularis 6
Geosiphon pyriformis 5
Mucor lusitanicus 5

======05_G_Taxonomy_report.tabular
Gigaspora margarita 45
Massospora cicadina 41
Gigaspora rosea 30
Entomophthora muscae 23
Diversispora epigaea 22
Rhizophagus clarus 12
Geosiphon pyriformis 9
Mucor irregularis 9
Rhizophagus irregularis 8
Cunninghamella bertholletiae 6

======06_Ol_Taxonomy_report.tabular
Mucor irregularis 60
Gigaspora margarita 40
Massospora cicadina 36
Gigaspora rosea 33
Entomophthora muscae 28
Mucor racemosus 28
Mucor circinelloides 25
Mucor mucedo 22
Mucor lusitanicus 21
Thamnidium elegans 20

======07_Pr_Taxonomy_report.tabular
Gigaspora margarita 41
Massospora cicadina 38
Entomophthora muscae 19
Rhizophagus clarus 18
Gigaspora rosea 17
Diversispora epigaea 15
Mucor piriformis 9
Linnemannia elongata 7
Podila humilis 6
Rhizophagus irregularis 6

======08_Ar_Taxonomy_report.tabular
Neoconidiobolus thromboides 35
Massospora cicadina 17
Gigaspora margarita 14
Gigaspora rosea 10
Entomophthora muscae 8
Cunninghamella blakesleeana 5
Rhizophagus irregularis 5
Backusella circina 4
Diversispora epigaea 4
Halteromyces radiatus 4

```



# Trinotate

Wiki: https://github.com/Trinotate/Trinotate/wiki/Software-installation-and-data-required

Pipeline https://perun.biochem.dal.ca/manuals/trinityrnaseq_r2013-02-25/annotation/Trinotate.html

Reuse script used for OPM analysis.


```{bash, eval=FALSE}

#!/bin/bash
#SBATCH --job-name=trinotate
#SBATCH --nodes=1
#SBATCH --partition=medium
#SBATCH --time=48:00:00
#SBATCH --cpus-per-task=8
#SBATCH --mem=40gb
#SBATCH --mail-type=END
#SBATCH --mail-user=johan.zicola@uni-goettingen.de
#SBATCH --output=trinotate.%J.out
#SBATCH --error=trinotate.%J.err


mem_gbytes=$(( $SLURM_MEM_PER_NODE/1024 ))

echo "Starting at $(date)"
echo "Job submitted to the ${SLURM_JOB_PARTITION} partition, the default partition on ${SLURM_CLUSTER_NAME}"
echo "Job name: ${SLURM_JOB_NAME}, Job ID: ${SLURM_JOB_ID}"
echo "I have ${SLURM_CPUS_ON_NODE} CPUs and ${mem_gbytes}GiB of RAM on compute node $(hostname)"

# Modules to load
module purge

module load rev/21.12 gcc/10.2.0 hmmer/3.3 ncbi-blast/2.11.0

# Test that all programs needed are installed
test_program(){
  command -v $1 >/dev/null || \
    { echo >&2 " $1 required but not installed. Aborting"; exit 1; }
}

test_program "diamond"
test_program "cmpress"
test_program "hmmpress"
test_program "tmhmm"
test_program "signalp6"
test_program "makeblastdb"

# Variables passed to sbatch
# NAME_SPECIES
# Output from Trinity $GENE_TRANS_MAP and $TRINITY_FASTA
# Output from Transdecoder $TRANSDECODER_PEP

# Variables
TRINOTATE_HOME="/usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1"

#NAME_SPECIES is provided as argument to sbatch
sqlite_db="/usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1/Trinotate_${NAME_SPECIES}.sqlite"


echo "############################"
echo "Trinotate create"
date
echo "############################"

$TRINOTATE_HOME/Trinotate --create \
    --db $sqlite_db \
    --trinotate_data_dir $TRINOTATE_HOME/TRINOTATE_DATA_DIR \
    --use_diamond

echo "###############################"
echo "Trinotate Init"
date
echo "###############################"


$TRINOTATE_HOME/Trinotate --db $sqlite_db --init \
           --gene_trans_map $GENE_TRANS_MAP \
           --transcript_fasta $TRINITY_FASTA \
           --transdecoder_pep $TRANSDECODER_PEP

echo "##############################"
echo  "Trinotate Run"
date
echo "##############################"
 
$TRINOTATE_HOME/Trinotate --db ${sqlite_db} --CPU ${SLURM_CPUS_ON_NODE} --run ALL \
  --trinotate_data_dir $TRINOTATE_HOME/TRINOTATE_DATA_DIR \
    --transcript_fasta $TRINITY_FASTA \
      --transdecoder_pep $TRANSDECODER_PEP \
      --use_diamond

echo "###########################"
echo "Generating report table"
date
echo "###########################"

$TRINOTATE_HOME/Trinotate --db ${sqlite_db} \
  --report --incl_pep --incl_trans > Trinotate_report.tsv

##################
## Misc value adds
##################

# Extract GO terms
${TRINOTATE_HOME}/util/extract_GO_assignments_from_Trinotate_xls.pl \
  --Trinotate_xls Trinotate_report.tsv -G -I > Trinotate_report.tsv.gene_ontology

# Generate trinotate report summary statistics
${TRINOTATE_HOME}/util/report_summary/trinotate_report_summary.pl \
  Trinotate_report.tsv Trinotate_report_stats

echo "##########################"
echo "done.  See annotation summary file:  Trinotate_report.tsv"
date
echo "##########################"



```

```{bash, eval=FALSE}


# Variables passed to sbatch
# Output from Trinity $GENE_TRANS_MAP and $TRINITY_FASTA
# Output from Transdecoder $TRANSDECODER_PEP

work_path="/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects"
while read i; do
  mkdir $i
  cd $i 
  sbatch --export=NAME_SPECIES=$i,GENE_TRANS_MAP=${work_path}/trinity_output/summary/${i}_trinity.fasta.gene_trans_map,TRINITY_FASTA=${work_path}/trinity_output/summary/${i}_trinity.fasta,TRANSDECODER_PEP=${work_path}/transdecoder_output/${i}_trinity.fasta.transdecoder.pep ../04_trinotate.sh
  cd ..
done < list_species.txt

```


Steps done before being killed

```{bash, eval=FALSE}

Running CMD: /usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1/Trinotate --db /usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1/Trinotate_Anchomenus_dorsalis.sqlite --LOAD_swissprot_blastp uniprot_sprot.diamond.blastp.outfmt6
-LOADING as per --LOAD_swissprot_blastp

CMD: /home/uni08/zicola/bin/Trinotate-Trinotate-v4.0.1/util/trinotateSeqLoader/Trinotate_BLAST_loader.pl --sqlite /usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1/Trinotate_Anchomenus_dorsalis.sqlite --outfmt6 uniprot_sprot.diamond.blastp.outfmt6 --prog blastp --dbtype Swissprot
Can't locate DBI.pm in @INC (you may need to install the DBI module) (@INC contains: /home/uni08/zicola/bin/Trinotate-Trinotate-v4.0.1/util/trinotateSeqLoader/../../PerlLib /usr/users/zicola/perl5/lib/perl5/x86_64-linux-thread-multi /usr/users/zicola/perl5/lib/perl5 /usr/local/lib64/perl5 /usr/local/share/perl5 /usr/lib64/perl5/vendor_perl /usr/share/perl5/vendor_perl /usr/lib64/perl5 /usr/share/perl5) at /home/uni08/zicola/bin/Trinotate-Trinotate-v4.0.1/util/trinotateSeqLoader/Trinotate_BLAST_loader.pl line 9.
BEGIN failed--compilation aborted at /home/uni08/zicola/bin/Trinotate-Trinotate-v4.0.1/util/trinotateSeqLoader/Trinotate_BLAST_loader.pl line 9.
Error, cmd: /home/uni08/zicola/bin/Trinotate-Trinotate-v4.0.1/util/trinotateSeqLoader/Trinotate_BLAST_loader.pl --sqlite /usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1/Trinotate_Anchomenus_dorsalis.sqlite --outfmt6 uniprot_sprot.diamond.blastp.outfmt6 --prog blastp --dbtype Swissprot died with ret 512 at /usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1/Trinotate line 338.
Error, cmd: /usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1/Trinotate --db /usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1/Trinotate_Anchomenus_dorsalis.sqlite --LOAD_swissprot_blastp uniprot_sprot.diamond.blastp.outfmt6 died with ret 512 No such file or directory at /home/uni08/zicola/bin/Trinotate-Trinotate-v4.0.1/PerlLib/Pipeliner.pm line 187.
        Pipeliner::run(Pipeliner=HASH(0x56249c3268b0)) called at /usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1/Trinotate line 736
        main::Trinotate_RUN_annotators("/usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1/Trinotate_An"..., ARRAY(0x56249bf92f20), 8, "/usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1/TRINOTATE_DA"..., "/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgs"..., "/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgs"...) called at /usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1/Trinotate line 308
-REPORT being generated.
CMD: /home/uni08/zicola/bin/Trinotate-Trinotate-v4.0.1/util/Trinotate_report_writer.pl --sqlite /usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1/Trinotate_Anchomenus_dorsalis.sqlite -E 1e-5 --pfam_cutoff DNC  --incl_pep  --incl_trans
Can't locate DBI.pm in @INC (you may need to install the DBI module) (@INC contains: /home/uni08/zicola/bin/Trinotate-Trinotate-v4.0.1/util/../PerlLib /usr/users/zicola/perl5/lib/perl5/x86_64-linux-thread-multi /usr/users/zicola/perl5/lib/perl5 /usr/local/lib64/perl5 /usr/local/share/perl5 /usr/lib64/perl5/vendor_perl /usr/share/perl5/vendor_perl /usr/lib64/perl5 /usr/share/perl5) at /home/uni08/zicola/bin/Trinotate-Trinotate-v4.0.1/util/../PerlLib/Sqlite_connect.pm line 15.


```



## Troubleshoot perl library installation

After the migration, I had to reinstall my perl libraries. I notice that some of the libraries I installed before cannot be reinstalled (`cpan install DBI`) due to a `cc1: fatal error: inaccessible plugin file /opt/sw/rev/21.12/haswell/gcc-4.8.5/gcc-10.2.0-potfoy/lib/gcc/x86_64-pc-linux-gnu/10.2.0/plugin/gcc-annobin.so expanded from short plugin name gcc-annobin: No such file or directory` error. I found on the web several posts about the annobin issue and one suggests installing a local Perl installation with Spack (https://github.com/openhpc/ohpc/issues/1364).

I may think this 'annobin' gcc plugin was present on the previous SCC nodes OS (Scientific Linux 7) but was probably removed in the current OS (Rocky Linux 8). It seems that the GCC plugin annobin can be installed on Rocky 8 (https://rockylinux.pkgs.org/8/rockylinux-devel-x86\_64/annobin-11.13-2.el8.x86\_64.rpm.html). Would it be possible to install it or do I have to install my own gcc?

Note that I tried with all versions of gcc available and all failed:

```
module purge
module load rev/20.12 gcc/7.5.0
cpan instal DBI
#cc1: error: unrecognized command line option '-fstack-clash-protection'
#cc1: error: unrecognized command line option '-fcf-protection'
#cc1: fatal error: inaccessible plugin file /opt/sw/rev/20.12/haswell/gcc-9.3.0/gcc-7.5.0-owlabd/lib/gcc/x86_64-pc-linux-gnu/7.5.0/plugin/gcc-annobin.so expanded from short plugin name gcc-annobin: No such file or directory

# Nope
module purge
module load rev/20.12 gcc/8.4.0
cpan instal DBI
#cc1: fatal error: inaccessible plugin file /opt/sw/rev/20.12/haswell/gcc-4.8.5/gcc-8.4.0-anizkf/lib/gcc/x86_64-pc-linux-gnu/8.4.0/plugin/gcc-annobin.so expanded from short plugin name gcc-annobin: No such file or directory

module purge
module load rev/20.12 gcc/9.3.0
cpan instal DBI
# cc1: fatal error: inaccessible plugin file /opt/sw/rev/20.12/haswell/gcc-8.4.0/gcc-9.3.0-a466o5/lib/gcc/x86_64-pc-linux-gnu/9.3.0/plugin/gcc-annobin.so expanded from short plugin name gcc-annobin: No such file or directory

module purge
module load rev/20.12  gcc/9.5.0
cpan instal DBI
# cc1: fatal error: inaccessible plugin file /opt/sw/rev/23.12/linux-scientific7-haswell/gcc-4.8.5/gcc-9.5.0-y5n2mx/lib/gcc/x86_64-pc-linux-gnu/9.5.0/plugin/gcc-annobin.so expanded from short plugin name gcc-annobin: No such file or directory


module purge
module load rev/20.12 gcc/10.2.0
cpan instal DBI
# cc1: fatal error: inaccessible plugin file /opt/sw/rev/20.12/haswell/gcc-8.4.0/gcc-10.2.0-tc66mg/lib/gcc/x86_64-pc-linux-gnu/10.2.0/plugin/gcc-annobin.so expanded from short plugin name gcc-annobin: No such file or directory

module purge
module load gcc/11.4.0
cpan instal DBI
#cc1: fatal error: inaccessible plugin file /opt/sw/rev/23.12/linux-scientific7-haswell/gcc-4.8.5/gcc-11.4.0-jgoe5h/lib/gcc/x86_64-pc-linux-gnu/11.4.0/plugin/gcc-annobin.so expanded from short plugin name gcc-annobin: No such file or directory
```

Red Hat documentation on Annobin: https://docs.redhat.com/en/documentation/red\_hat\_developer\_toolset/9/html/user\_guide/chap-annobin#chap-Annobin


I tried `spack install perl` but the installation failed

```
ImportError: cannot bootstrap the "clingo" Python module from spec "clingo-bootstrap@spack+python %gcc target=x86_64" due to the following failures:
    'spack-install' raised InstallError: Terminating after first install failure: ProcessError: Command exited with status 1:
    '/usr/users/zicola/.spack/bootstrap/store/linux-rocky8-x86_64/gcc-10.2.0/perl-5.34.0-hhpsqrp435rgen7twicgyvpz5tpy5tbs/bin/perl' 'Makefile.PL'
    Please run `spack -d spec zlib` for more verbose error messages
```

I tried several times `spack -d spec zlib` but it does not work.



Mnay posts about the issue with annobin. One recommendation is to install perl in local using spack.


Download archive https://metacpan.org/pod/DBI and install manually DBI-1.647.tgz


```{bash, eval=FALSE}
cd ~/bin
tar -zxvf DBI-1.647.tgz
cd DBI-1.647

perl Makefile.PL PREFIX=/usr/users/zicola/perl5/lib/perl5
make

#cc1: fatal error: inaccessible plugin file /opt/sw/rev/23.12/linux-scientific7-haswell/gcc-4.8.5/gcc-11.4.0-jgoe5h/lib/gcc/x86_64-pc-linux-gnu/11.4.0/plugin/gcc-annobin.so expanded from short plugin name gcc-annobin: No such file or directory

#make test
#make install

```


```{bash, eval=FALSE}

spack install perl

# Error message, recommends to type 'spack -d spec zlib'
spack -d spec zlib


```


### Solution => Use apptainer


# Test apptainer

sbatch script `sbatch_apptainer.sh`. Launch from dir `/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects`

```{bash, eval=FALSE}


#!/bin/bash
#SBATCH --job-name=trinotate
#SBATCH --nodes=1
#SBATCH --partition=medium
#SBATCH --time=00:05:00
#SBATCH --cpus-per-task=1
#SBATCH --mem=1G
#SBATCH --output=trinotate.%J.out
#SBATCH --error=trinotate.%J.err

module load apptainer

# Pass to apptainer the variable NEWHOME
export APPTAINERENV_NEWHOME=/usr/users/zicola/newhome
export APPTAINERENV_SPECIES=$NAMESPECIES

apptainer run -C -B /usr/users/npghpc/ \
   -B /usr/users/zicola /usr/users/zicola/bin/trinotate.v4.0.2.simg /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/test_apptainer.sh

```


shell script `test_apptainer.sh`

```{bash, eval=FALSE}

export HOME=$NEWHOME
export TRINOTATE_HOME=/usr/local/src/Trinotate
cd $HOME
pwd > output_pwd.txt

#${TRINOTATE_HOME}/Trinotate --help > help_trinotate.txt

```

Make the script executable `chmod +x test_apptainer.sh`

```{bash, eval=FALSE}

sbatch sbatch_apptainer.sh

cat ~/newhome/output_pwd.txt
/usr/users/zicola/newhome

# Worked
```




##Simple functions


Way to call Trinotate

```{bash, eval=FALSE}

module load apptainer

PATH_DIR="/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects"

export TRINOTATE_HOME=/usr/local/src/Trinotate

apptainer exec --bind $PATH_DIR -e ~/bin/trinotate.v4.0.2.simg $TRINOTATE_HOME/Trinotate 


# Direct run within the container
apptainer shell --bind "/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects" ~/bin/trinotate.v4.0.2.simg

cd ~
apptainer shell ~/bin/trinotate.v4.0.2.simg

```


Add extra step for tmhmm and signal in the pipeline used for OPM


## Create database

What this step does it just copy the `TrinotateBoilerplate.sqlite` (340 Mb) file from TRINOTATE_DATA_DIR into the name and folder we specify for the our database. Include --use_diamond if you plan to use diamond blast with Trinotate, otherwise NCBI blast+ will be used for database preparation.

https://github.com/Trinotate/Trinotate/wiki/Software-installation-and-data-required

```{bash, eval=FALSE}
PATH_DIR="/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinotate_output"

# Variables (I moved the folder TRINOTATE_DATA_DIR locally for the run to avoid path mess with apptainer)
# Was originally in /usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1

export TRINOTATE_DATA_DIR=/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinotate_output/TRINOTATE_DATA_DIR

export TRINOTATE_HOME=/usr/local/src/Trinotate

PATH_DIR="/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinotate_output"

while read NAME_SPECIES; do
   mkdir $NAME_SPECIES
   
   apptainer exec --bind $PATH_DIR -e ~/bin/trinotate.v4.0.2.simg $TRINOTATE_HOME/Trinotate \
      --db $NAME_SPECIES/database_${NAME_SPECIES}.sqlite \
      --create \
      --trinotate_data_dir $TRINOTATE_DATA_DIR \
      --use_diamond
done < list_species.txt

    
```



## Download manually eggnog.db.gz

I can't gunzip the eggnog.db.gz so I think it is corrupted. Redownload from the eggnog5.embl.de website

```{bash, eval=FALSE}
http://eggnog5.embl.de/download/emapperdb-5.0.2/
../
eggnog.db.gz                                       02-Mar-2021 14:01      6G
eggnog.taxa.tar.gz                                 02-Mar-2021 14:04     69M
eggnog_proteins.dmnd.gz                            02-Mar-2021 14:04      5G
mmseqs.tar.gz                                      02-Mar-2021 14:07      5G
pfam.tar.gz                                        02-Mar-2021 14:08    938M

```

There are from March 2nd 2021, as my files downloaded during the create step of Trinity

```{bash, eval=FALSE}
cd /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinotate_output/TRINOTATE_DATA_DIR/EGGNOG_DATA_DIR
# Old file
-rw-r----- 1 zicola UANP 6.4G Mar  2  2021 eggnog.db.gz
md5sum eggnog.db.gz
f58b698b3f0bfac44ec8301bb83eb517  eggnog.db.gz

mv eggnog.db.gz eggnog.db.gz.old

wget http://eggnog5.embl.de/download/emapperdb-5.0.2/eggnog.db.gz

md5sum eggnog.db.gz
60688b3e21d0afbab97002dfb78ac324  eggnog.db.gz
# Indeedm the md5 is different

srun -p medium --time=05:00:00 gunzip eggnog.db.gz
# Now 39 Gb
-rw-r----- 1 zicola UANP  39G Mar  2  2021 eggnog.db
```



## Trinotate


```{bash, eval=FALSE}
#!/bin/bash
#SBATCH --job-name=trinotate
#SBATCH --nodes=1
#SBATCH --partition=medium
#SBATCH --time=16:00:00
#SBATCH --cpus-per-task=10
#SBATCH --mem=40gb
#SBATCH --mail-type=END
#SBATCH --mail-user=johan.zicola@uni-goettingen.de
#SBATCH --output=trinotate.%J.out
#SBATCH --error=trinotate.%J.err

mem_gbytes=$(( $SLURM_MEM_PER_NODE/1024 ))

echo "Starting at $(date)"
echo "Job submitted to the ${SLURM_JOB_PARTITION} partition, the default partition on ${SLURM_CLUSTER_NAME}"
echo "Job name: ${SLURM_JOB_NAME}, Job ID: ${SLURM_JOB_ID}"
echo "I have ${SLURM_CPUS_ON_NODE} CPUs and ${mem_gbytes}GiB of RAM on compute node $(hostname)"

# Modules to load
module purge
module load apptainer

# Test that all programs needed are installed
test_program(){
  command -v $1 >/dev/null || \
    { echo >&2 " $1 required but not installed. Aborting"; exit 1; }
}

test_program "tmhmm"
test_program "signalp6"

# Variables passed to sbatch
# NAME_SPECIES
# Output from Trinity $GENE_TRANS_MAP and $TRINITY_FASTA
# Output from Transdecoder $TRANSDECODER_PEP

PATH_DIR="/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects"

# Variables (I moved the folder TRINOTATE_DATA_DIR locally for the run to avoid path mess with apptainer)
# Was originally in /usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1
export TRINOTATE_DATA_DIR=${PATH_DIR}/trinotate_output/TRINOTATE_DATA_DIR

export TRINOTATE_HOME=/usr/local/src/Trinotate

#NAME_SPECIES is provided as argument to sbatch (Do not put a uppercase to trinotate_ suffix, it won't work)
sqlite_db="${PATH_DIR}/trinotate_output/${NAME_SPECIES}/database_${NAME_SPECIES}.sqlite"


echo "###############################"
echo "Trinotate Init"
date
echo "###############################"

apptainer exec --bind $PATH_DIR -e ~/bin/trinotate.v4.0.2.simg $TRINOTATE_HOME/Trinotate \
   --db $sqlite_db --init \
   --gene_trans_map $GENE_TRANS_MAP \
   --transcript_fasta $TRINITY_FASTA \
   --transdecoder_pep $TRANSDECODER_PEP

echo "##############################"
echo  "Trinotate Run"
date
echo "##############################"

apptainer exec --bind $PATH_DIR -e ~/bin/trinotate.v4.0.2.simg $TRINOTATE_HOME/Trinotate \
   --CPU ${SLURM_CPUS_ON_NODE} \
   --db $sqlite_db \
   --run ALL \
   --trinotate_data_dir $TRINOTATE_DATA_DIR \
   --transcript_fasta $TRINITY_FASTA \
   --transdecoder_pep $TRANSDECODER_PEP \
   --use_diamond

echo "##############################"
echo  "Trinotate Run signalp"
date
echo "##############################"

signalp6 --fastafile $TRANSDECODER_PEP \
   --output_dir ${PATH_DIR}/trinotate_output/${NAME_SPECIES}/sigP6outdir \
   --format none --organism euk --mode fast


apptainer exec --bind $PATH_DIR -e ~/bin/trinotate.v4.0.2.simg $TRINOTATE_HOME/Trinotate \
   --db $sqlite_db \
   --LOAD_signalp ${PATH_DIR}/trinotate_output/${NAME_SPECIES}/sigP6outdir/output.gff3


echo "##############################"
echo  "Trinotate Run TMHMM"
date
echo "##############################"

tmhmm --short $TRANSDECODER_PEP > ${PATH_DIR}/trinotate_output/${NAME_SPECIES}/tmhmm.v2.out

apptainer exec --bind $PATH_DIR -e ~/bin/trinotate.v4.0.2.simg $TRINOTATE_HOME/Trinotate \
   --db $sqlite_db \
   --LOAD_tmhmmv2 ${PATH_DIR}/trinotate_output/${NAME_SPECIES}/tmhmm.v2.out

echo "###########################"
echo "Generating report table"
date
echo "###########################"

apptainer exec --bind $PATH_DIR -e ~/bin/trinotate.v4.0.2.simg $TRINOTATE_HOME/Trinotate \
   --db ${sqlite_db} \
   --report \
   --incl_pep \
   --incl_trans > ${PATH_DIR}/trinotate_output/${NAME_SPECIES}/Trinotate_report.tsv


##################
## Misc value adds
date
##################

# Extract GO terms

apptainer exec --bind $PATH_DIR -e ~/bin/trinotate.v4.0.2.simg \
   $TRINOTATE_HOME/util/extract_GO_assignments_from_Trinotate_xls.pl \
  --Trinotate_xls Trinotate_report.tsv \
  -G -I > ${PATH_DIR}/trinotate_output/${NAME_SPECIES}/Trinotate_report.tsv.gene_ontology

# Generate trinotate report summary statistics
apptainer exec --bind $PATH_DIR -e ~/bin/trinotate.v4.0.2.simg \
   $TRINOTATE_HOME/util/report_summary/trinotate_report_summary.pl \
  ${PATH_DIR}/trinotate_output/${NAME_SPECIES}/Trinotate_report.tsv \
  ${PATH_DIR}/trinotate_output/${NAME_SPECIES}/Trinotate_report_stats


echo "##########################"
echo "done.  See annotation summary file:  Trinotate_report.tsv"
date
echo "##########################"

```

Run one by one because I notice intermediate files and report are written in $HOME without running number, which will lead to potential conflicts for parallel runs.

```{bash, eval=FALSE}

work_path="/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects"

i="Amara_similata"
cd $i
sbatch --export=NAME_SPECIES=$i,GENE_TRANS_MAP=${work_path}/trinity_output/summary/${i}_trinity.fasta.gene_trans_map,TRINITY_FASTA=${work_path}/trinity_output/summary/${i}_trinity.fasta,TRANSDECODER_PEP=${work_path}/transdecoder_output/${i}_trinity.fasta.transdecoder.pep ../04_trinotate.sh

```



Job killed due to error

```{bash, eval=FALSE}
CMD: /usr/local/src/Trinotate/util/report_summary/../gene_ontology/Trinotate_GO_to_SLIM.pl /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinotate_output/Amara_similata/Trinotate_report_stats.GO > /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinotate_output/Amara_similata/Trinotate_report_stats.GO.slim
CMD: /usr/local/src/Trinotate/util/report_summary/util/make_cXp_html.pl /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinotate_output/Amara_similata/Trinotate_report_stats > /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinotate_output/Amara_similata/Trinotate_report_stats.cXp_summary.html
Can't use an undefined value as an ARRAY reference at /usr/local/src/Trinotate/util/report_summary/util/../../../TrinotateWeb/cgi-bin/PerlLib/CanvasXpress/Sunburst.pm line 77.
Error, cmd:
/usr/local/src/Trinotate/util/report_summary/util/make_cXp_html.pl /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinotate_output/Amara_similata/Trinotate_report_stats > /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinotate_output/Amara_similata/Trinotate_report_stats.cXp_summary.html
 died with ret (65280) at /usr/local/src/Trinotate/util/report_summary/../../PerlLib/Process_cmd.pm line 19, <$fh> line 1.
        Process_cmd::process_cmd("/usr/local/src/Trinotate/util/report_summary/util/make_cXp_ht"...) called at /usr/local/src/Trinotate/util/report_summary/trinotate_report_summary.pl line 102

```


```{bash, eval=FALSE}

apptainer exec --bind $PATH_DIR -e ~/bin/trinotate.v4.0.2.simg $TRINOTATE_HOME/Trinotate \
   --db ${sqlite_db} \
   --report \
   --incl_pep \
   --incl_trans > ${PATH_DIR}/trinotate_output/${NAME_SPECIES}/Trinotate_report.tsv

```



```{bash, eval=FALSE}
#!/bin/bash
#SBATCH --job-name=test_pwd
#SBATCH --nodes=1
#SBATCH --partition=medium
#SBATCH --time=00:00:10
#SBATCH --cpus-per-task=1
#SBATCH --mem=40mb
#SBATCH --output=trinotate.%J.out
#SBATCH --chdir="/usr/users/zicola"

pwd


```



## Problem with DB download


## Remake full database

I think I screwed up something in TRINOTATE_DATA_DIR so download everything anew.

```{bash, eval=FALSE}
module load apptainer

export TRINOTATE_DATA_DIR=/scratch/users/zicola/TRINOTATE_DATA_DIR
export TRINOTATE_HOME=/usr/local/src/Trinotate

PATH_DIR="/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinotate_output"

apptainer exec --bind /scratch/users/zicola -e ~/bin/trinotate.v4.0.2.simg $TRINOTATE_HOME/Trinotate \
      --db test.sqlite \
      --create \
      --trinotate_data_dir $TRINOTATE_DATA_DIR \
      --use_diamond

cd /scratch/users/zicola

mkdir TRINOTATE_DATA_DIR

srun --time=06:00:00 -p medium --mem=20Gb bash create_database.sh

```


Rfam files did not download sucessfully so I downloaded them manually from https://ftp.ebi.ac.uk/pub/databases/Rfam/CURRENT/

Kill the job at this point, then do the steps my self and relaunch

```{bash, eval=FALSE}
cd /scratch/users/zicola/trinotate

wget ftp://ftp.ebi.ac.uk/pub/databases/Rfam/CURRENT/Rfam.cm.gz

gunzip Rfam.cm.gz
cmpress Rfam.cm

Working...    done.
Pressed and indexed 4178 CMs and p7 HMM filters (4178 names and 4178 accessions).
Covariance models and p7 filters pressed into binary file:  Rfam.cm.i1m
SSI index for binary covariance model file:                 Rfam.cm.i1i
Optimized p7 filter profiles (MSV part)  pressed into:      Rfam.cm.i1f
Optimized p7 filter profiles (remainder) pressed into:      Rfam.cm.i1p


wget https://ftp.ebi.ac.uk/pub/databases/Rfam/CURRENT/Rfam.clanin

mv Rfam.* /scratch/users/zicola/trinotate/TRINOTATE_DATA_DIR

# Create a fake checkpoint so that it does not try to download the files again
touch TRINOTATE_DATA_DIR/__chckpts/cmpress_Rfam_cm.ok

srun --time=06:00:00 -o output2.txt -p medium --mem=20Gb bash create_database.sh

```


I also moved the EGGNOG files as they are heavy and I have alread downloaded them recently

```{bash, eval=FALSE}
cd /scratch/users/zicola/TRINOTATE_DATA_DIR
mv /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinotate_output/TRINOTATE_DATA_DIR_OLD/EGGNOG_DATA_DIR .
```


```{bash, eval=FALSE}
module load apptainer

export TRINOTATE_DATA_DIR=/scratch/users/zicola/TRINOTATE_DATA_DIR
export TRINOTATE_HOME=/usr/local/src/Trinotate

PATH_DIR="/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinotate_output"

apptainer exec --bind /scratch/users/zicola -e ~/bin/trinotate.v4.0.2.simg $TRINOTATE_HOME/Trinotate \
      --db test.sqlite \
      --create \
      --trinotate_data_dir $TRINOTATE_DATA_DIR \
      --use_diamond
```

```

-CREATING test.sqlite and populating data dir: /scratch/users/zicola/TRINOTATE_DATA_DIR
WARNING: SQLITE database test.sqlite already exists and wont be replaced.
-sqlite db test.sqlite already exists and is not being replaced. If a new boilerplate is required, copy db from: /scratch/users/zicola/TRINOTATE_DATA_DIR/TrinotateBoilerplate.sqlite
-- Skipping CMD: /usr/local/src/Trinotate/util/admin/Build_Trinotate_Boilerplate_SQLite_db.pl TrinotateBoilerplate, checkpoint [/scratch/users/zicola/TRINOTATE_DATA_DIR/__chckpts/build_boilerplate.ok] exists.
-- Skipping CMD: mv TrinotateBoilerplate.sqlite pfam2go go-basic.obo pfam2go.tab NOG.annotations.tsv.gz /scratch/users/zicola/TRINOTATE_DATA_DIR/, checkpoint [/scratch/users/zicola/TRINOTATE_DATA_DIR/__chckpts/copy_boilerplate_to_datadir.ok] exists.
-- Skipping CMD: makeblastdb -in uniprot_sprot.pep -dbtype prot, checkpoint [/scratch/users/zicola/TRINOTATE_DATA_DIR/__chckpts/make_sprot_blast_db.ok] exists.
-- Skipping CMD: diamond makedb --in uniprot_sprot.pep -d uniprot_sprot, checkpoint [/scratch/users/zicola/TRINOTATE_DATA_DIR/__chckpts/make_diamond_sprot_db.ok] exists.
-- Skipping CMD: mv uniprot_sprot* /scratch/users/zicola/TRINOTATE_DATA_DIR/, checkpoint [/scratch/users/zicola/TRINOTATE_DATA_DIR/__chckpts/relocate_sprot_db.ok] exists.
-- Skipping CMD: gunzip Pfam-A.hmm.gz, checkpoint [/scratch/users/zicola/TRINOTATE_DATA_DIR/__chckpts/gunzip_pfam.ok] exists.
-- Skipping CMD: hmmpress Pfam-A.hmm, checkpoint [/scratch/users/zicola/TRINOTATE_DATA_DIR/__chckpts/hmmpress_pfam_db.ok] exists.
-- Skipping CMD: mv Pfam-A.hmm* /scratch/users/zicola/TRINOTATE_DATA_DIR/, checkpoint [/scratch/users/zicola/TRINOTATE_DATA_DIR/__chckpts/relocate_pfam_db.ok] exists.
-- Skipping CMD: gunzip Pfam-A.hmm.gz, checkpoint [/scratch/users/zicola/TRINOTATE_DATA_DIR/__chckpts/gunzip_pfam.ok] exists.
-- Skipping CMD: hmmpress Pfam-A.hmm, checkpoint [/scratch/users/zicola/TRINOTATE_DATA_DIR/__chckpts/hmmpress_pfam_db.ok] exists.
-- Skipping CMD: mv Pfam-A.hmm* /scratch/users/zicola/TRINOTATE_DATA_DIR/, checkpoint [/scratch/users/zicola/TRINOTATE_DATA_DIR/__chckpts/relocate_pfam_db.ok] exists.
-- Skipping CMD: wget ftp://ftp.ebi.ac.uk/pub/databases/Rfam/CURRENT/Rfam.cm.gz, checkpoint [/scratch/users/zicola/TRINOTATE_DATA_DIR/__chckpts/wget_Rfam.cm.gz] exists.
-- Skipping CMD: gunzip Rfam.cm.gz, checkpoint [/scratch/users/zicola/TRINOTATE_DATA_DIR/__chckpts/gunzip_Rfam.cm.gz] exists.
-- Skipping CMD: wget ftp://ftp.ebi.ac.uk/pub/databases/Rfam/CURRENT/Rfam.clanin, checkpoint [/scratch/users/zicola/TRINOTATE_DATA_DIR/__chckpts/wget_clanin.cm.gz] exists.
-- Skipping CMD: cmpress Rfam.cm, checkpoint [/scratch/users/zicola/TRINOTATE_DATA_DIR/__chckpts/cmpress_Rfam_cm.ok] exists.
-- Skipping CMD: mv Rfam.* /scratch/users/zicola/TRINOTATE_DATA_DIR/, checkpoint [/scratch/users/zicola/TRINOTATE_DATA_DIR/__chckpts/relocate_Rfam_db.ok] exists.
-EGGNOG_MAPPER: installing eggnogmapper data at: /scratch/users/zicola/TRINOTATE_DATA_DIR/EGGNOG_DATA_DIR
* [Thu Jan 23 09:49:49 2025] Running CMD: download_eggnog_data.py --data_dir /scratch/users/zicola/TRINOTATE_DATA_DIR/EGGNOG_DATA_DIR -y
Skipping eggnog.db database (already present). Use -f to force download
Skipping eggnog.taxa.db database (already present). Use -f to force download
Skipping diamond database (or already present). Use -f to force download
Skipping novel families diamond database (or already present). Use -F and -f to force download
Skipping Pfam database (or already present). Use -P and -f to force download
Skipping MMseqs2 database (or already present). Use -M and -f to force download
No HMMER database requested. Use "-H -d taxid" to download the hmmer database for taxid
Finished.
```

It seems every it worked but I don't think the boiler plate included all files because of this checkpoint files.


```{bash, eval=FALSE}

module load apptainer

export TRINOTATE_DATA_DIR=/scratch/users/zicola/trinotate/TRINOTATE_DATA_DIR
export TRINOTATE_HOME=/usr/local/src/Trinotate

apptainer exec --bind /scratch/users/zicola -e ~/bin/trinotate.v4.0.2.simg $TRINOTATE_HOME/Trinotate \
      --db test.sqlite \
      --create \
      --trinotate_data_dir $TRINOTATE_DATA_DIR \
      --use_diamond


cd /scratch/users/zicola/trinotate

srun --time=06:00:00 -p medium --mem=10Gb bash create_database.sh

```


## Amara_similata

```{bash, eval=FALSE}
i="Amara_similata"
cd $i
sbatch --export=NAME_SPECIES=$i,GENE_TRANS_MAP=${work_path}/trinity_output/summary/${i}_trinity.fasta.gene_trans_map,TRINITY_FASTA=${work_path}/trinity_output/summary/${i}_trinity.fasta,TRANSDECODER_PEP=${work_path}/transdecoder_output/${i}_trinity.fasta.transdecoder.pep ../04_trinotate.sh
```


What happens in the background (using err file from slurm)

```{bash, eval=FALSE}

# Load transcripts into sqlite database
/usr/local/src/Trinotate/util/trinotateSeqLoader/TrinotateSeqLoader.pl \
   --sqlite /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinotate_output/Amara_si
milata/database_Amara_similata.sqlite \
   --gene_trans_map /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinity_output/summary/Amara_similata_trinity.fasta.gen
e_trans_map \
   --transcript_fasta /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinity_output/summary/Amara_similata_trinity.fasta \
      --transdecoder_pep /usr/use
rs/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/transdecoder_output/Amara_similata_trinity.fasta.transdecoder.pep \
   --bulk_load

# Diamong blastp
diamond blastp -d /scratch/users/zicola/TRINOTATE_DATA_DIR/uniprot_sprot \
   -q /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/transdecoder_output/Amara_similata_trinity.fasta.transdecoder.pep \
   -p 8 -k 1 -e 1e-5 -o \
   uniprot_sprot.diamond.blastp.outfmt6 --outfmt 6

# Load blastp
/usr/local/src/Trinotate/util/trinotateSeqLoader/Trinotate_BLAST_loader.pl --sqlite /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinotate_output/Amara_similata/database_Amara_similata.sqlite --outfmt6 uniprot_sprot.diamond.blastp.outfmt6 \
   --prog blastp \
   --dbtype Swissprot

# Diamong blastx
diamond blastx -d /scratch/users/zicola/TRINOTATE_DATA_DIR/uniprot_sprot -q /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinity_output/summary/Amara_similata_trinity.fasta \
   -p 8 -k 1 -e 1e-5 \
   -o uniprot_sprot.diamond.blastx.outfmt6 --outfmt 6

# Load blastx
/usr/local/src/Trinotate/util/trinotateSeqLoader/Trinotate_BLAST_loader.pl \
--sqlite /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinotate_output/Amara_similata/database_Amara_similata.sqlite --outfmt6 uniprot_sprot.diamond.blastx.outfmt6 \
   --prog blastx \
   --dbtype Swissprot

# hmmsearch (PFAM)
hmmsearch --cpu 8 --noali --domtblout TrinotatePFAM.out /scratch/users/zicola/TRINOTATE_DATA_DIR/Pfam-A.hmm /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/transdecoder_output/Amara_similata_trinity.fasta.transdecoder.pep > pfam.log

# PFAM
/usr/local/src/Trinotate/Trinotate --db /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinotate_output/Amara_similata/database_Amara_similata.sqlite --LOAD_pfam  TrinotatePFAM.out

# Load PFAM result
/usr/local/src/Trinotate/util/trinotateSeqLoader/Trinotate_PFAM_loader.pl --sqlite /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinotate_output/Amara_similata/database_Amara_similata.sqlite --pfam TrinotatePFAM.out

# Infernal
cmscan -Z 90 --cut_ga --rfam --nohmmonly --tblout infernal.out --fmt 2 --cpu 8 --clanin /scratch/users/zicola/TRINOTATE_DATA_DIR/Rfam.clanin /scratch/users/zicola/TRINOTATE_DATA_DIR/Rfam.cm /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinity_output/summary/Amara_similata_trinity.fasta > infernal.log

# Load
/usr/local/src/Trinotate/Trinotate --db /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinotate_output/Amara_similata/database_Amara_similata.sqlite --LOAD_infernal infernal.out

# emapper
emapper.py -i /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/transdecoder_output/Amara_similata_trinity.fasta.transdecoder.pep --cpu 8 --data_dir /scratch/users/zicola/TRINOTATE_DATA_DIR/EGGNOG_DATA_DIR -o eggnog_mapper

# Load
/usr/local/src/Trinotate/util/trinotateSeqLoader/Trinotate_EggnogMapper_loader.pl --sqlite /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinotate_output/Amara_similata/database_Amara_similata.sqlite --emapper eggnog_mapper.emapper.annotations

```




Still creating files in $HOME

```{bash, eval=FALSE}
zicola@gwdu101:~
$ ls -h | grep "Jan 23"
drwx------  5 zicola UANP         2 Jan 23  2024 R
-rw-r-----  1 zicola UANP      120M Jan 23 13:26 TrinotatePFAM.out
-rw-r-----  1 zicola UANP         0 Jan 23 12:50 __init.ok
drwxr-x---  2 zicola UANP         2 Jan 23 22:00 __trinotate_run_checkpts
-rw-r-----  1 zicola UANP       43M Jan 23 22:00 eggnog_mapper.emapper.annotations
-rw-r-----  1 zicola UANP      9.1M Jan 23 19:54 eggnog_mapper.emapper.hits
-rw-r-----  1 zicola UANP      2.3M Jan 23 22:00 eggnog_mapper.emapper.seed_orthologs
-rw-r-----  1 zicola UANP      129M Jan 23 19:41 infernal.log
-rw-r-----  1 zicola UANP       79K Jan 23 19:41 infernal.out
-rw-r-----  1 zicola UANP      134M Jan 23 13:26 pfam.log
-rw-r-----  1 zicola UANP      1.6M Jan 23 12:50 uniprot_sprot.diamond.blastp.outfmt6
-rw-r-----  1 zicola UANP      1.6M Jan 23 12:50 uniprot_sprot.diamond.blastx.outfmt6

```

It seems the flag `--workdir ${PATH_DIR}/trinotate_output/${NAME_SPECIES}` does not help to make Trinotate generate its output in the working directory


Check how many transcripts were annotated as rRNA:

```{bash, eval=FALSE}
grep "rRNA" infernal.out | wc -l
111


```




```{bash, eval=FALSE}

#!/bin/bash
#SBATCH --job-name=trinotate
#SBATCH --nodes=1
#SBATCH --partition=medium
#SBATCH --time=16:00:00
#SBATCH --cpus-per-task=1
#SBATCH --mem=20gb
#SBATCH --mail-type=END
#SBATCH --mail-user=johan.zicola@uni-goettingen.de
#SBATCH --output=trinotate.%J.out
#SBATCH --error=trinotate.%J.err

mem_gbytes=$(( $SLURM_MEM_PER_NODE/1024 ))

echo "Starting at $(date)"
echo "Job submitted to the ${SLURM_JOB_PARTITION} partition, the default partition on ${SLURM_CLUSTER_NAME}"
echo "Job name: ${SLURM_JOB_NAME}, Job ID: ${SLURM_JOB_ID}"
echo "I have ${SLURM_CPUS_ON_NODE} CPUs and ${mem_gbytes}GiB of RAM on compute node $(hostname)"

# Modules to load
module purge
module load apptainer

# Test that all programs needed are installed
test_program(){
  command -v $1 >/dev/null || \
    { echo >&2 " $1 required but not installed. Aborting"; exit 1; }
}

test_program "tmhmm"
test_program "signalp6"

# Variables passed to sbatch
# NAME_SPECIES
# Output from Trinity $GENE_TRANS_MAP and $TRINITY_FASTA
# Output from Transdecoder $TRANSDECODER_PEP

PATH_DIR="/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects"

# Variables (I moved the folder TRINOTATE_DATA_DIR locally for the run to avoid path mess with apptainer)
# Was originally in /usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1
export TRINOTATE_DATA_DIR=/scratch/users/zicola/TRINOTATE_DATA_DIR

export TRINOTATE_HOME=/usr/local/src/Trinotate

#NAME_SPECIES is provided as argument to sbatch (Do not put a uppercase to trinotate_ suffix, it won't work)
sqlite_db="${PATH_DIR}/trinotate_output/${NAME_SPECIES}/database_${NAME_SPECIES}.sqlite"


echo "##############################"
echo  "Trinotate Run signalp"
date
echo "##############################"

signalp6 --fastafile $TRANSDECODER_PEP \
   --output_dir ${PATH_DIR}/trinotate_output/${NAME_SPECIES}/sigP6outdir \
   --format none --organism euk --mode fast


apptainer exec --bind $PATH_DIR,$TRINOTATE_DATA_DIR -e ~/bin/trinotate.v4.0.2.simg $TRINOTATE_HOME/Trinotate \
   --db $sqlite_db \
   --LOAD_signalp ${PATH_DIR}/trinotate_output/${NAME_SPECIES}/sigP6outdir/output.gff3


echo "##############################"
echo  "Trinotate Run TMHMM"
date
echo "##############################"

tmhmm --short $TRANSDECODER_PEP > ${PATH_DIR}/trinotate_output/${NAME_SPECIES}/tmhmm.v2.out

apptainer exec --bind $PATH_DIR,$TRINOTATE_DATA_DIR -e ~/bin/trinotate.v4.0.2.simg $TRINOTATE_HOME/Trinotate \
   --db $sqlite_db \
   --LOAD_tmhmmv2 ${PATH_DIR}/trinotate_output/${NAME_SPECIES}/tmhmm.v2.out

echo "###########################"
echo "Generating report table"
date
echo "###########################"

apptainer exec --bind $PATH_DIR,$TRINOTATE_DATA_DIR -e ~/bin/trinotate.v4.0.2.simg $TRINOTATE_HOME/Trinotate \
   --db ${sqlite_db} \
   --report \
   --incl_pep \
   --incl_trans > ${PATH_DIR}/trinotate_output/${NAME_SPECIES}/Trinotate_report.tsv


##################
## Misc value adds
date
##################

# Extract GO terms

apptainer exec --bind $PATH_DIR,$TRINOTATE_DATA_DIR -e ~/bin/trinotate.v4.0.2.simg \
   $TRINOTATE_HOME/util/extract_GO_assignments_from_Trinotate_xls.pl \
  --Trinotate_xls Trinotate_report.tsv \
  -G -I > ${PATH_DIR}/trinotate_output/${NAME_SPECIES}/Trinotate_report.tsv.gene_ontology

# Generate trinotate report summary statistics
apptainer exec --bind $PATH_DIR,$TRINOTATE_DATA_DIR -e ~/bin/trinotate.v4.0.2.simg \
   $TRINOTATE_HOME/util/report_summary/trinotate_report_summary.pl \
  ${PATH_DIR}/trinotate_output/${NAME_SPECIES}/Trinotate_report.tsv \
  ${PATH_DIR}/trinotate_output/${NAME_SPECIES}/Trinotate_report_stats


echo "##########################"
echo "done.  See annotation summary file:  Trinotate_report.tsv"
date
echo "##########################"
```



# tmhmm alone


I just need to do tmhmm as I've already done signalp for Amara similara


```{bash, eval=FALSE}

#!/bin/bash
#SBATCH --job-name=trinotate_tmhmm
#SBATCH --nodes=1
#SBATCH --partition=medium
#SBATCH --time=06:00:00
#SBATCH --cpus-per-task=1
#SBATCH --mem=5G
#SBATCH --mail-type=END
#SBATCH --mail-user=johan.zicola@uni-goettingen.de
#SBATCH --output=trinotate_tmhmm.%J.out
#SBATCH --error=trinotate_tmhmm..%J.err

mem_gbytes=$(( $SLURM_MEM_PER_NODE/1024 ))

echo "Starting at $(date)"
echo "Job submitted to the ${SLURM_JOB_PARTITION} partition, the default partition on ${SLURM_CLUSTER_NAME}"
echo "Job name: ${SLURM_JOB_NAME}, Job ID: ${SLURM_JOB_ID}"
echo "I have ${SLURM_CPUS_ON_NODE} CPUs and ${mem_gbytes}GiB of RAM on compute node $(hostname)"

# Modules to load
module purge
module load apptainer

# Test that all programs needed are installed
test_program(){
  command -v $1 >/dev/null || \
    { echo >&2 " $1 required but not installed. Aborting"; exit 1; }
}

test_program "tmhmm"

# Path to trinotate executable
export TRINOTATE_HOME=/usr/local/src/Trinotate

# Path to Trinity and Transdecoder files
export WORKPATH=/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects
export GENE_TRANS_MAP=${WORKPATH}/trinity_output/summary/${SPECIES}_trinity.fasta.gene_trans_map
export TRINITY_FASTA=${WORKPATH}/trinity_output/summary/${SPECIES}_trinity.fasta
export TRANSDECODER_PEP=${WORKPATH}/transdecoder_output/${SPECIES}_trinity.fasta.transdecoder.pep 

# Database
export TRINOTATE_DATA_DIR=/scratch/users/zicola/TRINOTATE_DATA_DIR
export EGGNOG_DATA_DIR=/scratch/users/zicola/TRINOTATE_DATA_DIR/EGGNOG_DATA_DIR
export SQLITE_DB=${WORKPATH}/trinotate_output/${SPECIES}/database_${SPECIES}.sqlite

cd $WORKPATH

tmhmm --short $TRANSDECODER_PEP > ${WORKPATH}/trinotate_output/${SPECIES}/tmhmm.v2.out

apptainer exec --bind $WORKPATH -e ~/bin/trinotate.v4.0.2.simg $TRINOTATE_HOME/Trinotate \
   --db $SQLITE_DB \
   --LOAD_tmhmmv2 ${WORKPATH}/trinotate_output/${SPECIES}/tmhmm.v2.out
   

```

```{bash, eval=FALSE}

sbatch --export=SPECIES="Amara_similata" 05_tmhmm.sh

```

```{bash, eval=FALSE}

sacct -j 5431422   -o jobid,jobname,AllocCPUS,ReqMem,MaxRSS,Elapsed

sacct -j 5431422   -o jobid,jobname,AllocCPUS,ReqMem,MaxRSS,Elapsed
JobID           JobName  AllocCPUS     ReqMem     MaxRSS    Elapsed
------------ ---------- ---------- ---------- ---------- ----------
5431422      trinotate+          1         5G              00:27:02
5431422.bat+      batch          1                48284K   00:27:02
5431422.ext+     extern          1                     0   00:27:02

```

Took only 48 Mb of RAM and 27 min



# Troubelshoot path binding in apptainer

## Anchomenus_dorsalis

Now that I have TRINOTATE_DATA_DIR on another path, I need to add it to `--bind` so that the container can find the files.

```{bash, eval=FALSE}

module load apptainer

export TRINOTATE_DATA_DIR=/scratch/users/zicola/TRINOTATE_DATA_DIR
export TRINOTATE_HOME=/usr/local/src/Trinotate

export WORKDIR=/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinotate_output

PATH_DIR="/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinotate_output"

apptainer exec --bind $WORKDIR,$TRINOTATE_DATA_DIR -e ~/bin/trinotate.v4.0.2.simg $TRINOTATE_HOME/Trinotate \
      --db Anchomenus_dorsalis/database_Anchomenus_dorsalis.sqlite \
      --create \
      --trinotate_data_dir $TRINOTATE_DATA_DIR \
      --use_diamond

```

Worked!

Now run only the first steps

```{bash, eval=FALSE}
#!/bin/bash
#SBATCH --job-name=trinotate
#SBATCH --nodes=1
#SBATCH --partition=medium
#SBATCH --time=16:00:00
#SBATCH --cpus-per-task=8
#SBATCH --mem=10gb
#SBATCH --mail-type=END
#SBATCH --mail-user=johan.zicola@uni-goettingen.de
#SBATCH --output=trinotate.%J.out
#SBATCH --error=trinotate.%J.err

mem_gbytes=$(( $SLURM_MEM_PER_NODE/1024 ))

echo "Starting at $(date)"
echo "Job submitted to the ${SLURM_JOB_PARTITION} partition, the default partition on ${SLURM_CLUSTER_NAME}"
echo "Job name: ${SLURM_JOB_NAME}, Job ID: ${SLURM_JOB_ID}"
echo "I have ${SLURM_CPUS_ON_NODE} CPUs and ${mem_gbytes}GiB of RAM on compute node $(hostname)"

# Modules to load
module purge
module load apptainer

# Test that all programs needed are installed
test_program(){
  command -v $1 >/dev/null || \
    { echo >&2 " $1 required but not installed. Aborting"; exit 1; }
}

test_program "tmhmm"
test_program "signalp6"

# Variables passed to sbatch
# NAME_SPECIES
# Output from Trinity $GENE_TRANS_MAP and $TRINITY_FASTA
# Output from Transdecoder $TRANSDECODER_PEP

PATH_DIR="/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects"

# Variables (I moved the folder TRINOTATE_DATA_DIR locally for the run to avoid path mess with apptainer)
# Was originally in /usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1
export TRINOTATE_DATA_DIR=/scratch/users/zicola/TRINOTATE_DATA_DIR

export TRINOTATE_HOME=/usr/local/src/Trinotate

#NAME_SPECIES is provided as argument to sbatch (Do not put a uppercase to trinotate_ suffix, it won't work)
sqlite_db="${PATH_DIR}/trinotate_output/${NAME_SPECIES}/database_${NAME_SPECIES}.sqlite"


echo "###############################"
echo "Trinotate Init"
date
echo "###############################"

apptainer exec --workdir ${PATH_DIR}/trinotate_output/${NAME_SPECIES} --bind $PATH_DIR,$TRINOTATE_DATA_DIR -e ~/bin/trinotate.v4.0.2.simg $TRINOTATE_HOME/Trinotate \
   --db $sqlite_db --init \
   --gene_trans_map $GENE_TRANS_MAP \
   --transcript_fasta $TRINITY_FASTA \
   --transdecoder_pep $TRANSDECODER_PEP

echo "##############################"
echo  "Trinotate Run"
date
echo "##############################"

apptainer exec --workdir ${PATH_DIR}/trinotate_output/${NAME_SPECIES} --bind $PATH_DIR,$TRINOTATE_DATA_DIR -e ~/bin/trinotate.v4.0.2.simg $TRINOTATE_HOME/Trinotate \
   --CPU ${SLURM_CPUS_ON_NODE} \
   --db $sqlite_db \
   --run ALL \
   --trinotate_data_dir $TRINOTATE_DATA_DIR \
   --transcript_fasta $TRINITY_FASTA \
   --transdecoder_pep $TRANSDECODER_PEP \
   --use_diamond


```




```{bash, eval=FALSE}
work_path="/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects"

i="Anchomenus_dorsalis"
cd $i
sbatch --export=NAME_SPECIES=$i,GENE_TRANS_MAP=${work_path}/trinity_output/summary/${i}_trinity.fasta.gene_trans_map,TRINITY_FASTA=${work_path}/trinity_output/summary/${i}_trinity.fasta,TRANSDECODER_PEP=${work_path}/transdecoder_output/${i}_trinity.fasta.transdecoder.pep ../05_trinotate.sh


```


It seems signalP and tmhmmv2 worked but the not the report production (empty file).

Summary MEM needed

```{bash, eval=FALSE}

reportseff 5246016  --format +AllocTRES,TRESUsageInMax
    JobID    State       Elapsed  TimeEff   CPUEff   MemEff             AllocTRES                                          TRESUsageInMax
  5246016  COMPLETED    07:42:15   48.2%    98.8%    18.0%    billing=2,cpu=1,mem=20G,node=1   cpu=07:36:44,energy=0,fs/disk=2642600232,mem=3778500K,pages=2107,vmem=0


sacct -j 5246016 -o jobid,jobname,AllocCPUS,ReqMem,MaxRSS
JobID           JobName  AllocCPUS     ReqMem     MaxRSS
------------ ---------- ---------- ---------- ----------
5246016       trinotate          1        20G
5246016.bat+      batch          1              3778500K
5246016.ext+     extern          1                     0

```

Only 3.7 Gb max used and 7.5 hours needed.


```{bash, eval=FALSE}

# Bash
srun --pty -p medium --mem=4G --time=01:00:00  bash

module load apptainer


NAME_SPECIES="Anchomenus_dorsalis"

PATH_DIR="/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects"

# Variables (I moved the folder TRINOTATE_DATA_DIR locally for the run to avoid path mess with apptainer)
# Was originally in /usr/users/zicola/bin/Trinotate-Trinotate-v4.0.1
export TRINOTATE_DATA_DIR=/scratch/users/zicola/TRINOTATE_DATA_DIR

export TRINOTATE_HOME=/usr/local/src/Trinotate

#NAME_SPECIES is provided as argument to sbatch (Do not put a uppercase to trinotate_ suffix, it won't work)
sqlite_db="${PATH_DIR}/trinotate_output/${NAME_SPECIES}/database_${NAME_SPECIES}.sqlite"

apptainer exec --bind $PATH_DIR,$TRINOTATE_DATA_DIR -e ~/bin/trinotate.v4.0.2.simg $TRINOTATE_HOME/Trinotate \
   --db ${sqlite_db} \
   --report \
   --incl_pep \
   --incl_trans > ${PATH_DIR}/trinotate_output/${NAME_SPECIES}/Trinotate_report.tsv

```



```{bash, eval=FALSE}

# Try with local Trinity but it does not work because DBI perl module is not installed
~/bin/Trinotate-Trinotate-v4.0.1/Trinotate   --db ${sqlite_db} \
   --report \
   --incl_pep \
   --incl_trans > ${PATH_DIR}/trinotate_output/${NAME_SPECIES}/Trinotate_report.tsv

```


I looked at the database with a software (DB Browse for SQLite) and it seems some tables are empty. Signalp and tmhmm are properly filled.



```{bash, eval=FALSE}


work_path="/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects"

i="Anchomenus_dorsalis"
NAME_SPECIES=$i
GENE_TRANS_MAP=${work_path}/trinity_output/summary/${i}_trinity.fasta.gene_trans_map
TRINITY_FASTA=${work_path}/trinity_output/summary/${i}_trinity.fasta
TRANSDECODER_PEP=${work_path}/transdecoder_output/${i}_trinity.fasta.transdecoder.pep 
PATH_DIR="/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects"
sqlite_db="${PATH_DIR}/trinotate_output/${NAME_SPECIES}/database_${NAME_SPECIES}.sqlite"

apptainer exec --bind $PATH_DIR -e ~/bin/trinotate.v4.0.2.simg $TRINOTATE_HOME/Trinotate \
   --db $sqlite_db --init \
   --gene_trans_map $GENE_TRANS_MAP \
   --transcript_fasta $TRINITY_FASTA \
   --transdecoder_pep $TRANSDECODER_PEP
```


```{bash, eval=FALSE}

apptainer exec --bind $PATH_DIR,$TRINOTATE_DATA_DIR -e ~/bin/trinotate.v4.0.2.simg $TRINOTATE_HOME/Trinotate \
   --db ${sqlite_db} \
   --LOAD_EggnogMapper 


```



```{bash, eval=FALSE}

# Remove OK check point, otherwise the DB not created
rm /scratch/users/zicola/TRINOTATE_DATA_DIR/__chckpts/cp_trinotate_boilerplate_sqlite_*

apptainer exec --bind $PATH_DIR,$TRINOTATE_DATA_DIR -e ~/bin/trinotate.v4.0.2.simg $TRINOTATE_HOME/Trinotate \
   --db $NAME_SPECIES/database_${NAME_SPECIES}.sqlite \
   --create \
   --trinotate_data_dir $TRINOTATE_DATA_DIR \
   --use_diamond
   
# Works

# Delete the __init.ok log file (otherwise, nothing happens)
rm /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinotate_output/Anchomenus_dorsalis/__init.ok

apptainer exec --bind $PATH_DIR,$TRINOTATE_DATA_DIR -e ~/bin/trinotate.v4.0.2.simg $TRINOTATE_HOME/Trinotate \
   --db $sqlite_db --init \
   --gene_trans_map $GENE_TRANS_MAP \
   --transcript_fasta $TRINITY_FASTA \
   --transdecoder_pep $TRANSDECODER_PEP

# Worked, no 504 Mb instead of 340 Mb

# --LOAD_pfam
apptainer exec --bind $PATH_DIR,$TRINOTATE_DATA_DIR -e ~/bin/trinotate.v4.0.2.simg $TRINOTATE_HOME/Trinotate \
--db ${sqlite_db} --LOAD_pfam  /usr/users/zicola/TrinotatePFAM.out

# --LOAD_swissprot_blastp
apptainer exec --bind $PATH_DIR,$TRINOTATE_DATA_DIR -e ~/bin/trinotate.v4.0.2.simg $TRINOTATE_HOME/Trinotate \
--db ${sqlite_db} --LOAD_swissprot_blastp /usr/users/zicola/uniprot_sprot.diamond.blastp.outfmt6

# --LOAD_swissprot_blastx
apptainer exec --bind $PATH_DIR,$TRINOTATE_DATA_DIR -e ~/bin/trinotate.v4.0.2.simg $TRINOTATE_HOME/Trinotate \
--db ${sqlite_db} --LOAD_swissprot_blastx /usr/users/zicola/uniprot_sprot.diamond.blastx.outfmt6

# LOAD_EggnogMapper
apptainer exec --bind $PATH_DIR,$TRINOTATE_DATA_DIR -e ~/bin/trinotate.v4.0.2.simg $TRINOTATE_HOME/Trinotate \
--db ${sqlite_db} --LOAD_EggnogMapper /usr/users/zicola/eggnog_mapper.emapper.annotations

# --LOAD_signalp
apptainer exec --bind $PATH_DIR,$TRINOTATE_DATA_DIR -e ~/bin/trinotate.v4.0.2.simg $TRINOTATE_HOME/Trinotate \
   --db $sqlite_db \
   --LOAD_signalp ${PATH_DIR}/trinotate_output/${NAME_SPECIES}/sigP6outdir/output.gff3
   
# --LOAD_tmhmmv2
apptainer exec --bind $PATH_DIR,$TRINOTATE_DATA_DIR -e ~/bin/trinotate.v4.0.2.simg $TRINOTATE_HOME/Trinotate \
   --db $sqlite_db \
   --LOAD_tmhmmv2 ${PATH_DIR}/trinotate_output/${NAME_SPECIES}/tmhmm.v2.out
   
# --LOAD_infernal
apptainer exec --bind $PATH_DIR,$TRINOTATE_DATA_DIR -e ~/bin/trinotate.v4.0.2.simg $TRINOTATE_HOME/Trinotate \
   --db $sqlite_db \
   --LOAD_infernal /usr/users/zicola/infernal.out

# Get report
apptainer exec --bind $PATH_DIR,$TRINOTATE_DATA_DIR -e ~/bin/trinotate.v4.0.2.simg $TRINOTATE_HOME/Trinotate --db ${sqlite_db} \
   --report \
   --incl_pep \
   --incl_trans > ${PATH_DIR}/trinotate_output/${NAME_SPECIES}/Trinotate_report.tsv
   
# Extract GO terms

apptainer exec --bind $PATH_DIR -e ~/bin/trinotate.v4.0.2.simg \
   $TRINOTATE_HOME/util/extract_GO_assignments_from_Trinotate_xls.pl \
  --Trinotate_xls Trinotate_report.tsv \
  -G -I > ${PATH_DIR}/trinotate_output/${NAME_SPECIES}/Trinotate_report.tsv.gene_ontology

# Generate trinotate report summary statistics
apptainer exec --bind $PATH_DIR -e ~/bin/trinotate.v4.0.2.simg \
   $TRINOTATE_HOME/util/report_summary/trinotate_report_summary.pl \
  ${PATH_DIR}/trinotate_output/${NAME_SPECIES}/Trinotate_report.tsv \
  ${PATH_DIR}/trinotate_output/${NAME_SPECIES}/Trinotate_report_stats
   
```

Worked!!!

```{bash, eval=FALSE}
ll
total 851M
drwxr-x---  3 zicola UANP 4.1K Jan 25 21:12 ./
drwxr-x--- 10 zicola UANP    0 Jan 24 10:40 ../
-rw-r-----  1 zicola UANP 176M Jan 25 21:08 Trinotate_report.tsv
-rw-r-----  1 zicola UANP 4.7M Jan 25 21:12 Trinotate_report.tsv.gene_ontology
-rw-r-----  1 zicola UANP 4.7M Jan 25 21:12 Trinotate_report_stats.GO
-rw-r-----  1 zicola UANP  47K Jan 25 21:12 Trinotate_report_stats.GO.slim
-rw-r-----  1 zicola UANP 333K Jan 25 21:12 Trinotate_report_stats.cXp_summary.html
-rw-r-----  1 zicola UANP 2.6K Jan 25 21:12 Trinotate_report_stats.eggnog_counts
-rw-r-----  1 zicola UANP  527 Jan 25 21:12 Trinotate_report_stats.eggnog_counts.funcats
-rw-r-----  1 zicola UANP  87K Jan 25 21:12 Trinotate_report_stats.kegg.counts
-rw-r-----  1 zicola UANP  80K Jan 25 21:12 Trinotate_report_stats.pfam.counts
-rw-r-----  1 zicola UANP 3.2K Jan 25 21:12 Trinotate_report_stats.species_counts
-rw-r-----  1 zicola UANP 4.4K Jan 25 21:12 Trinotate_report_stats.taxonomy_counts
-rw-r-----  1 zicola UANP    0 Jan 25 20:42 __init.ok
-rw-r-----  1 zicola UANP 633M Jan 25 20:59 database_Anchomenus_dorsalis.sqlite
drwxr-x---  2 zicola UANP    0 Jan 24 17:48 sigP6outdir/
-rw-r-----  1 zicola UANP 3.7M Jan 24 18:23 tmhmm.v2.out
-rw-r-----  1 zicola UANP    0 Jan 25 20:52 tmp.emapper_bulk_load.1140688
-rw-r-----  1 zicola UANP    0 Jan 25 20:55 tmp.signalp_bulk_load.1152221
-rw-r-----  1 zicola UANP  27K Jan 23 22:00 trinotate.5222088.err
-rw-r-----  1 zicola UANP 3.2K Jan 23 22:00 trinotate.5222088.out
-rw-r-----  1 zicola UANP 386K Jan 24 18:23 trinotate.5246016.err
-rw-r-----  1 zicola UANP  969 Jan 24 18:23 trinotate.5246016.out



```

the `database_Anchomenus_dorsalis.sqlite` is 633 Mb at the end.

### sbatch_trinotate.sh

```{bash, eval=FALSE}
#!/bin/bash
#SBATCH --job-name=04_trinotate
#SBATCH --nodes=1
#SBATCH --partition=medium
#SBATCH --time=16:00:00
#SBATCH --cpus-per-task=10
#SBATCH --mem=12G
#SBATCH --mail-type=END
#SBATCH --mail-user=johan.zicola@uni-goettingen.de
#SBATCH --output=04_trinotate.%J.out
#SBATCH --error=04_trinotate.%J.err

mem_gbytes=$(( $SLURM_MEM_PER_NODE/1024 ))

echo "Starting at $(date)"
echo "Job submitted to the ${SLURM_JOB_PARTITION} partition, the default partition on ${SLURM_CLUSTER_NAME}"
echo "Job name: ${SLURM_JOB_NAME}, Job ID: ${SLURM_JOB_ID}"
echo "I have ${SLURM_CPUS_ON_NODE} CPUs and ${mem_gbytes}GiB of RAM on compute node $(hostname)"

# Modules to load
module purge
module load apptainer

# Create newhome
mkdir /usr/users/zicola/trinotate_${SPECIES}

export APPTAINERENV_NEWHOME=/usr/users/zicola/trinotate_${SPECIES}
export APPTAINERENV_SPECIES=$SPECIES

apptainer run -C -B /usr/users/npghpc/ \
   -B /usr/users/zicola -B /scratch/users/zicola/ /usr/users/zicola/bin/trinotate.v4.0.2.simg /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/04_trinotate.sh

```

Launch sbatch with the species argument.

```{bash, eval=FALSE}
sbatch --export=SPECIES="Amara_similata" sbatch_trinotate.sh
```

Works, now, all Trinotate output will go in `~/newhome` instead of `$HOME`.

```{bash, eval=FALSE}
ll ~/newhome/
total 51M
drwxr-x---  3 zicola UANP    6 Jan 27 19:36 ./
drwx------ 63 zicola UANP 9.3K Jan 27 19:33 ../
-rw-r-----  1 zicola UANP  20M Jan 27 19:40 TrinotatePFAM.out
-rw-r-----  1 zicola UANP    0 Jan 27 19:36 __init.ok
drwxr-x---  2 zicola UANP    0 Jan 27 19:36 __trinotate_run_checkpts/
-rw-r-----  1 zicola UANP  22M Jan 27 19:40 pfam.log
-rw-r-----  1 zicola UANP 1.6M Jan 27 19:36 uniprot_sprot.diamond.blastp.outfmt6
-rw-r-----  1 zicola UANP 1.6M Jan 27 19:36 uniprot_sprot.diamond.blastx.outfmt6
```


```{bash, eval=FALSE}
sacct -j JOBID -o jobid,jobname,AllocCPUS,ReqMem,MaxRSS,Elapsed

sacct -j 5411597  -o jobid,jobname,AllocCPUS,ReqMem,MaxRSS,Elapsed
JobID           JobName  AllocCPUS     ReqMem     MaxRSS    Elapsed
------------ ---------- ---------- ---------- ---------- ----------
5411597       trinotate         10        12G              12:43:41
5411597.bat+      batch         10              8463548K   12:43:41
5411597.ext+     extern         10                     0   12:43:41

```

12h40 needed.

```{bash, eval=FALSE}
ls ~/newhome/
total 475232
-rw-r----- 1 zicola UANP 124819074 Jan 27 20:11 TrinotatePFAM.out
-rw-r----- 1 zicola UANP         0 Jan 27 19:36 __init.ok
drwxr-x--- 2 zicola UANP         2 Jan 28 08:19 __trinotate_run_checkpts
-rw-r----- 1 zicola UANP  44173297 Jan 28 08:19 eggnog_mapper.emapper.annotations
-rw-r----- 1 zicola UANP   9483642 Jan 28 02:27 eggnog_mapper.emapper.hits
-rw-r----- 1 zicola UANP   2365038 Jan 28 08:19 eggnog_mapper.emapper.seed_orthologs
-rw-r----- 1 zicola UANP 135214516 Jan 28 02:16 infernal.log
-rw-r----- 1 zicola UANP     80487 Jan 28 02:16 infernal.out
-rw-r----- 1 zicola UANP 140305066 Jan 27 20:11 pfam.log
-rw-r----- 1 zicola UANP   1583830 Jan 27 19:36 uniprot_sprot.diamond.blastp.outfmt6
-rw-r----- 1 zicola UANP   1673060 Jan 27 19:36 uniprot_sprot.diamond.blastx.outfmt6

```

```{bash, eval=FALSE}
cd /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects

587M Jan 28 08:19 database_Amara_similata.sqlite

```
