---
title: "Untitled"
author: "Johan Zicola"
date: '2025-01-06'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Objective

Di and Kathi proceeded to the RNA extraction and mRNA-seq library preparation of 8 different insect species. The libraries were sequenced in 100+100 PE mode by BGI.

Insects collected and identified by Di:

+ Anchomenus dorsalis
+ Pterostichus melanarius
+ Zabrus tenebrioides
+ Amara similata
+ Harpalus affinis
+ Harpalus eous
+ Poecilus cupreus
+ Aphidius rhopalosiphi

Besides Aphidius rhopalosiphi (parasitic wasp, Hymenoptera), all other 7 species are Coleoptera.

The goal is to compare transcripts of different species and see if they are targeted by the designed dsRNAs used to kill the CSFB.

* Assemble the transcripts for each species
* Get the full list of transcripts and the list of longest ORFs




Off-target analysis done by the company:

* Map the designed dsRNAs (50 or so designed) to these transcripts with allowed mismatches (do 0, 1, and 2)
* Get a list of candidate target transcripts

Check Cedden 2024 (https://resjournals.onlinelibrary.wiley.com/doi/10.1111/imb.12942) for the dsRNA design and off-target analysis. He developped an online tool to design dsRNAs (https://dsrip.uni-goettingen.de/efficiency).

## Data

Data are in `/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects`

Summary statistics for raw data:

| Sample | Length(bp) | TotalReads            |
|--------|------------|-----------------------|
| 01_Ad  | 100;100    |           64,769,548  |
| 02_Pm  | 100;100    |           56,305,457  |
| 03_Zt  | 100;100    |           50,054,875  |
| 04_Asp | 100;100    |           42,581,758  |
| 05_G   | 100;100    |           41,668,796  |
| 06_Ol  | 100;100    |           57,482,693  |
| 07_Pr  | 100;100    |           47,960,411  |
| 08_Ar  | 100;100    |           44,601,099  |

We have between 44 and 64 M paired-end reads.

fastqc

```{bash, eval=FALSE}
f *fq.gz > list_fastq.txt

while read i; do srun --partition medium fastqc $i -o fastqc_reports/ & done < list_fastq.txt
```


# Pipeline

https://southgreenplatform.github.io/tutorials//bioanalysis/trinity/#trimmomatic

* Trim data with cutadapt to remove adapters

## Trimming

Note that Trinity will skip reads of less than 25 bp so I can set the minimum size to 25 nt.

Remove 3' adapters from read 1 and read 2 and trim by quality, allowing no Ns.

```{bash, eval=FALSE}

#!/bin/bash
#
#SBATCH --job-name=trimming_cutadapt
#SBATCH --nodes=1
#SBATCH --ntasks=1 
#SBATCH --mem=16000
#SBATCH --output=slurm.%A.%a.out
#SBATCH --error=slurm.%A.%a.err
#SBATCH --partition=medium
#SBATCH --time=02:00:00
#SBATCH --constraint=scratch
#SBATCH --mail-type=END
#SBATCH --mail-user=johan.zicola@uni-goettingen.de
#SBATCH --array=1-8

module load cutadapt/4.4

file_name=$(find . -type f -name "*fq.gz" | \
cut -d_ -f1-4 | sort | uniq | sed -n ${SLURM_ARRAY_TASK_ID}p)

cutadapt -a AGATCGGAAGAGCACACGTCTGAACTCCAGTCA \
    -A AGATCGGAAGAGCGTCGTGTAGGGAAAGAGTGT \
    --quality-base 33 --max-n 0 --minimum-length 25 -o ${file_name}_1.trimmed.fq \
    -p ${file_name}_2.trimmed.fq ${file_name}_1.fq.gz ${file_name}_2.fq.gz
    
```

The Ar sample shows a polyA adaptor contamination that is gradual from the beginning to the end of the read. It also contains a lower GC content than all other samples. 

Number of reads after trimming:

```{bash, eval=FALSE}
for i in *fq; do srun --partition medium fastqc $i -o fastqc_output/ & done

cut -f1,6 multiqc_general_stats.txt
```


| Sample      | TotalReads  |
|-------------|-------------|
| 01_Ad_L1_1  | 63,577,928  |
| 01_Ad_L1_2  | 63,577,928  |
| 02_Pm_L1_1  | 55,269,387  |
| 02_Pm_L1_2  | 55,269,387  |
| 03_Zt_L1_1  | 49,128,635  |
| 03_Zt_L1_2  | 49,128,635  |
| 04_Asp_L1_1 | 41,800,974  |
| 04_Asp_L1_2 | 41,800,974  |
| 05_G_L1_1   | 40,906,340  |
| 05_G_L1_2   | 40,906,340  |
| 06_Ol_L1_1  | 56,417,880  |
| 06_Ol_L1_2  | 56,417,880  |
| 07_Pr_L1_1  | 47,070,196  |
| 07_Pr_L1_2  | 47,070,196  |
| 08_Ar_L1_1  | 43,759,260  |
| 08_Ar_L1_2  | 43,759,260  |




# Remove empty reads

```{bash, eval=FALSE}

while read i; do
  srun -p medium -t 01:20:00 --cpus-per-task=10 cutadapt -j 10 --minimum-length 25 \
  -o ${i}_L1_1.trimmed.min25.fq -p ${i}_L1_2.trimmed.min25.fq \
  ${i}_L1_1.trimmed.fq ${i}_L1_2.trimmed.fq &
done < list.txt

```




# Trimming with trimmomatic

```{bash, eval=FALSE}

module load rev/21.12 trimmomatic openjdk

PATH_ADAPTER="/usr/product/bioinfo/SL_7.0/BIOINFORMATICS/TRIMMOMATIC/0.36/adapters/"

srun -p medium -t 00:30:00 --mem=12G --cpus-per-task=10 \
java -Xmx10G -jar /usr/product/bioinfo/SL_7.0/BIOINFORMATICS/TRIMMOMATIC/0.36/trimmomatic-0.36.jar \
PE -phred33 -threads 10 \
raw_data/01_Ad/01_Ad_L1_1.fq.gz \
raw_data/01_Ad/01_Ad_L1_2.fq.gz \
trimmed_reads_trimmomatic/01_Ad_L1_1.paired.fq.gz \
trimmed_reads_trimmomatic/01_Ad_L1_1.unpaired.fq.gz \
trimmed_reads_trimmomatic/01_Ad_L1_2.paired.fq.gz \
trimmed_reads_trimmomatic/01_Ad_L1_2.unpaired.fq.gz \
ILLUMINACLIP:"/usr/product/bioinfo/SL_7.0/BIOINFORMATICS/TRIMMOMATIC/0.36/adapters/TruSeq2-SE.fa":2:30:10 LEADING:3 TRAILING:3 SLIDINGWINDOW:4:15 MINLEN:50 &

Input Read Pairs: 64769548 Both Surviving: 63780516 (98.47%) Forward Only Surviving: 228552 (0.35%) Reverse Only Surviving: 744814 (1.15%) Dropped: 15666 (0.02%)
TrimmomaticPE: Completed successfully

seqkit stats 01_Ad_L1_1.paired.fq.gz 01_Ad_L1_2.paired.fq.gz
processed files:  2 / 2 [======================================] ETA: 0s. done
file                     format  type    num_seqs        sum_len  min_len  avg_len  max_len
01_Ad_L1_1.paired.fq.gz  FASTQ   DNA   63,780,516  6,351,162,309       50     99.6      100
01_Ad_L1_2.paired.fq.gz  FASTQ   DNA   63,780,516  6,350,850,563       50     99.6      100

# Same number of reads in each file

module load fastqc/0.11.4

srun -p medium -t 00:50:00 fastqc 01_Ad_L1_2.paired.fq.gz &
srun -p medium -t 00:50:00 fastqc 01_Ad_L1_1.paired.fq.gz &

```


# Remove rRNA

https://github.com/sortmerna/sortmerna

Download v4.3 database (https://github.com/biocore/sortmerna/releases/download/v4.3.4/database.tar.gz)

### Index database

The docu recommends to use `smr_v4.3_default_db.fasta`

```{r}
mkdir ~/genomes/silva_rRNA_maize/smr_database && cd ~/genomes/silva_rRNA_maize/smr_database

wget https://github.com/biocore/sortmerna/releases/download/v4.3.4/database.tar.gz

tar -zxvf database.tar.gz

module load rev/21.12 sortmerna/20170809

srun -t 00:50:00 --mem=20G -p medium indexdb_rna --ref smr_v4.3_default_db.fasta,smr_v4.3_default_db.idx -v

```


```{bash, eval=FALSE}
\ls -1sh smr_v4.3_default*
140M smr_v4.3_default_db.fasta
662M smr_v4.3_default_db.idx.bursttrie_0.dat
1.0M smr_v4.3_default_db.idx.kmer_0.dat
1.2G smr_v4.3_default_db.idx.pos_0.dat
4.0M smr_v4.3_default_db.idx.stats
```

4 index files are generated with a *dat extension.

### Merge reads

I tested with the R1 and R2 in separate files using two --read flags but the output contained only R2 sequences. When I merge the R1 and R2, it works. I just need to merge R1 and R2 and unmerge them afterwards

```{bash, eval=FALSE}

cd /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trimmed_reads

mkdir merged_fastq

module load rev/21.12 sortmerna/20170809

while read i; do
if [ -e merged_fastq/${i}.merged.fq ]; then
  echo "$i already processed"
else
  echo "srun -p medium --mem=2Gb --time=00:20:00 bash /usr/product/bioinfo/SL_7.0/BIOINFORMATICS/SORTMERNA/sortmerna_20170809/scripts/merge-paired-reads.sh ${i}_L1_1.trimmed.fq ${i}_L1_2.trimmed.fq  merged_fastq/${i}.merged.fq" && srun -p medium --mem=2Gb --time=00:20:00 bash /usr/product/bioinfo/SL_7.0/BIOINFORMATICS/SORTMERNA/sortmerna_20170809/scripts/merge-paired-reads.sh ${i}_L1_1.trimmed.fq ${i}_L1_2.trimmed.fq  merged_fastq/${i}.merged.fq &
fi
done < list.txt


```


### sortmerna

```{r}
#!/bin/bash
#SBATCH --job-name=sortmerna
#SBATCH --nodes=1
#SBATCH --time=01:00:00
#SBATCH --cpus-per-task=10
#SBATCH --mem=10gb
#SBATCH --partition=medium
#SBATCH --mail-type=END
#SBATCH --mail-user=johan.zicola@uni-goettingen.de
#SBATCH --output=sortmerna.%J.out
#SBATCH --error=sortmerna.%J.err
#SBATCH --array=1-8

mem_gbytes=$(( $SLURM_MEM_PER_NODE/1024 ))

echo "Starting at $(date)"
echo "Job submitted to the ${SLURM_JOB_PARTITION} partition, the default partition on ${SLURM_CLUSTER_NAME}"
echo "Job name: ${SLURM_JOB_NAME}, Job ID: ${SLURM_JOB_ID}"
echo "I have ${SLURM_CPUS_ON_NODE} CPUs and ${mem_gbytes}GiB of RAM on compute node $(hostname)"

module purge

module load rev/21.12 sortmerna/20170809

PATH_DB="/usr/users/zicola/genomes/silva_rRNA_maize/smr_database"

i=$(sed -n ${SLURM_ARRAY_TASK_ID}p trimmed_reads/list.txt)

sortmerna -a 1 --paired_in --fastx \
    --ref ${PATH_DB}/smr_v4.3_default_db.fasta,${PATH_DB}/smr_v4.3_default_db.idx \
    --reads trimmed_reads/merged_fastq/${i}.merged.fq \
    --aligned rRNA_removed/${i}.aligned \
    --other rRNA_removed/${i}.mRNA

```

```{bash, eval=FALSE}

srun --pty -p medium --mem=4G --cpus-per-task=10 --time=01:00:00  bash

i="02_Pm"

module purge

module load rev/21.12 sortmerna/20170809

PATH_DB="/usr/users/zicola/genomes/silva_rRNA_maize/smr_database"

sortmerna -a 10 --paired_in --fastx \
    --ref ${PATH_DB}/smr_v4.3_default_db.fasta,${PATH_DB}/smr_v4.3_default_db.idx \
    --reads trimmed_reads/merged_fastq/${i}.merged.fq \
    --aligned rRNA_removed/${i}.aligned \
    --other rRNA_removed/${i}.mRNA

```




## Trinity (v2.15.1) with apptainer

```{bash, eval=FALSE}
#!/bin/bash
#SBATCH --job-name=Trinity
#SBATCH --nodes=1
#SBATCH --time=12:00:00
#SBATCH --cpus-per-task=8
#SBATCH --mem=110gb
#SBATCH --partition=medium
#SBATCH --mail-type=END
#SBATCH --mail-user=johan.zicola@uni-goettingen.de
#SBATCH --output=Trinity.%J.out
#SBATCH --error=Trinity.%J.err

mem_gbytes=$(( $SLURM_MEM_PER_NODE/1024 ))

echo "Starting at $(date)"
echo "Job submitted to the ${SLURM_JOB_PARTITION} partition, the default partition on ${SLURM_CLUSTER_NAME}"
echo "Job name: ${SLURM_JOB_NAME}, Job ID: ${SLURM_JOB_ID}"
echo "I have ${SLURM_CPUS_ON_NODE} CPUs and ${mem_gbytes}GiB of RAM on compute node $(hostname)"

module purge

module load apptainer

PATH_DIR="/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects"

apptainer exec --bind $PATH_DIR -e /usr/users/zicola/bin/trinityrnaseq.v2.15.1.simg  Trinity \
          --seqType fq \
          --SS_lib_type RF \
          --left ${PATH_DIR}/$INPUT_FASTQ_LEFT  \
          --right ${PATH_DIR}/$INPUT_FASTQ_RIGHT \
          --max_memory 100G --CPU $SLURM_CPUS_PER_TASK \
          --output ${PATH_DIR}/$OUTPUT_DIR

```


```{r}
# Ad
sbatch  --export=INPUT_FASTQ_LEFT=trimmed_reads/01_Ad_L1_1.trimmed.fq,INPUT_FASTQ_RIGHT=trimmed_reads/01_Ad_L1_2.trimmed.fq,OUTPUT_DIR=trinity_output/Ad/trinity_out_dir trinity_run.sh
```

First run finished. It took 9 hours (08:58:28)

```{bash, eval=FALSE}
seqkit stats trinity_out_dir.Trinity.fasta
file                           format  type  num_seqs      sum_len  min_len  avg_len  max_len
trinity_out_dir.Trinity.fasta  FASTA   DNA    103,137  110,493,798      184  1,071.3   33,453

# Number of genes
cut -f1 trinity_out_dir.Trinity.fasta.gene_trans_map | uniq | wc -l
67047

# Rename fasta
mv trinity_out_dir.Trinity.fasta Ad_trinity.Trinity.fasta
mv trinity_out_dir.Trinity.fasta.gene_trans_map Ad_trinity.Trinity.fasta.gene_trans_map

```




```{r}
# Pm
sbatch  --export=INPUT_FASTQ_LEFT=trimmed_reads/01_Pm_L1_1.trimmed.fq,INPUT_FASTQ_RIGHT=trimmed_reads/01_Pm_L1_2.trimmed.fq,OUTPUT_DIR=trinity_output/Pm/Pm_trinity trinity_run.sh
```



```{bash, eval=FALSE}


# Pm
sbatch  --export=INPUT_FASTQ_LEFT=trimmed_reads/01_Ad_L1_1.trimmed.fq,INPUT_FASTQ_RIGHT=trimmed_reads/01_Ad_L1_2.trimmed.fq,OUTPUT_DIR=trinity_output/Pm trinity_run.sh

# Zt
sbatch  --export=INPUT_FASTQ_LEFT=trimmed_reads/01_Ad_L1_1.trimmed.fq,INPUT_FASTQ_RIGHT=trimmed_reads/01_Ad_L1_2.trimmed.fq,OUTPUT_DIR=trinity_output/Zt trinity_run.sh

# Asp
sbatch  --export=INPUT_FASTQ_LEFT=trimmed_reads/01_Ad_L1_1.trimmed.fq,INPUT_FASTQ_RIGHT=trimmed_reads/01_Ad_L1_2.trimmed.fq,OUTPUT_DIR=trinity_output/Asp trinity_run.sh

# G
sbatch  --export=INPUT_FASTQ_LEFT=trimmed_reads/01_Ad_L1_1.trimmed.fq,INPUT_FASTQ_RIGHT=trimmed_reads/01_Ad_L1_2.trimmed.fq,OUTPUT_DIR=trinity_output/G trinity_run.sh

# Ol
sbatch  --export=INPUT_FASTQ_LEFT=trimmed_reads/01_Ad_L1_1.trimmed.fq,INPUT_FASTQ_RIGHT=trimmed_reads/01_Ad_L1_2.trimmed.fq,OUTPUT_DIR=trinity_output/Ol trinity_run.sh

# Pr
sbatch  --export=INPUT_FASTQ_LEFT=trimmed_reads/01_Ad_L1_1.trimmed.fq,INPUT_FASTQ_RIGHT=trimmed_reads/01_Ad_L1_2.trimmed.fq,OUTPUT_DIR=trinity_output/Pr trinity_run.sh

# Ar
sbatch  --export=INPUT_FASTQ_LEFT=trimmed_reads/01_Ad_L1_1.trimmed.fq,INPUT_FASTQ_RIGHT=trimmed_reads/01_Ad_L1_2.trimmed.fq,OUTPUT_DIR=trinity_output/Ar trinity_run.sh


```





# Off-target analysis

Use Doga's script (https://figshare.com/articles/dataset/_i_in_silico_i_off-target_analysis_of_lethal_genes_in_cabbage_stem_flea_beetle/25283683?file=44671225) (scripts/off_target_prediction_lethal.py)

# COI1 validation


Di mentioned that to verify if we have the correct species, one needs to look at the cytochrome oxidase subunit 1 (COI) sequence.

Anchomenus dorsalis
https://www.ncbi.nlm.nih.gov/nuccore/HQ953549.1

Pterostichus melanarius
https://www.ncbi.nlm.nih.gov/nuccore/MH020387.1

Zabrus tenebrioides
https://www.ncbi.nlm.nih.gov/nuccore/ON623806.2

Amara similata
https://www.ncbi.nlm.nih.gov/nuccore/MH020396.1

Harpalus affinis
https://www.ncbi.nlm.nih.gov/nuccore/OK638137.1

Harpalus eous
https://www.ncbi.nlm.nih.gov/nuccore/LC764887.1

Poecilus cupreus
https://www.ncbi.nlm.nih.gov/nuccore/MN454673.1

Aphidius rhopalosiphi
https://www.ncbi.nlm.nih.gov/nuccore/JN164779.1


