---
title: "Untitled"
author: "Johan Zicola"
output:
  html_notebook:
    toc: yes
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Objective

Di and Kathi proceeded to the RNA extraction and mRNA-seq library preparation of 8 different insect species. The libraries were sequenced in 100+100 PE mode by BGI.



Transcriptome assemblies of eight non-target insect species present in autumn rapeseed fields 

# Insect species

Insects collected and identified by Di:

| library | species                  | basionym             | order       | NCBI taxonomy ID |
|---------|--------------------------|----------------------|-------------|------------------|
| 01_Ad   | Anchomenus dorsalis      |                      | Coleoptera  | 586010           |
| 02_Pm   | Pterostichus melanarius  |                      | Coleoptera  | 60768            |
| 03_Zt   | Zabrus tenebrioides      | Carabus tenebrioides | Coleoptera  | 272042           |
| 04_Asp  | Amara similata           | Harpalus similatus   | Coleoptera  | 272013           |
| 05_G    | Harpalus affinis         |                      | Coleoptera  | 247283           |
| 06_Ol   | Harpalus eous            |                      | Coleoptera  | 2107211          |
| 07_Pr   | Poecilus cupreus         | Pterostichus cupreus | Coleoptera  | 270614           |
| 08_Ar   | Aphidius rhopalosiphi    |                      | Hymenoptera | 55891            |

+ morphology and species attributes: https://www.gbif.org/species/search?q=
+ Full taxonomy: https://www.ncbi.nlm.nih.gov/Taxonomy/Browser/wwwtax.cgi

Besides Aphidius rhopalosiphi (parasitic wasp, Hymenoptera), all other 7 species are Coleoptera.

The goal is to compare transcripts of different species and see if they are targeted by the designed dsRNAs used to kill the CSFB.

* Assemble the transcripts for each species
* Get the full list of transcripts and the list of longest ORFs


Off-target analysis done by the company:

* Map the designed dsRNAs (50 or so designed) to these transcripts with allowed mismatches (do 0, 1, and 2)
* Get a list of candidate target transcripts

Check Cedden 2024 (https://resjournals.onlinelibrary.wiley.com/doi/10.1111/imb.12942) for the dsRNA design and off-target analysis. He developped an online tool to design dsRNAs (https://dsrip.uni-goettingen.de/efficiency).

# Data

Data are in `/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects`

Summary statistics for raw data:

| Sample | Length(bp) | TotalReads            |
|--------|------------|-----------------------|
| 01_Ad  | 100;100    |           64,769,548  |
| 02_Pm  | 100;100    |           56,305,457  |
| 03_Zt  | 100;100    |           50,054,875  |
| 04_Asp | 100;100    |           42,581,758  |
| 05_G   | 100;100    |           41,668,796  |
| 06_Ol  | 100;100    |           57,482,693  |
| 07_Pr  | 100;100    |           47,960,411  |
| 08_Ar  | 100;100    |           44,601,099  |

We have between 44 and 64 M paired-end reads.

fastqc

```{bash, eval=FALSE}
f *fq.gz > list_fastq.txt

while read i; do srun --partition medium fastqc $i -o fastqc_reports/ & done < list_fastq.txt
```


# Pipeline

https://southgreenplatform.github.io/tutorials//bioanalysis/trinity/

https://training.galaxyproject.org/topics/transcriptomics/tutorials/full-de-novo/tutorial.html

* Trim data with cutadapt to remove adapters

# Trimming

Note that Trinity will skip reads of less than 25 bp so I can set the minimum size to 25 nt.

Remove 3' adapters from read 1 and read 2 and trim by quality, allowing no Ns.

```{bash, eval=FALSE}

#!/bin/bash
#
#SBATCH --job-name=trimming_cutadapt
#SBATCH --nodes=1
#SBATCH --ntasks=1 
#SBATCH --mem=16000
#SBATCH --output=slurm.%A.%a.out
#SBATCH --error=slurm.%A.%a.err
#SBATCH --partition=medium
#SBATCH --time=02:00:00
#SBATCH --constraint=scratch
#SBATCH --mail-type=END
#SBATCH --mail-user=johan.zicola@uni-goettingen.de
#SBATCH --array=1-8


mem_gbytes=$(( $SLURM_MEM_PER_NODE/1024 ))

echo "Starting at $(date)"
echo "Job submitted to the ${SLURM_JOB_PARTITION} partition, the default partition on ${SLURM_CLUSTER_NAME}"
echo "Job name: ${SLURM_JOB_NAME}, Job ID: ${SLURM_JOB_ID}"
echo "I have ${SLURM_CPUS_ON_NODE} CPUs and ${mem_gbytes}GiB of RAM on compute node $(hostname)"

module load cutadapt/4.4

file_name=$(find . -type f -name "*fq.gz" | \
cut -d_ -f1-4 | sort | uniq | sed -n ${SLURM_ARRAY_TASK_ID}p)

cutadapt -a AGATCGGAAGAGCACACGTCTGAACTCCAGTCA \
    -A AGATCGGAAGAGCGTCGTGTAGGGAAAGAGTGT \
    --quality-base 33 --max-n 0 --minimum-length 25 -o ${file_name}_1.trimmed.fq \
    -p ${file_name}_2.trimmed.fq ${file_name}_1.fq.gz ${file_name}_2.fq.gz
    
```

The Ar sample shows a polyA adaptor contamination that is gradual from the beginning to the end of the read. It also contains a lower GC content than all other samples. Ar (Aphidius rhopalosiphi) is the only Hymenoptera among the 8 species (all others are Coleptera) so it may explain why it differs in GC content.

Number of reads after trimming:

```{bash, eval=FALSE}
for i in *fq; do srun --partition medium fastqc $i -o fastqc_output/ & done

cut -f1,6 multiqc_general_stats.txt
```


| Sample      | TotalReads  |
|-------------|-------------|
| 01_Ad_L1_1  | 63,577,928  |
| 01_Ad_L1_2  | 63,577,928  |
| 02_Pm_L1_1  | 55,269,387  |
| 02_Pm_L1_2  | 55,269,387  |
| 03_Zt_L1_1  | 49,128,635  |
| 03_Zt_L1_2  | 49,128,635  |
| 04_Asp_L1_1 | 41,800,974  |
| 04_Asp_L1_2 | 41,800,974  |
| 05_G_L1_1   | 40,906,340  |
| 05_G_L1_2   | 40,906,340  |
| 06_Ol_L1_1  | 56,417,880  |
| 06_Ol_L1_2  | 56,417,880  |
| 07_Pr_L1_1  | 47,070,196  |
| 07_Pr_L1_2  | 47,070,196  |
| 08_Ar_L1_1  | 43,759,260  |
| 08_Ar_L1_2  | 43,759,260  |


# Trinity (v2.15.1)

Use apptainer, the easiest solution to deal with all Trinity dependencies.

```{bash, eval=FALSE}
#!/bin/bash
#SBATCH --job-name=Trinity
#SBATCH --nodes=1
#SBATCH --time=12:00:00
#SBATCH --cpus-per-task=8
#SBATCH --mem=110gb
#SBATCH --partition=medium
#SBATCH --mail-type=END
#SBATCH --mail-user=johan.zicola@uni-goettingen.de
#SBATCH --output=Trinity.%J.out
#SBATCH --error=Trinity.%J.err

mem_gbytes=$(( $SLURM_MEM_PER_NODE/1024 ))

echo "Starting at $(date)"
echo "Job submitted to the ${SLURM_JOB_PARTITION} partition, the default partition on ${SLURM_CLUSTER_NAME}"
echo "Job name: ${SLURM_JOB_NAME}, Job ID: ${SLURM_JOB_ID}"
echo "I have ${SLURM_CPUS_ON_NODE} CPUs and ${mem_gbytes}GiB of RAM on compute node $(hostname)"

module purge

module load apptainer

PATH_DIR="/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects"

apptainer exec --bind $PATH_DIR -e /usr/users/zicola/bin/trinityrnaseq.v2.15.1.simg  Trinity \
          --max_memory 100G \
          --CPU $SLURM_CPUS_PER_TASK \
          --seqType fq \
          --SS_lib_type RF \
          --left ${PATH_DIR}/$INPUT_FASTQ_LEFT  \
          --right ${PATH_DIR}/$INPUT_FASTQ_RIGHT \
          --output ${PATH_DIR}/$OUTPUT_DIR

```

Loop of over each sample:

```{bash, eval=FALSE}

while read i; do
  echo "sbatch  --export=INPUT_FASTQ_LEFT=trimmed_reads/${i}_L1_1.trimmed.fq,INPUT_FASTQ_RIGHT=trimmed_reads/${i}_L1_2.trimmed.fq,OUTPUT_DIR=trinity_output/${i}/${i}_trinity trinity_run.sh"
  sbatch  --export=INPUT_FASTQ_LEFT=trimmed_reads/${i}_L1_1.trimmed.fq,INPUT_FASTQ_RIGHT=trimmed_reads/${i}_L1_2.trimmed.fq,OUTPUT_DIR=trinity_output/${i}/${i}_trinity trinity_run.sh
done < list

```


```{bash, eval=FALSE}
while read i; do seqkit stats $i; done <<<$(f *trinity.Trinity.fasta )
file                                 format  type  num_seqs      sum_len  min_len  avg_len  max_len
./02_Pm/02_Pm_trinity.Trinity.fasta  FASTA   DNA    140,145  116,247,368      177    829.5   26,406
file                               format  type  num_seqs      sum_len  min_len  avg_len  max_len
./05_G/05_G_trinity.Trinity.fasta  FASTA   DNA    104,390  100,387,029      177    961.7   25,018
file                                 format  type  num_seqs     sum_len  min_len  avg_len  max_len
./03_Zt/03_Zt_trinity.Trinity.fasta  FASTA   DNA     76,934  86,902,887      182  1,129.6   26,778
file                                   format  type  num_seqs     sum_len  min_len  avg_len  max_len
./04_Asp/04_Asp_trinity.Trinity.fasta  FASTA   DNA     72,481  82,598,703      174  1,139.6   24,631
file                                 format  type  num_seqs      sum_len  min_len  avg_len  max_len
./01_Ad/01_Ad_trinity.Trinity.fasta  FASTA   DNA    103,137  110,493,798      184  1,071.3   33,453

```



## 01_Ad

```{bash, eval=FALSE}
============ Job Information ==========================
JobID: 4909774
JobName: Trinity
Submitted: 2025-01-09T11:00:42
Started: 2025-01-09T11:01:32
Ended: 2025-01-09T20:00:00
Elapsed: 539 min, Limit: 2880 min, Difference: 2341 min
CPUs: 8, Nodes: 1
Estimated consumption: 143.73 core-hours
=======================================================
```


## 02_Pm

```{bash, eval=FALSE}
============ Job Information ==========================
JobID: 5037510
JobName: Trinity
Submitted: 2025-01-15T18:04:04
Started: 2025-01-15T18:30:58
Ended: 2025-01-16T05:29:00
Elapsed: 659 min, Limit: 720 min, Difference: 61 min
CPUs: 8, Nodes: 1
Estimated consumption: 175.73 core-hours
=======================================================

```

## 03_Zt

```{bash, eval=FALSE}
============ Job Information ==========================
JobID: 5037511
JobName: Trinity
Submitted: 2025-01-15T18:04:04
Started: 2025-01-15T18:33:29
Ended: 2025-01-16T02:21:09
Elapsed: 468 min, Limit: 720 min, Difference: 252 min
CPUs: 8, Nodes: 1
Estimated consumption: 124.80 core-hours
=======================================================
```


## 04_Asp

```{bash, eval=FALSE}
============ Job Information ==========================
JobID: 5037512
JobName: Trinity
Submitted: 2025-01-15T18:04:04
Started: 2025-01-15T18:43:26
Ended: 2025-01-16T01:05:36
Elapsed: 383 min, Limit: 720 min, Difference: 337 min
CPUs: 8, Nodes: 1
Estimated consumption: 102.13 core-hours
=======================================================

```


## 05_G

```{bash, eval=FALSE}
============ Job Information ==========================
JobID: 5037513
JobName: Trinity
Submitted: 2025-01-15T18:04:04
Started: 2025-01-16T01:05:50
Ended: 2025-01-16T09:00:03
Elapsed: 475 min, Limit: 720 min, Difference: 245 min
CPUs: 8, Nodes: 1
Estimated consumption: 126.67 core-hours
=======================================================
```


## 06_Ol

```{bash, eval=FALSE}
============ Job Information ==========================
JobID: 5037514
JobName: Trinity
Submitted: 2025-01-15T18:04:04
Started: 2025-01-16T09:00:47
Ended: 2025-01-16T16:55:44
Elapsed: 475 min, Limit: 720 min, Difference: 245 min
CPUs: 8, Nodes: 1
Estimated consumption: 126.67 core-hours
=======================================================
```

## 07_Pr

```{bash, eval=FALSE}

============ Job Information ==========================
JobID: 5037515
JobName: Trinity
Submitted: 2025-01-15T18:04:04
Started: 2025-01-16T09:51:54
Ended: 2025-01-16T17:41:27
Elapsed: 470 min, Limit: 720 min, Difference: 250 min
CPUs: 8, Nodes: 1
Estimated consumption: 125.33 core-hours
=======================================================
```


## 08_Ar

Time out

```{bash, eval=FALSE}
============ Job Information ==========================
JobID: 5037516
JobName: Trinity
Submitted: 2025-01-15T18:04:05
Started: 2025-01-16T09:51:54
Ended: 2025-01-16T21:52:17
Elapsed: 721 min, Limit: 720 min, Difference: -1 min
CPUs: 8, Nodes: 1
Estimated consumption: 192.27 core-hours
=======================================================
```

Based on the forum, I can rerun Trinity command and it will start from the last completed checkpoint


```{bash, eval=FALSE}
i="08_Ar"
sbatch  --export=INPUT_FASTQ_LEFT=trimmed_reads/${i}_L1_1.trimmed.fq,INPUT_FASTQ_RIGHT=trimmed_reads/${i}_L1_2.trimmed.fq,OUTPUT_DIR=trinity_output/${i}/${i}_trinity trinity_run.sh
```


```{bash, eval=FALSE}
============ Job Information ==========================
JobID: 5056408
JobName: Trinity
Submitted: 2025-01-17T08:47:18
Started: 2025-01-17T08:47:24
Ended: 2025-01-17T11:05:41
Elapsed: 139 min, Limit: 720 min, Difference: 581 min
CPUs: 8, Nodes: 1
Estimated consumption: 37.07 core-hours
=======================================================
```

Finished in 139 min for the remaining steps.


# Summary assemblies

```{bash, eval=FALSE}

f *_trinity.Trinity.fasta -exec seqkit stats {} \; | grep -v "file"
./02_Pm/02_Pm_trinity.Trinity.fasta  FASTA   DNA    140,145  116,247,368      177    829.5   26,406
./05_G/05_G_trinity.Trinity.fasta  FASTA   DNA    104,390  100,387,029      177    961.7   25,018
./03_Zt/03_Zt_trinity.Trinity.fasta  FASTA   DNA     76,934  86,902,887      182  1,129.6   26,778
./04_Asp/04_Asp_trinity.Trinity.fasta  FASTA   DNA     72,481  82,598,703      174  1,139.6   24,631
./06_Ol/06_Ol_trinity.Trinity.fasta  FASTA   DNA    100,481  96,388,352      178    959.3   20,450
./08_Ar/08_Ar_trinity.Trinity.fasta  FASTA   DNA    249,888  141,043,361      188    564.4   35,915
./07_Pr/07_Pr_trinity.Trinity.fasta  FASTA   DNA     92,572  108,290,220      180  1,169.8   24,578
./01_Ad/01_Ad_trinity.Trinity.fasta  FASTA   DNA    103,137  110,493,798      184  1,071.3   33,453


```


# Quality check TrinityStats

```{bash, eval=FALSE}

path_output="/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinity_output"

module load apptainer

apptainer exec --bind $path_output -e /usr/users/zicola/bin/trinityrnaseq.v2.15.1.simg /usr/local/bin/util/TrinityStats.pl 01_Ad/01_Ad_trinity.Trinity.fasta

```


## 01_Ad

```
################################
## Counts of transcripts, etc.
################################
Total trinity 'genes':  67047
Total trinity transcripts:   103137
Percent GC: 36.52

########################################
Stats based on ALL transcript contigs:
########################################

   Contig N10: 6347
   Contig N20: 4615
   Contig N30: 3602
   Contig N40: 2835
   Contig N50: 2225

   Median contig length: 460
   Average contig: 1071.33
   Total assembled bases: 110493798


#####################################################
## Stats based on ONLY LONGEST ISOFORM per 'GENE':
#####################################################

   Contig N10: 5485
   Contig N20: 3842
   Contig N30: 2821
   Contig N40: 2091
   Contig N50: 1507

   Median contig length: 348
   Average contig: 751.81
   Total assembled bases: 50406345

```


## 02_Pm


```
################################
## Counts of transcripts, etc.
################################
Total trinity 'genes':  94235
Total trinity transcripts:   140145
Percent GC: 36.70

########################################
Stats based on ALL transcript contigs:
########################################

   Contig N10: 5401
   Contig N20: 3962
   Contig N30: 2946
   Contig N40: 2241
   Contig N50: 1687

   Median contig length: 372
   Average contig: 829.48
   Total assembled bases: 116247368


#####################################################
## Stats based on ONLY LONGEST ISOFORM per 'GENE':
#####################################################

   Contig N10: 4755
   Contig N20: 3178
   Contig N30: 2225
   Contig N40: 1580
   Contig N50: 1027

   Median contig length: 318
   Average contig: 620.79
   Total assembled bases: 58500213

```

## 03_Zt

```
################################
## Counts of transcripts, etc.
################################
Total trinity 'genes':  51952
Total trinity transcripts:   76934
Percent GC: 36.81

########################################
Stats based on ALL transcript contigs:
########################################

   Contig N10: 6342
   Contig N20: 4654
   Contig N30: 3615
   Contig N40: 2885
   Contig N50: 2285

   Median contig length: 505
   Average contig: 1129.58
   Total assembled bases: 86902887


#####################################################
## Stats based on ONLY LONGEST ISOFORM per 'GENE':
#####################################################

   Contig N10: 5679
   Contig N20: 3983
   Contig N30: 3002
   Contig N40: 2313
   Contig N50: 1753

   Median contig length: 377
   Average contig: 837.57
   Total assembled bases: 43513583

```

## 04_Asp

```
################################
## Counts of transcripts, etc.
################################
Total trinity 'genes':  45431
Total trinity transcripts:   72481
Percent GC: 37.03

########################################
Stats based on ALL transcript contigs:
########################################

   Contig N10: 6191
   Contig N20: 4539
   Contig N30: 3567
   Contig N40: 2817
   Contig N50: 2233

   Median contig length: 533
   Average contig: 1139.59
   Total assembled bases: 82598703


#####################################################
## Stats based on ONLY LONGEST ISOFORM per 'GENE':
#####################################################

   Contig N10: 5577
   Contig N20: 3920
   Contig N30: 2963
   Contig N40: 2276
   Contig N50: 1750

   Median contig length: 390
   Average contig: 862.61
   Total assembled bases: 39189450

```

## 05_G

```
################################
## Counts of transcripts, etc.
################################
Total trinity 'genes':  63019
Total trinity transcripts:   104390
Percent GC: 36.59

########################################
Stats based on ALL transcript contigs:
########################################

   Contig N10: 5600
   Contig N20: 3984
   Contig N30: 3006
   Contig N40: 2372
   Contig N50: 1843

   Median contig length: 453
   Average contig: 961.65
   Total assembled bases: 100387029


#####################################################
## Stats based on ONLY LONGEST ISOFORM per 'GENE':
#####################################################

   Contig N10: 4898
   Contig N20: 3317
   Contig N30: 2445
   Contig N40: 1830
   Contig N50: 1357

   Median contig length: 352
   Average contig: 728.03
   Total assembled bases: 45879553

```

## 06_Ol

```
################################
## Counts of transcripts, etc.
################################
Total trinity 'genes':  64282
Total trinity transcripts:   100481
Percent GC: 35.70

########################################
Stats based on ALL transcript contigs:
########################################

   Contig N10: 5486
   Contig N20: 3965
   Contig N30: 3028
   Contig N40: 2375
   Contig N50: 1866

   Median contig length: 447
   Average contig: 959.27
   Total assembled bases: 96388352


#####################################################
## Stats based on ONLY LONGEST ISOFORM per 'GENE':
#####################################################

   Contig N10: 4825
   Contig N20: 3342
   Contig N30: 2471
   Contig N40: 1872
   Contig N50: 1369

   Median contig length: 357
   Average contig: 731.99
   Total assembled bases: 47053619

```

## 07_Pr

```
################################
## Counts of transcripts, etc.
################################
Total trinity 'genes':  56798
Total trinity transcripts:   92572
Percent GC: 37.40

########################################
Stats based on ALL transcript contigs:
########################################

   Contig N10: 6481
   Contig N20: 4676
   Contig N30: 3736
   Contig N40: 2960
   Contig N50: 2341

   Median contig length: 538
   Average contig: 1169.79
   Total assembled bases: 108290220


#####################################################
## Stats based on ONLY LONGEST ISOFORM per 'GENE':
#####################################################

   Contig N10: 5761
   Contig N20: 4107
   Contig N30: 3008
   Contig N40: 2283
   Contig N50: 1693

   Median contig length: 376
   Average contig: 829.30
   Total assembled bases: 47102779


```

## 08_Ar

```
################################
## Counts of transcripts, etc.
################################
Total trinity 'genes':  202721
Total trinity transcripts:   249888
Percent GC: 27.13

########################################
Stats based on ALL transcript contigs:
########################################

   Contig N10: 3219
   Contig N20: 2170
   Contig N30: 1504
   Contig N40: 986
   Contig N50: 680

   Median contig length: 350
   Average contig: 564.43
   Total assembled bases: 141043361


#####################################################
## Stats based on ONLY LONGEST ISOFORM per 'GENE':
#####################################################

   Contig N10: 2421
   Contig N20: 1436
   Contig N30: 897
   Contig N40: 645
   Contig N50: 507

   Median contig length: 334
   Average contig: 478.33
   Total assembled bases: 96967451


```

# Contamination removal

## NCBI FCS GX

Get for each species the taxonomy id on https://www.ncbi.nlm.nih.gov/Taxonomy/Browser/wwwtax.cgi

+ Anchomenus dorsalis (NCBI taxonomy ID 586010)
+ Pterostichus melanarius (NCBI taxonomy ID 60768)
+ Zabrus tenebrioides (NCBI taxonomy ID 272042)
+ Amara similata (=Harpalus similatus) (NCBI taxonomy ID 272013)
+ Harpalus affinis (NCBI taxonomy ID 247283)
+ Harpalus eous (NCBI taxonomy ID 2107211)
+ Poecilus cupreus (=Pterostichus cupreus) (NCBI taxonomy ID 270614)
+ Aphidius rhopalosiphi (NCBI taxonomy ID 55891)

Then use the NCBI FCS GX tool to detect contaminations (https://usegalaxy.org/)

Rename the two output files with the name of the libaries.

```{bash, eval=FALSE}

for i in *Action_report.tabular; do
  sample=$(echo "$i" | cut -d_ -f1-2)
  echo "grep "^TRINITY_" $i | cut -f1 | sort | uniq > ${sample}_list_contamination.txt"
  grep "^TRINITY_" $i | cut -f1 | sort | uniq > ${sample}_list_contamination.txt
done

 wc -l *list_contamination.txt
  3452 01_Ad_list_contamination.txt
  2053 02_Pm_list_contamination.txt
   356 03_Zt_list_contamination.txt
    57 04_Asp_list_contamination.txt
   397 05_G_list_contamination.txt
   667 06_Ol_list_contamination.txt
   200 07_Pr_list_contamination.txt
  1133 08_Ar_list_contamination.txt
  8315 total
```


### Fungal presence

```{bash, eval=FALSE}

for i in *_Taxonomy_report.tabular; do 
  echo -e "\n======$i"
  awk '$9=="fung:fungi" {print $0}' $i | \
  cut -f6,7 | sort | uniq -c | sort | sed 's/^\s*//' | \
  awk '{print $2,$3,$1}'  | sort -k3nr | head
done

======01_Ad_Taxonomy_report.tabular
Mucor circinelloides 454
Mucor velutinosus 414
Mucor lusitanicus 410
Mucor irregularis 267
Mucor racemosus 118
Pilaira anomala 100
Mucor plumbeus 90
Thamnidium elegans 80
Mucor piriformis 69
Mucor mucedo 66

======02_Pm_Taxonomy_report.tabular
Mucor indicus 225
Mucor circinelloides 82
Gigaspora margarita 79
Mucor racemosus 68
Mucor lusitanicus 65
Mucor velutinosus 62
Choanephora cucurbitarum 56
Mucor irregularis 56
Massospora cicadina 54
Gilbertella persicaria 53

======03_Zt_Taxonomy_report.tabular
Massospora cicadina 32
Gigaspora margarita 25
Entomophthora muscae 21
Gigaspora rosea 21
Mucor irregularis 16
Diversispora epigaea 15
Rhizophagus clarus 12
Mucor lusitanicus 11
Mucor mucedo 11
Mucor racemosus 7

======04_Asp_Taxonomy_report.tabular
Massospora cicadina 32
Gigaspora margarita 31
Gigaspora rosea 28
Diversispora epigaea 17
Entomophthora muscae 15
Mucor circinelloides 7
Thamnidium elegans 7
Mucor irregularis 6
Geosiphon pyriformis 5
Mucor lusitanicus 5

======05_G_Taxonomy_report.tabular
Gigaspora margarita 45
Massospora cicadina 41
Gigaspora rosea 30
Entomophthora muscae 23
Diversispora epigaea 22
Rhizophagus clarus 12
Geosiphon pyriformis 9
Mucor irregularis 9
Rhizophagus irregularis 8
Cunninghamella bertholletiae 6

======06_Ol_Taxonomy_report.tabular
Mucor irregularis 60
Gigaspora margarita 40
Massospora cicadina 36
Gigaspora rosea 33
Entomophthora muscae 28
Mucor racemosus 28
Mucor circinelloides 25
Mucor mucedo 22
Mucor lusitanicus 21
Thamnidium elegans 20

======07_Pr_Taxonomy_report.tabular
Gigaspora margarita 41
Massospora cicadina 38
Entomophthora muscae 19
Rhizophagus clarus 18
Gigaspora rosea 17
Diversispora epigaea 15
Mucor piriformis 9
Linnemannia elongata 7
Podila humilis 6
Rhizophagus irregularis 6

======08_Ar_Taxonomy_report.tabular
Neoconidiobolus thromboides 35
Massospora cicadina 17
Gigaspora margarita 14
Gigaspora rosea 10
Entomophthora muscae 8
Cunninghamella blakesleeana 5
Rhizophagus irregularis 5
Backusella circina 4
Diversispora epigaea 4
Halteromyces radiatus 4

```



## NCBI FCS adaptor

Download Galaxy file and replace spaces by underscore and remove colons.

+ In Galaxy, select the output files Adaptor report => Build Dataset List
+ Download the list of files
+ Note which dataset number matches to which annotation

The software use this list of sequence: https://ftp.ncbi.nlm.nih.gov/pub/UniVec/UniVec

```{bash, eval=FALSE}

unzip '67_ Adaptor_reports.zip'

# Replace spaces by underscore
find . -depth -name '*' | while IFS= read -r f ; do mv -i "$f" "$(dirname "$f")/$(basename "$f"|tr ':' )" ; done

# Remove colons
find . -name "*:*" -exec rename ':' '' {} \;

# Remove prefix NCBI_FCS_Adaptor_on_data_

 rename 'NCBI_FCS_Adaptor_on_data_' '' *tabular


# Rename each file manually to replace Dataset number by library names


# Check number of transcri
for i in *tabular; do echo $i; grep -v -n "^#accession" $i | wc -l; done
NCBI_FCS_Adaptor_on_data_01_Ad_Adaptor_report.tabular
0
NCBI_FCS_Adaptor_on_data_02_Pm_Adaptor_report.tabular
3
NCBI_FCS_Adaptor_on_data_03_Zt_Adaptor_report.tabular
2
NCBI_FCS_Adaptor_on_data_04_Asp_Adaptor_report.tabular
0
NCBI_FCS_Adaptor_on_data_05_G_Adaptor_report.tabular
4
NCBI_FCS_Adaptor_on_data_06_Ol_Adaptor_report.tabular
1
NCBI_FCS_Adaptor_on_data_07_Pr_Adaptor_report.tabular
0
NCBI_FCS_Adaptor_on_data_08_Ar_Adaptor_report.tabular
9

# Check most abundant adapter type
cat *.tabular | cut -f5 | sort | uniq -c | sed 's/^\s*//' | awk '{print $2,$1}'  | sort -k2nr | grep -v "^name"
CONTAMINATION_SOURCE_TYPE_ADAPTOR:NGB01114.1:Illumina 12
CONTAMINATION_SOURCE_TYPE_ADAPTOR:NGB00201.1:NotI 2
CONTAMINATION_SOURCE_TYPE_ADAPTOR:NGB00749.1:Illumina 1
CONTAMINATION_SOURCE_TYPE_ADAPTOR:NGB00754.1:Illumina 1
CONTAMINATION_SOURCE_TYPE_ADAPTOR:NGB00846.1:NEBNext 1
CONTAMINATION_SOURCE_TYPE_ADAPTOR:NGB00972.1:Pacific 1
CONTAMINATION_SOURCE_TYPE_ADAPTOR:NGB01094.1:Rubicon 1

```

14 transcripts with Illumina adaptors.

Keep only transcript ID:

```{bash, eval=FALSE}
for i in *tabular; do 
  grep -v "^#" $i | cut -f1 > ${i%.*}.clean.tabular
done
```


## Remove contaminated transcripts

Combine contaminated transcripts with adaptors and other species and cleanup fasta and map files

```{bash, eval=FALSE}

while read i; do
cat fcs_adaptor_reports/${i}_Adaptor_report.clean.tabular \
  fcs_gx_reports/${i}_list_contamination.txt | sort | uniq > ${i}_contaminants.txt
done < list.txt

wc -l *_contaminants.txt
  3452 01_Ad_contaminants.txt
  2056 02_Pm_contaminants.txt
   358 03_Zt_contaminants.txt
    57 04_Asp_contaminants.txt
   401 05_G_contaminants.txt
   667 06_Ol_contaminants.txt
   200 07_Pr_contaminants.txt
  1142 08_Ar_contaminants.txt

```

```{bash, eval=FALSE}

while read i; do
  seqkit grep -w -v -f ncbi_fcs_gx_adaptor/${i}_contaminants.txt \
  trinity_output/${i}/${i}_trinity.Trinity.fasta > trinity_output/${i}/${i}_trinity.Trinity.cleaned.fasta
done < list.txt

[INFO] 3452 patterns loaded from file
[INFO] 2056 patterns loaded from file
[INFO] 358 patterns loaded from file
[INFO] 57 patterns loaded from file
[INFO] 401 patterns loaded from file
[INFO] 667 patterns loaded from file
[INFO] 200 patterns loaded from file
[INFO] 1142 patterns loaded from file

```

Also fix the gene_trans_map

```{bash, eval=FALSE}
# Important to use -w, otherwise, partial match will remove transcripts
# that are not in the contamination list!
while read i; do
  grep -w -v -f ncbi_fcs_gx_adaptor/${i}_contaminants.txt trinity_output/${i}/${i}_trinity.Trinity.fasta.gene_trans_map > trinity_output/${i}/${i}_trinity.Trinity.cleaned.fasta.gene_trans_map
done < list.txt

```

Gather final files

```{bash, eval=FALSE}

cd /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinity_output

find . -type f -name *cleaned.fasta -exec cp {} summary/ \;

find . -type f -name *cleaned.fasta.gene_trans_map -exec cp {} summary/ \;

```

Rename with insect species

```{bash, eval=FALSE}

cat library_to_species.txt
01_Ad Anchomenus_dorsalis
02_Pm Pterostichus_melanarius
03_Zt Zabrus_tenebrioides
04_Asp  Amara_similata
05_G  Harpalus_affinis
06_Ol Harpalus_eous
07_Pr Poecilus_cupreus
08_Ar Aphidius_rhopalosiphi


while read i; do
  library=$(echo "$i" | cut -f1)
  species=$(echo "$i" | cut -f2)
  echo "rename $library $species *fasta*"
  rename $library $species *fasta*
done < library_to_species.txt


# Remove the .Trinity.cleaned.
rename '.Trinity.cleaned' '' *fasta*

seqkit stats *fasta
processed files:  8 / 8 [======================================] ETA: 0s. done
file                                   format  type  num_seqs      sum_len  min_len  avg_len  max_len
Amara_similata_trinity.fasta           FASTA   DNA     72,424   82,575,412      174  1,140.2   24,631
Anchomenus_dorsalis_trinity.fasta      FASTA   DNA     99,685  108,787,099      184  1,091.3   33,453
Aphidius_rhopalosiphi_trinity.fasta    FASTA   DNA    248,746  140,588,291      188    565.2   35,915
Harpalus_affinis_trinity.fasta         FASTA   DNA    103,989  100,209,791      177    963.7   25,018
Harpalus_eous_trinity.fasta            FASTA   DNA     99,814   96,095,807      178    962.7   20,450
Poecilus_cupreus_trinity.fasta         FASTA   DNA     92,372  108,133,517      180  1,170.6   24,578
Pterostichus_melanarius_trinity.fasta  FASTA   DNA    138,089  115,188,695      177    834.2   26,406
Zabrus_tenebrioides_trinity.fasta      FASTA   DNA     76,576   86,687,902      182  1,132.1   26,778


```

| file                                  | format | type | num_seqs | sum_len     | min_len | avg_len  | max_len |
|---------------------------------------|--------|------|----------|-------------|---------|----------|---------|
| Amara_similata_trinity.fasta          | FASTA  | DNA  | 72,424   | 82,575,412  | 174     | 1,140.20 | 24,631  |
| Anchomenus_dorsalis_trinity.fasta     | FASTA  | DNA  | 99,685   | 108,787,099 | 184     | 1,091.30 | 33,453  |
| Aphidius_rhopalosiphi_trinity.fasta   | FASTA  | DNA  | 248,746  | 140,588,291 | 188     | 565.2    | 35,915  |
| Harpalus_affinis_trinity.fasta        | FASTA  | DNA  | 103,989  | 100,209,791 | 177     | 963.7    | 25,018  |
| Harpalus_eous_trinity.fasta           | FASTA  | DNA  | 99,814   | 96,095,807  | 178     | 962.7    | 20,450  |
| Poecilus_cupreus_trinity.fasta        | FASTA  | DNA  | 92,372   | 108,133,517 | 180     | 1,170.60 | 24,578  |
| Pterostichus_melanarius_trinity.fasta | FASTA  | DNA  | 138,089  | 115,188,695 | 177     | 834.2    | 26,406  |
| Zabrus_tenebrioides_trinity.fasta     | FASTA  | DNA  | 76,576   | 86,687,902  | 182     | 1,132.10 | 26,778  |



Number of genes:

```{bash, eval=FALSE}
for i in *gene_trans_map; do echo $(echo "$i" | cut -d_ -f1,2); cut -f1 $i| sort | uniq | wc -l; done
Amara_similata
45382
Anchomenus_dorsalis
64702
Aphidius_rhopalosiphi
201906
Harpalus_affinis
62673
Harpalus_eous
63803
Poecilus_cupreus
56672
Pterostichus_melanarius
92836
Zabrus_tenebrioides
51663

```

| species                 | nb_genes |
|-------------------------|----------|
| Amara_similata          | 45,382   |
| Anchomenus_dorsalis     | 64,702   |
| Aphidius_rhopalosiphi   | 201,906  |
| Harpalus_affinis        | 62,673   |
| Harpalus_eous           | 63,803   |
| Poecilus_cupreus        | 56,672   |
| Pterostichus_melanarius | 92,836   |
| Zabrus_tenebrioides     | 51,663   |



## Summary contamination

| library | species                 | Transcripts w/ conta | Transcripts w/o conta | diff1 | Genes w/ conta | Genes w/o conta | diff2 |
|---------|-------------------------|----------------------|-----------------------|-------|----------------|-----------------|-------|
| 01_Ad   | Anchomenus_dorsalis     | 103,137              | 99,685                | 3,452 | 67,047         | 64,702          | 2,345 |
| 02_Pm   | Pterostichus_melanarius | 140,145              | 138,089               | 2,056 | 94,235         | 92,835          | 1,400 |
| 03_Zt   | Zabrus_tenebrioides     | 76,934               | 76,576                | 358   | 51,952         | 51,663          | 289   |
| 04_Asp  | Amara_similata          | 72,481               | 72,424                | 57    | 45,431         | 45,382          | 49    |
| 05_G    | Harpalus_affinis        | 104,390              | 103,989               | 401   | 63,019         | 62,673          | 346   |
| 06_Ol   | Harpalus_eous           | 100,481              | 99,814                | 667   | 64,282         | 63,803          | 479   |
| 07_Pr   | Poecilus_cupreus        | 92,572               | 92,372                | 200   | 56,798         | 56,672          | 126   |
| 08_Ar   | Aphidius_rhopalosiphi   | 249,888              | 248,746               | 1,142 | 202,721        | 201,906         | 815   |



# Transdecoder

Reuse script used for OPM analysis. Transdecoder needs a different output folder for each file, otherwise, it will override files of other instances.

```{bash, eval=FALSE}

#!/bin/bash
#SBATCH --job-name=TransDecoder
#SBATCH --nodes=1
#SBATCH --partition=medium
#SBATCH --cpus-per-task=1
#SBATCH --time=00:40:00
#SBATCH --mem=20gb
#SBATCH --mail-type=END
#SBATCH --mail-user=johan.zicola@uni-goettingen.de
#SBATCH --output=TransDecoder.%J.out
#SBATCH --error=TransDecoder.%J.err
#SBATCH --array=1-8

mem_gbytes=$(( $SLURM_MEM_PER_NODE/1024 ))

echo "Starting at $(date)"
echo "Job submitted to the ${SLURM_JOB_PARTITION} partition, the default partition on ${SLURM_CLUSTER_NAME}"
echo "Job name: ${SLURM_JOB_NAME}, Job ID: ${SLURM_JOB_ID}"
echo "I have ${SLURM_CPUS_ON_NODE} CPUs and ${mem_gbytes}GiB of RAM on compute node $(hostname)"

output_folder="/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/transdecoder_output"
trinity_folder="/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinity_output/summary"

file_name=$(sed -n ${SLURM_ARRAY_TASK_ID}p ${trinity_folder}/library_to_species.txt | cut -f2)
mkdir ${output_folder}/${file_name}

~/bin/TransDecoder-TransDecoder-v5.7.0/TransDecoder.LongOrfs \
    --gene_trans_map ${trinity_folder}/${file_name}_trinity.fasta.gene_trans_map \
    -t ${trinity_folder}/${file_name}_trinity.fasta \
    --output_dir ${output_folder}/${file_name} 
    && \
    ~/bin/TransDecoder-TransDecoder-v5.7.0/TransDecoder.Predict \
    -t ${trinity_folder}/${file_name}_trinity.fasta \
    --output_dir ${output_folder}/${file_name}

```



# Trinotate apptainer

https://southgreenplatform.github.io/trainings/trinityTrinotate/TP-annotation/

> Transcrits assembled using Trinity can be easily annotate using trinotate https://github.com/Trinotate/Trinotate.github.io/wiki.
> Trinotate use different methods for functional annotation including homology search to known sequence data (BLAST+/SwissProt), protein domain identification (HMMER/PFAM), protein signal peptide and transmembrane domain prediction (signalP/tmHMM), and take advantage from annotation databases (eggNOG/GO/Kegg). These data are integrated into a SQLite database which allows to create an annotation report for a transcriptome.

Download last singularity image on https://data.broadinstitute.org/Trinity/TRINOTATE_SINGULARITY/ (trinotate.v4.0.2)

See the trick https://github.com/Trinotate/Trinotate/wiki/TrinotateViaSingularity

Run TmHMM and SignalPv6 separately: https://github.com/Trinotate/Trinotate/wiki/Running_Loading_Separately

> Certain software tools such as those provided by DTU Health Tech have restricted licensing that prevent distribution, and so these cannot be included in our Docker or Singularity images and must be obtained/installed separately by users. These, in particular, include both TmHMM and SignalP utilities. While these can be installed natively and automatically run by Trinotate, they cannot be automatically run via the Docker/Singularity images. Instructions below provide for execution and loading of the results into the Trinotate sqlite database.


Version of signalp6 and tmhmm used:

```{bash, eval=FALSE}

signalp6 --version
SignalP 6.0 Signal peptide prediction tool 6.0h

which tmhmm
~/bin/tmhmm-2.0c/bin/tmhmm

```


## Download manually eggnog.db.gz

I can't gunzip the eggnog.db.gz so I think it is corrupted. Redownload from the eggnog5.embl.de website

```{bash, eval=FALSE}

cd /scratch/users/zicola/TRINOTATE_DATA_DIR/EGGNOG_DATA_DIR

http://eggnog5.embl.de/download/emapperdb-5.0.2/
../
eggnog.db.gz                                       02-Mar-2021 14:01      6G
eggnog.taxa.tar.gz                                 02-Mar-2021 14:04     69M
eggnog_proteins.dmnd.gz                            02-Mar-2021 14:04      5G
mmseqs.tar.gz                                      02-Mar-2021 14:07      5G
pfam.tar.gz                                        02-Mar-2021 14:08    938M

gunzip *

```


## 04_trinotate.sh


First steps are in apptainer:
- Create database
- Initialize database
- Run all but tmhmm and signalp softwares and load in database

```{bash, eval=FALSE}

# New home not to fill my real home with Trinotate output files
export HOME=$NEWHOME

# Path to trinotate executable
export TRINOTATE_HOME=/usr/local/src/Trinotate

# Path to Trinity and Transdecoder files
export WORKPATH=/scratch/users/zicola/F22FTSEUHT1495_LIBgsmfR_judi_8_insects
export GENE_TRANS_MAP=${WORKPATH}/trinity_output/summary/${SPECIES}_trinity.fasta.gene_trans_map
export TRINITY_FASTA=${WORKPATH}/trinity_output/summary/${SPECIES}_trinity.fasta
export TRANSDECODER_PEP=${WORKPATH}/transdecoder_output/${SPECIES}_trinity.fasta.transdecoder.pep 

# Database
export TRINOTATE_DATA_DIR=/scratch/users/zicola/TRINOTATE_DATA_DIR
export EGGNOG_DATA_DIR=/scratch/users/zicola/TRINOTATE_DATA_DIR/EGGNOG_DATA_DIR
export SQLITE_DB=${WORKPATH}/trinotate_output/${SPECIES}/database_${SPECIES}.sqlite

# Move to home before launching Trinotate
cd $HOME

$TRINOTATE_HOME/Trinotate \
   --db $SQLITE_DB \
   --create \
   --use_diamond

$TRINOTATE_HOME/Trinotate \
   --db $SQLITE_DB --init \
   --gene_trans_map $GENE_TRANS_MAP \
   --transcript_fasta $TRINITY_FASTA \
   --transdecoder_pep $TRANSDECODER_PEP

$TRINOTATE_HOME/Trinotate \
   --CPU 10 \
   --db $SQLITE_DB \
   --run ALL \
   --transcript_fasta $TRINITY_FASTA \
   --transdecoder_pep $TRANSDECODER_PEP \
   --use_diamond

```

Remember to make the shell script executable.

```{bash, eval=FALSE}
chmod +x 04_trinotate.sh
```


### sbatch_04_trinotate.sh

```{bash, eval=FALSE}
#!/bin/bash
#SBATCH --job-name=04_trinotate
#SBATCH --nodes=1
#SBATCH --partition=medium
#SBATCH --time=48:00:00
#SBATCH --cpus-per-task=10
#SBATCH --mem=12G
#SBATCH --mail-type=END
#SBATCH --mail-user=johan.zicola@uni-goettingen.de
#SBATCH --output=job-%x-%j.out
#SBATCH --error=job-%x-%j.err

mem_gbytes=$(( $SLURM_MEM_PER_NODE/1024 ))

echo "Starting at $(date)"
echo "Job submitted to the ${SLURM_JOB_PARTITION} partition, the default partition on ${SLURM_CLUSTER_NAME}"
echo "Job name: ${SLURM_JOB_NAME}, Job ID: ${SLURM_JOB_ID}"
echo "I have ${SLURM_CPUS_ON_NODE} CPUs and ${mem_gbytes}GiB of RAM on compute node $(hostname)"

# Modules to load
module purge
module load apptainer

# Create newhome
mkdir /usr/users/zicola/trinotate_${SPECIES}

export APPTAINERENV_NEWHOME=/usr/users/zicola/trinotate_${SPECIES}
export APPTAINERENV_SPECIES=$SPECIES

apptainer run -C -B /usr/users/zicola -B /scratch/users/zicola/ /usr/users/zicola/bin/trinotate.v4.0.2.simg /scratch/users/zicola/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/04_trinotate.sh

```

Launch sbatch with the species argument.

```{bash, eval=FALSE}
sbatch --export=SPECIES="Amara_similata" sbatch_trinotate.sh
```

Works, now, all Trinotate output will go in `~/newhome` instead of `$HOME`.

```{bash, eval=FALSE}
ll ~/newhome/
total 51M
drwxr-x---  3 zicola UANP    6 Jan 27 19:36 ./
drwx------ 63 zicola UANP 9.3K Jan 27 19:33 ../
-rw-r-----  1 zicola UANP  20M Jan 27 19:40 TrinotatePFAM.out
-rw-r-----  1 zicola UANP    0 Jan 27 19:36 __init.ok
drwxr-x---  2 zicola UANP    0 Jan 27 19:36 __trinotate_run_checkpts/
-rw-r-----  1 zicola UANP  22M Jan 27 19:40 pfam.log
-rw-r-----  1 zicola UANP 1.6M Jan 27 19:36 uniprot_sprot.diamond.blastp.outfmt6
-rw-r-----  1 zicola UANP 1.6M Jan 27 19:36 uniprot_sprot.diamond.blastx.outfmt6
```


```{bash, eval=FALSE}
sacct -j JOBID -o jobid,jobname,AllocCPUS,ReqMem,MaxRSS,Elapsed

sacct -j 5411597  -o jobid,jobname,AllocCPUS,ReqMem,MaxRSS,Elapsed
JobID           JobName  AllocCPUS     ReqMem     MaxRSS    Elapsed
------------ ---------- ---------- ---------- ---------- ----------
5411597       trinotate         10        12G              12:43:41
5411597.bat+      batch         10              8463548K   12:43:41
5411597.ext+     extern         10                     0   12:43:41

```

12h40 needed.

```{bash, eval=FALSE}
ls ~/newhome/
total 475232
-rw-r----- 1 zicola UANP 124819074 Jan 27 20:11 TrinotatePFAM.out
-rw-r----- 1 zicola UANP         0 Jan 27 19:36 __init.ok
drwxr-x--- 2 zicola UANP         2 Jan 28 08:19 __trinotate_run_checkpts
-rw-r----- 1 zicola UANP  44173297 Jan 28 08:19 eggnog_mapper.emapper.annotations
-rw-r----- 1 zicola UANP   9483642 Jan 28 02:27 eggnog_mapper.emapper.hits
-rw-r----- 1 zicola UANP   2365038 Jan 28 08:19 eggnog_mapper.emapper.seed_orthologs
-rw-r----- 1 zicola UANP 135214516 Jan 28 02:16 infernal.log
-rw-r----- 1 zicola UANP     80487 Jan 28 02:16 infernal.out
-rw-r----- 1 zicola UANP 140305066 Jan 27 20:11 pfam.log
-rw-r----- 1 zicola UANP   1583830 Jan 27 19:36 uniprot_sprot.diamond.blastp.outfmt6
-rw-r----- 1 zicola UANP   1673060 Jan 27 19:36 uniprot_sprot.diamond.blastx.outfmt6

```

```{bash, eval=FALSE}
cd /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects

587M Jan 28 08:19 database_Amara_similata.sqlite

```


## 05_trinotate.sh

Second step:
- Run tmhmm and load results in database
- Run signalp and load results in database
- Generate Trinotate report and GO annotations

```{bash, eval=FALSE}

#!/bin/bash
#SBATCH --job-name=05_trinotate
#SBATCH --nodes=1
#SBATCH --partition=medium
#SBATCH --time=20:00:00
#SBATCH --cpus-per-task=1
#SBATCH --mem=10G
#SBATCH --mail-type=END
#SBATCH --mail-user=johan.zicola@uni-goettingen.de
#SBATCH --output=job-%x-%j.out
#SBATCH --error=job-%x-%j.err

mem_gbytes=$(( $SLURM_MEM_PER_NODE/1024 ))

echo "Starting at $(date)"
echo "Job submitted to the ${SLURM_JOB_PARTITION} partition, the default partition on ${SLURM_CLUSTER_NAME}"
echo "Job name: ${SLURM_JOB_NAME}, Job ID: ${SLURM_JOB_ID}"
echo "I have ${SLURM_CPUS_ON_NODE} CPUs and ${mem_gbytes}GiB of RAM on compute node $(hostname)"

# Modules to load
module purge
module load apptainer

# Test that all programs needed are installed
test_program(){
  command -v $1 >/dev/null || \
    { echo >&2 " $1 required but not installed. Aborting"; exit 1; }
}

test_program "tmhmm"
test_program "signalp6"

# Path to trinotate executable
export TRINOTATE_HOME=/usr/local/src/Trinotate

# Path to Trinity and Transdecoder files
export WORKPATH=/scratch/users/zicola/F22FTSEUHT1495_LIBgsmfR_judi_8_insects
export GENE_TRANS_MAP=${WORKPATH}/trinity_output/summary/${SPECIES}_trinity.fasta.gene_trans_map
export TRINITY_FASTA=${WORKPATH}/trinity_output/summary/${SPECIES}_trinity.fasta
export TRANSDECODER_PEP=${WORKPATH}/transdecoder_output/${SPECIES}_trinity.fasta.transdecoder.pep 

# Database
export TRINOTATE_DATA_DIR=/scratch/users/zicola/TRINOTATE_DATA_DIR
export EGGNOG_DATA_DIR=/scratch/users/zicola/TRINOTATE_DATA_DIR/EGGNOG_DATA_DIR
export SQLITE_DB=${WORKPATH}/trinotate_output/${SPECIES}/database_${SPECIES}.sqlite

echo "##############################"
echo  "Trinotate Run signalp"
date
echo "##############################"

cd ${WORKPATH}/trinotate_output/${SPECIES}/

signalp6 --fastafile $TRANSDECODER_PEP \
   --output_dir ${WORKPATH}/trinotate_output/${SPECIES}/sigP6outdir \
   --format none --organism euk --mode fast


apptainer exec -C -B /usr/users/zicola -B $WORKPATH -B $TRINOTATE_DATA_DIR -e ~/bin/trinotate.v4.0.2.simg \
   $TRINOTATE_HOME/Trinotate \
   --db $SQLITE_DB \
   --LOAD_signalp ${WORKPATH}/trinotate_output/${SPECIES}/sigP6outdir/output.gff3

echo "##############################"
echo  "Trinotate Run TMHMM"
date
echo "##############################"

tmhmm --short $TRANSDECODER_PEP > ${WORKPATH}/trinotate_output/${SPECIES}/tmhmm.v2.out

apptainer exec -C -B /usr/users/zicola -B $WORKPATH -B $TRINOTATE_DATA_DIR -e ~/bin/trinotate.v4.0.2.simg \
   $TRINOTATE_HOME/Trinotate \
   --db $SQLITE_DB \
   --LOAD_tmhmmv2 ${WORKPATH}/trinotate_output/${SPECIES}/tmhmm.v2.out

echo "###########################"
echo "Generating report table"
date
echo "###########################"

apptainer exec -C -B /usr/users/zicola -B $WORKPATH -B $TRINOTATE_DATA_DIR -e ~/bin/trinotate.v4.0.2.simg \
   $TRINOTATE_HOME/Trinotate \
   --db ${SQLITE_DB} \
   --report \
   --incl_pep \
   --incl_trans > ${WORKPATH}/trinotate_output/${SPECIES}/Trinotate_report.tsv

# Take about 10 min

##################
## Misc value adds
date
##################

# Extract GO terms

apptainer exec -C -B /usr/users/zicola -B $WORKPATH -B $TRINOTATE_DATA_DIR -e ~/bin/trinotate.v4.0.2.simg \
   $TRINOTATE_HOME/util/extract_GO_assignments_from_Trinotate_xls.pl \
  --Trinotate_xls ${WORKPATH}/trinotate_output/${SPECIES}/Trinotate_report.tsv \
  -G -I > ${WORKPATH}/trinotate_output/${SPECIES}/Trinotate_report.tsv.gene_ontology

# Tke few seconds

# Generate trinotate report summary statistics
apptainer exec -C -B /usr/users/zicola -B $WORKPATH -e ~/bin/trinotate.v4.0.2.simg \
   $TRINOTATE_HOME/util/report_summary/trinotate_report_summary.pl \
  ${WORKPATH}/trinotate_output/${SPECIES}/Trinotate_report.tsv \
  ${WORKPATH}/trinotate_output/${SPECIES}/Trinotate_report_stats

echo "##########################"
echo "done.  See annotation summary file:  Trinotate_report.tsv"
date
echo "##########################"

```


## Run


### Step 04

```{bash, eval=FALSE}

sbatch --export=SPECIES="Anchomenus_dorsalis" sbatch_trinotate.sh

sbatch --export=SPECIES="Amara_similata" sbatch_trinotate.sh

sbatch --export=SPECIES="Aphidius_rhopalosiphi" sbatch_trinotate.sh

sbatch --export=SPECIES="Pterostichus_melanarius" sbatch_trinotate.sh

sbatch --export=SPECIES="Zabrus_tenebrioides" sbatch_trinotate.sh

sbatch --export=SPECIES="Harpalus_affinis" sbatch_trinotate.sh

sbatch --export=SPECIES="Harpalus_eous" sbatch_trinotate.sh

sbatch --export=SPECIES="Poecilus_cupreus" sbatch_trinotate.sh

```


Job for Aphidius_rhopalosiphi overtime the 16 hours and was killed during infernal run. Relaunch with a 24h time. Trinotate should consider the checkpoint files in $HOME and start from infernal.

```{bash, eval=FALSE}

rm ~/trinotate_Aphidius_rhopalosiphi/infernal*

\ls -1 trinotate_Aphidius_rhopalosiphi/__trinotate_run_checkpts/
diamond.uniprot_sprot.blastp.loaded.ok
diamond.uniprot_sprot.blastp.ok
diamond.uniprot_sprot.blastx.loaded.ok
diamond.uniprot_sprot.blastx.ok
pfam_hmmsearch.ok
pfam_hmmsearch_loading.ok
pipeliner.36.cmds

cd /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects

sbatch --export=SPECIES="Aphidius_rhopalosiphi" sbatch_trinotate.sh

```

Yes, it skips all steps until infernal and continue there.


Job killed again due to time out but now at the emapper part. Relaunch again


```{bash, eval=FALSE}

sbatch --export=SPECIES="Aphidius_rhopalosiphi" sbatch_trinotate.sh

```

```{bash, eval=FALSE}
# emapper.py  -i /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/transdecoder_output/Aphidius_rhopalosiphi_trinity.fasta.transdecoder.pep --cpu 10 --data_dir /scratch/users/zicola/TRINOTATE_DATA_DIR/EGGNOG_DATA_DIR -o eggnog_mapper
Output files detected in disk. Use --resume or --override to continue

```

Fail, I need first to remove the eggmapper content in `~/trinotate_Aphidius_rhopalosiphi` that was killed



### Step 05

Wait this job is finished before launching `05_trinotate.sh`

```{bash, eval=FALSE}

sbatch --export=SPECIES="Pterostichus_melanarius" 05_trinotate.sh

sbatch --export=SPECIES="Zabrus_tenebrioides" 05_trinotate.sh

sbatch --export=SPECIES="Harpalus_affinis" 05_trinotate.sh

sbatch --export=SPECIES="Harpalus_eous" 05_trinotate.sh

sbatch --export=SPECIES="Anchomenus_dorsalis" 05_trinotate.sh

sbatch --export=SPECIES="Poecilus_cupreus" 05_trinotate.sh

sbatch --export=SPECIES="Amara_similata" 05_trinotate.sh
# Done


sbatch --export=SPECIES="Aphidius_rhopalosiphi" 05_trinotate.sh

```


# BUSCO

Based on the odb10 list (`busco --list-datasets`), there are no coleoptera database. I would therefore use insecta_odb12 to be more general. However, for Aphidius_rhopalosiphi, I can use the hymenoptera_odb10 database.

See doc on https://busco.ezlab.org/busco_userguide.html#running-busco

```{bash, eval=FALSE}

busco --list-datasets

```


## Get longest isoforms

```{bash, eval=FALSE}

cd /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinity_output/summary

export WORKDIR=/usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinity_output/summary

for i in *fasta; do
   apptainer exec -e -C -B $WORKDIR /usr/users/zicola/bin/trinityrnaseq.v2.15.1.simg \
/usr/local/bin/util/misc/get_longest_isoform_seq_per_trinity_gene.pl $WORKDIR/$i > $WORKDIR/longest_isoforms/${i%.*}.longest_isoforms.fasta
done

```

```{bash, eval=FALSE}
seqkit stats *
processed files:  9 / 9 [======================================] ETA: 0s. done
file                                                    format  type  num_seqs     sum_len  min_len  avg_len  max_len
Amara_similata_trinity.longest_isoforms.fasta           FASTA   DNA     45,382  39,170,740      201    863.1   24,631
Anchomenus_dorsalis_trinity.longest_isoforms.fasta      FASTA   DNA     64,702  49,429,920      201      764   33,453
Aphidius_rhopalosiphi_trinity.longest_isoforms.fasta    FASTA   DNA    201,906  96,690,872      201    478.9   35,915
Harpalus_affinis_trinity.longest_isoforms.fasta         FASTA   DNA     62,673  45,743,303      201    729.9   25,018
Harpalus_eous_trinity.longest_isoforms.fasta            FASTA   DNA     63,803  46,868,568      201    734.6   20,450
Poecilus_cupreus_trinity.longest_isoforms.fasta         FASTA   DNA     56,672  47,028,774      201    829.8   24,578
Pterostichus_melanarius_trinity.longest_isoforms.fasta  FASTA   DNA     92,836  57,883,850      201    623.5   26,406
Zabrus_tenebrioides_trinity.longest_isoforms.fasta      FASTA   DNA     51,663  43,363,104      201    839.3   26,778
longest_isoforms.fasta                                  FASTA   DNA     51,663  43,363,104      201    839.3   26,778

```


Compared to the number of peptides from Transdecoder

```{bash, eval=FALSE}
cd /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/transdecoder_output

seqkit stats *pep

processed files:  8 / 8 [======================================] ETA: 0s. done
file                                                    format  type     num_seqs     sum_len  min_len  avg_len  max_len
Amara_similata_trinity.fasta.transdecoder.pep           FASTA   Protein    30,556  14,245,018       88    466.2    7,989
Anchomenus_dorsalis_trinity.fasta.transdecoder.pep      FASTA   Protein    44,842  18,688,372       87    416.8   11,078
Aphidius_rhopalosiphi_trinity.fasta.transdecoder.pep    FASTA   Protein    40,022  14,313,010       85    357.6   11,722
Harpalus_affinis_trinity.fasta.transdecoder.pep         FASTA   Protein    41,824  17,869,618       86    427.3    7,821
Harpalus_eous_trinity.fasta.transdecoder.pep            FASTA   Protein    37,651  15,771,161       86    418.9    5,529
Poecilus_cupreus_trinity.fasta.transdecoder.pep         FASTA   Protein    42,557  19,020,139       87    446.9    7,179
Pterostichus_melanarius_trinity.fasta.transdecoder.pep  FASTA   Protein    47,305  18,248,733       85    385.8    8,424
Zabrus_tenebrioides_trinity.fasta.transdecoder.pep      FASTA   Protein    34,677  14,622,474       85    421.7    8,600

```



## Run BUSCO

```{bash, eval=FALSE}
#!/bin/bash
#SBATCH --job-name=busco
#SBATCH --partition=medium
#SBATCH --time 00:30:00
#SBATCH --cpus-per-task=8
#SBATCH --mem=8G
#SBATCH --mail-type=END
#SBATCH --mail-user=johan.zicola@uni-goettingen.de
#SBATCH --output=busco.%J.out
#SBATCH --error=busco.%J.err

module load busco/5.4.3

busco -i $TRINITY_FILE -m transcriptome --lineage_dataset insecta_odb10 --cpu 8 --out busco_output/$SPECIES

```


```{bash, eval=FALSE}

cd /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinity_output/summary

while read i; do
sbatch --export=TRINITY_FILE=longest_isoforms/${i}_trinity.longest_isoforms.fasta,SPECIES=$i ../../06_busco.sh
done < species.txt

sbatch --export=TRINITY_FILE=longest_isoforms/Pterostichus_melanarius_trinity.longest_isoforms.fasta,SPECIES=Pterostichus_melanarius ../../06_busco.sh

```


## Separate run for Aphidius_rhopalosiphi

Do a second run for Aphidius which is a hymenoptera

hymenoptera_odb10

```{bash, eval=FALSE}
srun --time=00:20:00 --cpus-per-task=8 --mem=8G -p medium busco -i longest_isoforms/Aphidius_rhopalosiphi_trinity.longest_isoforms.fasta -m transcriptome --lineage_dataset hymenoptera_odb10 --cpu 8 --out busco_output/Aphidius_rhopalosiphi_hymenoptera_odb10 &

```



## Summarize BUSCO results

```{bash, eval=FALSE}
cd /usr/users/npghpc/data/NGS_data/BGI_data/F22FTSEUHT1495_LIBgsmfR_judi_8_insects/trinity_output/summary/busco_output/

mkdir my_summaries

find . -maxdepth 2 -name "short_summary*txt" -exec cp {} my_summaries/ \;

python /usr/product/bioinfo/BUSCO/5.4.3/bin/generate_plot.py -wd my_summaries

```

Run R code from `my_summaries/busco_figure.R`:

```{r}
######################################
#
# BUSCO summary figure
# @version 4.0.0
# @since BUSCO 2.0.0
#
# Copyright (c) 2016-2022, Evgeny Zdobnov (ez@ezlab.org)
# Licensed under the MIT license. See LICENSE.md file.
#
######################################

# Load the required libraries
library(ggplot2)
library("grid")

# !!! CONFIGURE YOUR PLOT HERE !!!
# Output
my_output <- paste("my_summaries/","busco_figure.png",sep="/")
my_width <- 20
my_height <- 15
my_unit <- "cm"

# Colors
my_colors <- c("#56B4E9", "#3492C7", "#F0E442", "#F04442")
# Bar height ratio
my_bar_height <- 0.75

# Legend
my_title <- "BUSCO Assessment Results"

# Font
my_family <- "sans"
my_size_ratio <- 1

# !!! SEE YOUR DATA HERE !!!
# Your data as generated by python, remove or add more
my_species <- c('Anchomenus_dorsalis', 'Anchomenus_dorsalis', 'Anchomenus_dorsalis', 'Anchomenus_dorsalis', 'Pterostichus_melanarius', 'Pterostichus_melanarius', 'Pterostichus_melanarius', 'Pterostichus_melanarius', 'Aphidius_rhopalosiphi_hymenoptera_odb10', 'Aphidius_rhopalosiphi_hymenoptera_odb10', 'Aphidius_rhopalosiphi_hymenoptera_odb10', 'Aphidius_rhopalosiphi_hymenoptera_odb10', 'Aphidius_rhopalosiphi', 'Aphidius_rhopalosiphi', 'Aphidius_rhopalosiphi', 'Aphidius_rhopalosiphi', 'Amara_similata', 'Amara_similata', 'Amara_similata', 'Amara_similata', 'Zabrus_tenebrioides', 'Zabrus_tenebrioides', 'Zabrus_tenebrioides', 'Zabrus_tenebrioides', 'Poecilus_cupreus', 'Poecilus_cupreus', 'Poecilus_cupreus', 'Poecilus_cupreus', 'Harpalus_eous', 'Harpalus_eous', 'Harpalus_eous', 'Harpalus_eous', 'Harpalus_affinis', 'Harpalus_affinis', 'Harpalus_affinis', 'Harpalus_affinis')
my_species <- factor(my_species)
my_species <- factor(my_species,levels(my_species)[c(length(levels(my_species)):1)]) # reorder your species here just by changing the values in the vector :
my_percentage <- c(84.9, 13.1, 1.0, 1.0, 91.1, 5.6, 1.5, 1.8, 74.1, 1.6, 7.7, 16.6, 91.1, 3.0, 2.3, 3.6, 96.0, 0.7, 1.1, 2.2, 93.8, 2.9, 0.8, 2.5, 94.7, 3.1, 0.7, 1.5, 94.3, 1.9, 1.5, 2.3, 95.4, 1.7, 1.0, 1.9)
my_values <- c(1160, 179, 14, 14, 1246, 76, 20, 25, 4441, 96, 459, 995, 1245, 41, 32, 49, 1313, 9, 15, 30, 1282, 40, 11, 34, 1295, 43, 10, 19, 1289, 26, 20, 32, 1304, 23, 14, 26)

######################################
######################################
######################################
# Code to produce the graph
labsize = 1
if (length(levels(my_species)) > 10){
 labsize = 0.66
}
print("Plotting the figure ...")
category <- c(rep(c("S","D","F","M"),c(1)))
category <-factor(category)
category = factor(category,levels(category)[c(4,1,2,3)])
df = data.frame(my_species,my_percentage,my_values,category)

figure <- ggplot() +

  geom_bar(aes(y = my_percentage, x = my_species, fill = category), position = position_stack(reverse = TRUE), data = df, stat="identity", width=my_bar_height) +
  coord_flip() +
  theme_gray(base_size = 8) +
  scale_y_continuous(labels = c("0","20","40","60","80","100"), breaks = c(0,20,40,60,80,100)) +
  scale_fill_manual(values = my_colors,labels =c(" Complete (C) and single-copy (S)  ",
                                                 " Complete (C) and duplicated (D)",
                                                 " Fragmented (F)  ",
                                                 " Missing (M)")) +
  ggtitle(my_title) +
  xlab("") +
  ylab("\n%BUSCOs") +

  theme(plot.title = element_text(family=my_family, hjust=0.5, colour = "black", size = rel(2.2)*my_size_ratio, face = "bold")) +
  theme(legend.position="top",legend.title = element_blank()) +
  theme(legend.text = element_text(family=my_family, size = rel(1.2)*my_size_ratio)) +
  theme(panel.background = element_rect(color="#FFFFFF", fill="white")) +
  theme(panel.grid.minor = element_blank()) +
  theme(panel.grid.major = element_blank()) +
  theme(axis.text.y = element_text(family=my_family, colour = "black", size = rel(1.66)*my_size_ratio)) +
  theme(axis.text.x = element_text(family=my_family, colour = "black", size = rel(1.66)*my_size_ratio)) +
  theme(axis.line = element_line(size=1*my_size_ratio, colour = "black")) +
  theme(axis.ticks.length = unit(.85, "cm")) +
  theme(axis.ticks.y = element_line(colour="white", size = 0)) +
  theme(axis.ticks.x = element_line(colour="#222222")) +
  theme(axis.ticks.length = unit(0.4, "cm")) +
  theme(axis.title.x = element_text(family=my_family, size=rel(1.2)*my_size_ratio)) +

  guides(fill = guide_legend(override.aes = list(colour = NULL))) +
  guides(fill=guide_legend(nrow=2,byrow=TRUE))

  for(i in rev(c(1:length(levels(my_species))))){
    detailed_values <- my_values[my_species==my_species[my_species==levels(my_species)[i]]]
    total_buscos <- sum(detailed_values)
    figure <- figure +
    annotate("text", label=paste("C:", detailed_values[1] + detailed_values[2], " [S:", detailed_values[1], ", D:", detailed_values[2], "], F:", detailed_values[3], ", M:", detailed_values[4], ", n:", total_buscos, sep=""),
             y=3, x = i, size = labsize*4*my_size_ratio, colour = "black", hjust=0, family=my_family)
  }

figure


```

Using the hymenoptera database for _Aphidium rhopalosiphi_ yields to lower completedness than while using insecta database. It can be due to the fact that we use only the adult stage and that only a subset of genes are expressed in the adult analyzed. 


Without hymenoptera analysis

```{bash, eval=FALSE}
rm my_summaries/short_summary.specific.hymenoptera_odb10.Aphidius_rhopalosiphi_hymenoptera_odb10.txt

python /usr/product/bioinfo/BUSCO/5.4.3/bin/generate_plot.py -wd my_summaries
```

```{r}
######################################
#
# BUSCO summary figure
# @version 4.0.0
# @since BUSCO 2.0.0
#
# Copyright (c) 2016-2022, Evgeny Zdobnov (ez@ezlab.org)
# Licensed under the MIT license. See LICENSE.md file.
#
######################################

# Load the required libraries
library(ggplot2)
library("grid")

# !!! CONFIGURE YOUR PLOT HERE !!!
# Output
my_output <- paste("my_summaries/","busco_figure.png",sep="/")
my_width <- 20
my_height <- 15
my_unit <- "cm"

# Colors
my_colors <- c("#56B4E9", "#3492C7", "#F0E442", "#F04442")
# Bar height ratio
my_bar_height <- 0.75

# Legend
my_title <- "BUSCO Assessment Results"

# Font
my_family <- "sans"
my_size_ratio <- 1

# !!! SEE YOUR DATA HERE !!!
# Your data as generated by python, remove or add more
my_species <- c('Anchomenus_dorsalis', 'Anchomenus_dorsalis', 'Anchomenus_dorsalis', 'Anchomenus_dorsalis', 'Pterostichus_melanarius', 'Pterostichus_melanarius', 'Pterostichus_melanarius', 'Pterostichus_melanarius', 'Aphidius_rhopalosiphi', 'Aphidius_rhopalosiphi', 'Aphidius_rhopalosiphi', 'Aphidius_rhopalosiphi', 'Amara_similata', 'Amara_similata', 'Amara_similata', 'Amara_similata', 'Zabrus_tenebrioides', 'Zabrus_tenebrioides', 'Zabrus_tenebrioides', 'Zabrus_tenebrioides', 'Poecilus_cupreus', 'Poecilus_cupreus', 'Poecilus_cupreus', 'Poecilus_cupreus', 'Harpalus_eous', 'Harpalus_eous', 'Harpalus_eous', 'Harpalus_eous', 'Harpalus_affinis', 'Harpalus_affinis', 'Harpalus_affinis', 'Harpalus_affinis')
my_species <- factor(my_species)
my_species <- factor(my_species,levels(my_species)[c(length(levels(my_species)):1)]) # reorder your species here just by changing the values in the vector :
my_percentage <- c(84.9, 13.1, 1.0, 1.0, 91.1, 5.6, 1.5, 1.8, 91.1, 3.0, 2.3, 3.6, 96.0, 0.7, 1.1, 2.2, 93.8, 2.9, 0.8, 2.5, 94.7, 3.1, 0.7, 1.5, 94.3, 1.9, 1.5, 2.3, 95.4, 1.7, 1.0, 1.9)
my_values <- c(1160, 179, 14, 14, 1246, 76, 20, 25, 1245, 41, 32, 49, 1313, 9, 15, 30, 1282, 40, 11, 34, 1295, 43, 10, 19, 1289, 26, 20, 32, 1304, 23, 14, 26)

######################################
######################################
######################################
# Code to produce the graph
labsize = 1
if (length(levels(my_species)) > 10){
 labsize = 0.66
}
print("Plotting the figure ...")
category <- c(rep(c("S","D","F","M"),c(1)))
category <-factor(category)
category = factor(category,levels(category)[c(4,1,2,3)])
df = data.frame(my_species,my_percentage,my_values,category)

figure <- ggplot() +

  geom_bar(aes(y = my_percentage, x = my_species, fill = category), position = position_stack(reverse = TRUE), data = df, stat="identity", width=my_bar_height) +
  coord_flip() +
  theme_gray(base_size = 8) +
  scale_y_continuous(labels = c("0","20","40","60","80","100"), breaks = c(0,20,40,60,80,100)) +
  scale_fill_manual(values = my_colors,labels =c(" Complete (C) and single-copy (S)  ",
                                                 " Complete (C) and duplicated (D)",
                                                 " Fragmented (F)  ",
                                                 " Missing (M)")) +
  ggtitle(my_title) +
  xlab("") +
  ylab("\n%BUSCOs") +

  theme(plot.title = element_text(family=my_family, hjust=0.5, colour = "black", size = rel(2.2)*my_size_ratio, face = "bold")) +
  theme(legend.position="top",legend.title = element_blank()) +
  theme(legend.text = element_text(family=my_family, size = rel(1.2)*my_size_ratio)) +
  theme(panel.background = element_rect(color="#FFFFFF", fill="white")) +
  theme(panel.grid.minor = element_blank()) +
  theme(panel.grid.major = element_blank()) +
  theme(axis.text.y = element_text(family=my_family, colour = "black", size = rel(1.66)*my_size_ratio)) +
  theme(axis.text.x = element_text(family=my_family, colour = "black", size = rel(1.66)*my_size_ratio)) +
  theme(axis.line = element_line(size=1*my_size_ratio, colour = "black")) +
  theme(axis.ticks.length = unit(.85, "cm")) +
  theme(axis.ticks.y = element_line(colour="white", size = 0)) +
  theme(axis.ticks.x = element_line(colour="#222222")) +
  theme(axis.ticks.length = unit(0.4, "cm")) +
  theme(axis.title.x = element_text(family=my_family, size=rel(1.2)*my_size_ratio)) +

  guides(fill = guide_legend(override.aes = list(colour = NULL))) +
  guides(fill=guide_legend(nrow=2,byrow=TRUE))

  for(i in rev(c(1:length(levels(my_species))))){
    detailed_values <- my_values[my_species==my_species[my_species==levels(my_species)[i]]]
    total_buscos <- sum(detailed_values)
    figure <- figure +
    annotate("text", label=paste("C:", detailed_values[1] + detailed_values[2], " [S:", detailed_values[1], ", D:", detailed_values[2], "], F:", detailed_values[3], ", M:", detailed_values[4], ", n:", total_buscos, sep=""),
             y=3, x = i, size = labsize*4*my_size_ratio, colour = "black", hjust=0, family=my_family)
  }

figure

#ggsave(figure, file=my_output, width = my_width, height = my_height, unit = my_unit)
#print("Done")

```

We see a higher rate of duplication for _Anchoenus dorsalis_.


# ##################################

# Off-target analysis

This job will be done by the partner company Eupheria Biotech GmbH (https://www.fisaonline.de/einrichtungen-finden/details/?tx_fisaresearch_institutions%5Baction%5D=institutionDetails&tx_fisaresearch_institutions%5Bcontroller%5D=Institutions&tx_fisaresearch_institutions%5Bi_id%5D=5804&cHash=f826a8e38b4decbcd499614c8e5f051e#more)



Use Doga's script (https://figshare.com/articles/dataset/_i_in_silico_i_off-target_analysis_of_lethal_genes_in_cabbage_stem_flea_beetle/25283683?file=44671225) (scripts/off_target_prediction_lethal.py)

# COI1 validation


Di mentioned that to verify if we have the correct species, one needs to look at the cytochrome oxidase subunit 1 (COI) sequence.

01_Ad
Anchomenus dorsalis (NCBI taxonomy ID 586010)
https://www.ncbi.nlm.nih.gov/nuccore/HQ953549.1

https://www.ncbi.nlm.nih.gov/Taxonomy/Browser/wwwtax.cgi?id=586010

02_Pt
Pterostichus melanarius (NCBI taxonomy ID 60768)
https://www.ncbi.nlm.nih.gov/nuccore/MH020387.1

03_Zt
Zabrus tenebrioides (NCBI taxonomy ID 272042)
https://www.ncbi.nlm.nih.gov/nuccore/ON623806.2

04_Asp
Amara similata (=Harpalus similatus) (NCBI taxonomy ID 272013)
https://www.ncbi.nlm.nih.gov/nuccore/MH020396.1

05_G
Harpalus affinis (NCBI taxonomy ID 247283)
https://www.ncbi.nlm.nih.gov/nuccore/OK638137.1

06_Ol
Harpalus eous (NCBI taxonomy ID 2107211)
https://www.ncbi.nlm.nih.gov/nuccore/LC764887.1

07_Pr
Poecilus cupreus (=Pterostichus cupreus) (NCBI taxonomy ID 270614)
https://www.ncbi.nlm.nih.gov/nuccore/MN454673.1

08_Ar
Aphidius rhopalosiphi (NCBI taxonomy ID 55891)
https://www.ncbi.nlm.nih.gov/nuccore/JN164779.1


